---
title: "Bay Scallop Larve - D Stage Abnormality Rates"
author: "Katie McFarland"
date: "12/13/2022"
output: html_document
---

# 20231217 - started manuscript Github repostiory 'EAD-ASEB-Airradians_Larvae_crossgen_OA'
# SGurr edits
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
# knitr::opts_knit$set(root.dir = "C:/Users/katherine.mcfarland/Documents/GitHub/EAD-ASEB-Airradians_multigen_OA/larvae") # Katie's
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis") # Sam's
```

```{r, echo=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(rcompanion)
library(FSA)
library(car)
library(forcats)
library(kableExtra) # nice Rmd tables
library(emmeans)
library(ggpubr)
```

### Load Data as df
```{r}
#Here we are assing our data fram the name "df" so that we do not have to type out the full file name everytime we want to call it.
df<-read.csv("Data/Dstage.data.csv", header = T) %>% 
  # change to numeric for the DV and as factor for IDVs
  dplyr::mutate(Percent.deformed = as.numeric(Percent.deformed),
                Rep              = as.factor(Rep),
                Year             = as.factor(Year),
                Run              = as.factor(Run)) %>% 
  
  # rename the variables for consistancy using 'case_when'
  dplyr::mutate(Parent = case_when(Parent ==  "LOW" ~ "Low pCO2", 
                                   Parent ==  "MODERATE" ~ "Moderate pCO2", 
                                   Parent ==  "HIGH" ~ "High pCO2")) %>% 
  dplyr::mutate(Treatment = case_when(Treatment ==  "Low OA" ~ "Low pCO2", 
                                      Treatment ==  "Moderate OA" ~ "Moderate pCO2", 
                                      Treatment ==  "High OA" ~ "High pCO2")) %>% 
  
  # change the levels for order when plotting - call as factors too!
  dplyr::mutate(Parent = as.factor(forcats::fct_relevel(Parent, 
                                                        c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  dplyr::mutate(Treatment = as.factor(forcats::fct_relevel(Treatment, 
                                                           c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  
  # rename them
  dplyr::rename(Parent_pCO2 = Parent) %>% 
  dplyr::rename(Larvae_pCO2  = Treatment) # done 


#to look at columns names of your data frame
head(df)

#Sturcture of the data - to check that variables are properly assigned to facotr or variable
str(df)
```
# About data 
### Summary: 

  - (1) F1 larvae under three OA levels, *no grow out*

  - (2) F1 larvae under three OA levels, *grown out to adult stage*

  - (3) F2 larvae full cross 3 x 3 parent x offspring OA, *no grow out*

  - (4) F2 larvae under two OA levels, *grown out to adult stage*

  - (5) F2 larvae under 3 x 3 parent x offspring OA, *no grow out*

  - (6) F3 larvae under 3 x 3 parent x offsring OA, *no grow out past juvenile takedown of the experiment*

### * NOTE: N = 3 for Runs 3 and 6 for parental x offspring pCO2 treamtents 
```{r about data}

# number of runs 
unique(df$Run) # 1,2,3,4,6 - what are these?
unique(df$Generation) # 1, 2, 3 - generations for larval exposures 

# Lets just see our replication strength per treatment in this summary file..
Rep_Summary_all <- df %>% dplyr::select(c(Generation, Year,  Run, Parent_pCO2, Larvae_pCO2)) %>% 
                           dplyr::group_by_all() %>% 
                           dplyr::summarise(n=n())

unique(Rep_Summary_all$n) # we see 3-5 reps per treatment, veery niiiice!

# print the runs
# summary: 
# Runs 1 and 2 == Generation 1 matched parent x offspring exposure for the multigenerational rearing 
# Run 3 == Generation 2  full cross parent x offspring exposure *** TARGET FOR THIS PAPER! 
# Run 4 == Generation 2 matched parent x offspring exposure for the multigenerational rearing 
# Run 6 == Generation 3  full cross parent x offspring exposure *** TARGET FOR THIS PAPER! 

Rep_Summary_R1 <- Rep_Summary_all %>%  dplyr::filter(Run %in% 1) # larval [hase pf ] grow out multigenerational
Rep_Summary_R1 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_R2 <- Rep_Summary_all %>%  dplyr::filter(Run %in% 2) # larval phase of grow out multigenerational
Rep_Summary_R2 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")


Rep_Summary_R3 <- Rep_Summary_all %>%  dplyr::filter(Run %in% 3) #  full cross design 
Rep_Summary_R3 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")


Rep_Summary_R4 <- Rep_Summary_all %>%  dplyr::filter(Run %in% 4) # larval phase of grow out multigenerational
Rep_Summary_R4 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")


Rep_Summary_R6 <- Rep_Summary_all %>%  dplyr::filter(Run %in% 6) # full cross design - match parent x pffspring contributed to F3s juvenile grow out
Rep_Summary_R6 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

```

# Divide and conquer - we have 6 different experimens as follows:

```{r what expeirments do we have data for?}
unique(df$Run)
# 1 2 3 4 6
```
    
  
```{r call target datasets}

df_Exp1 <- df %>% filter(Run=="1") # F1 three OA 
df_Exp2 <- df %>% filter(Run=="2") # F1 three OA, repeated
df_Exp3 <- df %>% filter(Run=="3") # F2 full factorial
df_Exp4 <- df %>% filter(Run=="4") # F2 two OA 
df_Exp6 <- df %>% filter(Run=="6") # F3 - full factorial
```

# Plot data as barplots mean +- SE for each run

```{r F1 Experiment #1}


# Summarise Percent Deformities for plotting 
# Remember - we DO NOT need the Parent_pCO2 column here - just Larvae pCO2
Exp1_mean_deformaties <- df_Exp1 %>% 
                          dplyr::select(Larvae_pCO2,Percent.deformed) %>% 
                          na.omit() %>% 
                          dplyr::group_by(Larvae_pCO2) %>% 
                          dplyr::summarise(mean_Perc_def = mean(Percent.deformed), 
                                           n           = n(),
                                           sd_Perc_def   = sd(Percent.deformed),
                                           se_Perc_def   = sd_Perc_def/(sqrt(n)))
# kable read the head of this for pdf knit
Exp1_mean_deformaties %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")


# Barplots
Exp1_barplot <- ggplot(Exp1_mean_deformaties) +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_def-se_Perc_def, 
                                           ymax=mean_Perc_def+se_Perc_def), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_def,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="Experiment #1 (F1)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Abnormalities") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=10),
                              legend.position="none") 
print(Exp1_barplot)


# Barplots
Exp1_boxplot <- ggplot(df_Exp1,
                       aes(x=Larvae_pCO2, 
                                         y=Percent.deformed,
                                         fill=factor(Larvae_pCO2)), 
                                  stat="identity") +
                        geom_boxplot() +
                        geom_jitter(width = 0.25) +
                        labs(title="Experiment #1 (F1)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Abnormalities") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                           values=c("forestgreen","orange", "purple")) +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=10),
                              legend.position="none") 
print(Exp1_boxplot)

# MEan SE
Exp1_MeanSE <- ggplot(df_Exp1,
                       aes(x=Larvae_pCO2, 
                                         y=Percent.deformed,
                                         fill=factor(Larvae_pCO2)), 
                                  stat="identity") +
                       geom_point(aes(colour = Larvae_pCO2), 
                                        position = position_dodge2(width = 0.2)) + 
                       stat_summary(fun.y="mean", size = 0.8,
                                          position = position_dodge2(width = 1)) +
                       stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)), 
                                          fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                          geom = 'errorbar', width = 0.25, size = 1,
                                          position = position_dodge2(width = 0.2),
                                    color=c("forestgreen","orange", "purple")) + 
                        labs(title="Experiment #1 (F1)", 
                            x ="pCO2 Larvae", 
                            y = "Proportion deformed") +
                        # scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                        #                    values=c("forestgreen","orange", "purple")) +
                        scale_color_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                           values=c("forestgreen","orange", "purple")) +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.title.x=element_blank(),
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=8),
                              plot.title = element_text(size=12),
                              legend.position="none") 
print(Exp1_MeanSE)
```

```{r F1 Experiment #2}


# Summarise Percent Deformities for plotting 
# Remember - we DO NOT need the Parent_pCO2 column here - just Larvae pCO2
Exp2_mean_deformaties <- df_Exp2 %>% 
                          dplyr::select(Larvae_pCO2,Percent.deformed) %>% 
                          na.omit() %>% 
                          dplyr::group_by(Larvae_pCO2) %>% 
                          dplyr::summarise(mean_Perc_def = mean(Percent.deformed), 
                                           n           = n(),
                                           sd_Perc_def   = sd(Percent.deformed),
                                           se_Perc_def   = sd_Perc_def/(sqrt(n)))
# kable read the head of this for pdf knit
Exp2_mean_deformaties %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")


# Barplots
Exp2_barplot <- ggplot(Exp2_mean_deformaties) +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_def-se_Perc_def, 
                                           ymax=mean_Perc_def+se_Perc_def), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_def,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="Experiment #1 (F1)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Abnormalities") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.title.x=element_blank(),
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=10),
                              legend.position="none") 
print(Exp2_barplot)


# Mean SE
Exp2_MeanSE <- ggplot(df_Exp2,
                       aes(x=Larvae_pCO2, 
                                         y=Percent.deformed,
                                         fill=factor(Larvae_pCO2)), 
                                  stat="identity") +
                       geom_point(aes(colour = Larvae_pCO2), 
                                        position = position_dodge2(width = 0.2)) + 
                       stat_summary(fun.y="mean", size = 0.8,
                                          position = position_dodge2(width = 1)) +
                       stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)), 
                                          fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                          geom = 'errorbar', width = 0.25, size = 1,
                                          position = position_dodge2(width = 0.2),
                                    color=c("forestgreen","orange", "purple")) + 
                        labs(title="Experiment #2 (F1)", 
                            x ="pCO2 Larvae", 
                            y = "Percent Abnormalities") +
                        # scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                        #                    values=c("forestgreen","orange", "purple")) +
                        scale_color_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                           values=c("forestgreen","orange", "purple")) +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(),
                              axis.title.y=element_blank(),
                              axis.title.x=element_blank(),
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=8),
                              plot.title = element_text(size=12),
                              legend.position="none") 
print(Exp2_MeanSE)
```

```{r F2 Experiment #3}

# Summarise Percent Deformities for plotting 
Exp3_mean_deformaties <- df_Exp3 %>% 
                          dplyr::select(Parent_pCO2,Larvae_pCO2,Percent.deformed) %>% 
                          na.omit() %>% 
                          dplyr::group_by(Parent_pCO2,Larvae_pCO2) %>% 
                          dplyr::summarise(mean_Perc_def = mean(Percent.deformed), 
                                           n           = n(),
                                           sd_Perc_def   = sd(Percent.deformed),
                                           se_Perc_def   = sd_Perc_def/(sqrt(n)))

# Barplots
Exp3_barplot <- ggplot(Exp3_mean_deformaties) +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_def-se_Perc_def, 
                                           ymax=mean_Perc_def+se_Perc_def), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_def,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="Experiment #3 (F2)", 
                            x ="pCO2 Larvae", 
                            y = "Percent Abnormalities") +
                        scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=8),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

print(Exp3_barplot)


# Mean SE
Exp3_MeanSE <- ggplot(df_Exp3,aes(x=Larvae_pCO2, 
                                 y=Percent.deformed, 
                                 colour=Larvae_pCO2)) +
                             # scale_linetype(c("dotted","solid")) +
                        scale_colour_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        geom_point(aes(colour = Larvae_pCO2), 
                                        position = position_dodge2(width = 0.2)) + 
                        stat_summary(fun.y="mean", size = 0.8, color = "black",
                                          position = position_dodge2(width = 1)) +
                        stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)), 
                                          fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                          geom = 'errorbar', width = 0.25, size = 1,
                                          position = position_dodge2(width = 1)) +
                        labs(title="Experiment #3 (F2)", 
                            x ="pCO2 Larvae", 
                            y = "Percent Abnormalities") +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.title.y=element_blank(),
                              axis.title.x=element_blank(),
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=8),
                              plot.title = element_text(size=12),
                              legend.position="none")  +
                        facet_wrap(~Parent_pCO2)
print(Exp3_MeanSE)

```

```{r F2 Experiment #4}


# Summarise Percent Deformities for plotting 
# Remember - we DO NOT need the Parent_pCO2 column here - just Larvae pCO2
Exp4_mean_deformaties <- df_Exp4 %>% 
                          dplyr::select(Larvae_pCO2,Percent.deformed) %>% 
                          na.omit() %>% 
                          dplyr::group_by(Larvae_pCO2) %>% 
                          dplyr::summarise(mean_Perc_def = mean(Percent.deformed), 
                                           n           = n(),
                                           sd_Perc_def   = sd(Percent.deformed),
                                           se_Perc_def   = sd_Perc_def/(sqrt(n)))
# kable read the head of this for pdf knit
Exp4_mean_deformaties %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")


# Barplots
Exp4_barplot <- ggplot(Exp4_mean_deformaties) +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_def-se_Perc_def, 
                                           ymax=mean_Perc_def+se_Perc_def), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_def,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="Experiment #4 (F2)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Abnormalities") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=10),
                              legend.position="none") 
print(Exp4_barplot)


# Mean SE
Exp4_MeanSE <- ggplot(df_Exp4,
                       aes(x=Larvae_pCO2, 
                                         y=Percent.deformed,
                                         fill=factor(Larvae_pCO2)), 
                                  stat="identity") +
                       geom_point(aes(colour = Larvae_pCO2), 
                                        position = position_dodge2(width = 0.2)) + 
                       stat_summary(fun.y="mean", size = 0.8,
                                          position = position_dodge2(width = 1)) +
                       stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)), 
                                          fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                          geom = 'errorbar', width = 0.25, size = 1,
                                          position = position_dodge2(width = 0.2),
                                    color=c("forestgreen","orange")) + 
                        labs(title="Experiment #4 (F2)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Proportion deformed") +
                        # scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                        #                    values=c("forestgreen","orange", "purple")) +
                        scale_color_manual(breaks=c("Low pCO2", "Moderate pCO2"), 
                                           values=c("forestgreen","orange")) +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=10),
                              plot.title = element_text(size=12),
                              legend.position="none") 
print(Exp4_MeanSE)
```

```{r F3 Experiment #6}

# Summarise Percent Deformities for plotting 
Exp6_mean_deformaties <- df_Exp6 %>% 
                          dplyr::select(Parent_pCO2,Larvae_pCO2,Percent.deformed) %>% 
                          na.omit() %>% 
                          dplyr::group_by(Parent_pCO2,Larvae_pCO2) %>% 
                          dplyr::summarise(mean_Perc_def = mean(Percent.deformed), 
                                           n           = n(),
                                           sd_Perc_def   = sd(Percent.deformed),
                                           se_Perc_def   = sd_Perc_def/(sqrt(n)))

# Barplots
Exp6_barplot <- ggplot(Exp6_mean_deformaties) +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_def-se_Perc_def, 
                                           ymax=mean_Perc_def+se_Perc_def), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_def,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="Experiment #6 (F3)", 
                            x ="pCO2 Larvae", 
                            y = "Percent Abnormalities") +
                        scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=8),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

print(Exp6_barplot)


# Mean SE
Exp6_MeanSE <- ggplot(df_Exp6,aes(x=Larvae_pCO2, 
                                 y=Percent.deformed, 
                                 colour=Larvae_pCO2)) +
                             # scale_linetype(c("dotted","solid")) +
                        scale_colour_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        geom_point(aes(colour = Larvae_pCO2), 
                                        position = position_dodge2(width = 0.2)) + 
                        stat_summary(fun.y="mean", size = 0.8, color = "black",
                                          position = position_dodge2(width = 1)) +
                        stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)), 
                                          fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                          geom = 'errorbar', width = 0.25, size = 1,
                                          position = position_dodge2(width = 1)) +
                        labs(title="Experiment #6 (F3)", 
                            x ="pCO2 Larvae", 
                            y = "Percent Abnormalities") +
                        scale_x_discrete(labels=c("L", "M", "H")) +
                        scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.title.y=element_blank(),
                              axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=8),
                              plot.title = element_text(size=12),
                              legend.position="none")  +
                        facet_wrap(~Parent_pCO2)
print(Exp6_MeanSE)

```

```{r Plots ggarrange them all}

# grid arrange plots for ouput 

library(ggplot2)
library(gridExtra)
library(grid)

Exp1_MeanSE
Exp2_MeanSE
Exp3_MeanSE
Exp4_MeanSE
Exp6_MeanSE

pdf("Output/DStage_abnormalities/Percent_Abnormalities.pdf", width=8, height=6)
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 4))) # 3 rows, 5 columns
grid.text("Percent Abnormalities", vp = viewport(layout.pos.row = 1, layout.pos.col = 1:2))
print(Exp1_MeanSE, vp = vplayout(1, 1))  
print(Exp2_MeanSE, vp = vplayout(1, 2))  
print(Exp3_MeanSE, vp = vplayout(1, 3:4))   
print(Exp4_MeanSE, vp = vplayout(2, 2))
print(Exp6_MeanSE, vp = vplayout(2, 3:4))
dev.off()

```


# Statistics 

## Generation 2
### * Full model: Percent.deformed ~ Parent_pCO2 * Larvae_pCO2
```{r Gen 2 Full model} 
library(emmeans)
aovMOD_Gen2 <- lm(Percent.deformed ~ Parent_pCO2 * Larvae_pCO2, data=df_Gen2)
shapiro.test(resid(aovMOD_Gen2)) # 0.07377 - norm
leveneTest(aovMOD_Gen2) # 0.3222- pass
summary(aov(aovMOD_Gen2))
#                         Df Sum Sq Mean Sq F value   Pr(>F)    
# Parent_pCO2              2  0.036  0.0180   2.157    0.145    
# Larvae_pCO2              2  3.546  1.7730 212.856 2.98e-13 ***
# Parent_pCO2:Larvae_pCO2  4  0.010  0.0025   0.304    0.872    
# Residuals               18  0.150  0.0083
aovMOD_Gen2_posthoc <- emmeans(aovMOD_Gen2, pairwise~Larvae_pCO2, adjust="tukey")
multcomp::cld(aovMOD_Gen2_posthoc$emmeans,alpha = 0.5, Letters = letters)
# Larvae_pCO2   emmean     SE df lower.CL upper.CL .group
# Low pCO2       0.163 0.0304 18   0.0993    0.227  a    
# Moderate pCO2  0.193 0.0304 18   0.1293    0.257  a    
# High pCO2      0.947 0.0304 18   0.8826    1.010   b

(0.947- 0.163) * 100 # 78.4 % more mortality under high than low pCO2, regardless of parental history
(0.947- 0.193) * 100 # 75.4 % more mortality under high than moderate pCO2, regardless of parental history
```

### * Partial models (within history): Percent.deformed ~  Larvae_pCO2 (run for each history)
### * why do this? - we see that high exposure had an overarching effect on percent deformities when looking at all data, was a particular history affected by this encounter significantly as to others? Technically this should come out via an interaction but still of merit to explore here
```{r Gen 2 within history} 

# dataframes filterered for history 
df_Gen2_LOWhist  <- df_Gen2 %>%  dplyr::filter(Parent_pCO2 == "Low pCO2")
df_Gen2_MODhist  <- df_Gen2 %>%  dplyr::filter(Parent_pCO2 == "Moderate pCO2")
df_Gen2_HIGHhist <- df_Gen2 %>%  dplyr::filter(Parent_pCO2 == "High pCO2")

# Run mods 
aovMOD_Gen2_LOW  <- lm(Percent.deformed ~ Larvae_pCO2, data=df_Gen2_LOWhist)
aovMOD_Gen2_MOD  <- lm(Percent.deformed ~ Larvae_pCO2, data=df_Gen2_MODhist)
aovMOD_Gen2_HIGH <- lm(Percent.deformed ~ Larvae_pCO2, data=df_Gen2_HIGHhist)

# check assumptions
shapiro.test(resid(aovMOD_Gen2_LOW)) # 0.5787 - norm
leveneTest(aovMOD_Gen2_LOW)  # 0.9353 - pass

shapiro.test(resid(aovMOD_Gen2_MOD)) # 0.5217 - norm
leveneTest(aovMOD_Gen2_MOD) # 0.4866 - pass

shapiro.test(resid(aovMOD_Gen2_HIGH)) # 0.01795 - non norm - run non parametric Kruskal Wallis
leveneTest(aovMOD_Gen2_HIGH) # 0.1122 - pass

# non parametrics (if needed)
aovKW_Gen2_HIGH <- kruskal.test(Percent.deformed ~ Larvae_pCO2, data=df_Gen2_HIGHhist)

# Summary models 
summary(aov(aovMOD_Gen2_LOW)) 
#             Df Sum Sq Mean Sq F value   Pr(>F)    
# Larvae_pCO2  2 1.3093  0.6547   282.2 1.16e-06 ***
# Residuals    6 0.0139  0.0023

summary(aov(aovMOD_Gen2_MOD))
#             Df Sum Sq Mean Sq F value   Pr(>F)    
# Larvae_pCO2  2 1.0260  0.5130    44.9 0.000246 ***
# Residuals    6 0.0686  0.0114

aovKW_Gen2_HIGH
# Kruskal-Wallis chi-squared = 5.6, df = 2, p-value = 0.06081

# Conclusion 
# Low and Moderate history show a significant affect of larvae pCO2 on percent deformities
# High history are unaffected by larvae pCO2

```


## Generation 3
### * Full model: Percent.deformed ~ Parent_pCO2 * Larvae_pCO2
```{r Gen 3 Full model} 
library(emmeans)
aovMOD_Gen3 <- lm(Percent.deformed ~ Parent_pCO2 * Larvae_pCO2, data=df_Gen3)
shapiro.test(resid(aovMOD_Gen3)) # 0.8472 - norm
leveneTest(aovMOD_Gen3) # 0.4374- pass
summary(aov(aovMOD_Gen3))
#                         Df  Sum Sq Mean Sq F value  Pr(>F)   
# Parent_pCO2              2 0.01863 0.00931   0.667 0.52602   
# Larvae_pCO2              2 0.17192 0.08596   6.160 0.00973 **
# Parent_pCO2:Larvae_pCO2  4 0.12531 0.03133   2.245 0.10708   
# Residuals               17 0.23724 0.01396
aovMOD_Gen3_posthoc <- emmeans(aovMOD_Gen3, pairwise~Larvae_pCO2, adjust="tukey")
multcomp::cld(aovMOD_Gen3_posthoc$emmeans,alpha = 0.5, Letters = letters)
# Larvae_pCO2   emmean     SE df lower.CL upper.CL .group
# Low pCO2       0.312 0.0394 17    0.229    0.395  a    
# Moderate pCO2  0.445 0.0394 17    0.362    0.528   b   
# High pCO2      0.522 0.0425 17    0.432    0.612    c

(0.445- 0.312) * 100 # 13.3 % more mortality under moderate than low pCO2, regardless of parental history
(0.522- 0.312) * 100 # 21 % more mortality under high than low pCO2, regardless of parental history
(0.522- 0.445) * 100 # 7.7 % more mortality under high than moderate pCO2, regardless of parental history
```

### * Partial models (within history): Percent.deformed ~  Larvae_pCO2 (run for each history)
### * why do this? - we see that high exposure had an overarching effect on percent deformities when looking at all data, was a particular history affected by this encounter significantly as to others? Technically this should come out via an interaction but still of merit to explore here
```{r Gen 3 within history} 

# dataframes filterered for history 
df_Gen3_LOWhist  <- df_Gen3 %>%  dplyr::filter(Parent_pCO2 == "Low pCO2")
df_Gen3_MODhist  <- df_Gen3 %>%  dplyr::filter(Parent_pCO2 == "Moderate pCO2")
df_Gen3_HIGHhist <- df_Gen3 %>%  dplyr::filter(Parent_pCO2 == "High pCO2")

# Run mods 
aovMOD_Gen3_LOW  <- lm(Percent.deformed ~ Larvae_pCO2, data=df_Gen3_LOWhist)
aovMOD_Gen3_MOD  <- lm(Percent.deformed ~ Larvae_pCO2, data=df_Gen3_MODhist)
aovMOD_Gen3_HIGH <- lm(Percent.deformed ~ Larvae_pCO2, data=df_Gen3_HIGHhist)

# check assumptions
shapiro.test(resid(aovMOD_Gen3_LOW)) # 0.9852 - norm
leveneTest(aovMOD_Gen3_LOW)  # 0.07482  - pass

shapiro.test(resid(aovMOD_Gen3_MOD)) # 0.6621 - norm
leveneTest(aovMOD_Gen3_MOD) # 0.7494 - pass

shapiro.test(resid(aovMOD_Gen3_HIGH)) # 0.7328 - norm
leveneTest(aovMOD_Gen3_HIGH) # 0.8681 - pass

# Summary models 
summary(aov(aovMOD_Gen3_LOW))
#             Df  Sum Sq Mean Sq F value Pr(>F)  
# Larvae_pCO2  2 0.18733 0.09366   5.115 0.0618 .
# Residuals    5 0.09156 0.01831

summary(aov(aovMOD_Gen3_MOD))
#             Df Sum Sq Mean Sq F value Pr(>F)
# Larvae_pCO2  2 0.1096 0.05481   2.769  0.141
# Residuals    6 0.1187 0.01979

summary(aov(aovMOD_Gen3_HIGH))
#             Df   Sum Sq  Mean Sq F value Pr(>F)
# Larvae_pCO2  2 0.000285 0.000142   0.032  0.969
# Residuals    6 0.026931 0.004489

# Conclusion 
# Low showed a marginal effects of pCO2 on percent deformities
# Moderate and High history are unaffected by larvae pCO2
```



```{r}
df$Percent.deformed_trans <- asin(sqrt(df$Percent.deformed))

leveneTest(Percent.deformed_trans ~ Larvae_pCO2, data=df)
shapiro.test(df$Percent.deformed_trans)
```



# Stats across genrations
Data did not meet the assumptions of normality so the non parametric Scheirer Ray Hare test was used to test for signifcance among treatments and prental exposures.
```{r}
scheirerRayHare(Percent.deformed ~ Treatment*Run,
                data = df)

scheirerRayHare(Percent.deformed ~ Treatment*Parent,
                data = df)

scheirerRayHare(Percent.deformed ~ Run*Parent,
                data = df)

dunnTest(Percent.deformed ~ Run,
                data = df,
              method="bh")
```


### Low parental exposure across generations (all larval treatmeents)
```{r}
df_low <- df%>%
  filter(Parent %in% c("LOW"))

leveneTest(Percent.deformed_trans ~ Run, data=df_low)
shapiro.test(df_low$Percent.deformed_trans)

scheirerRayHare(Percent.deformed ~ Treatment*Run,
                data = df_low)


dunnTest(Percent.deformed ~ Run,
                data = df_low,
              method="bh")
```

### Low larval exposure and Low parental exposure ("control") across generations
```{r}

df_low_low <- df_low%>%
  filter(Treatment %in% c("Low OA"))

leveneTest(Percent.deformed_trans ~ Run, data=df_low_low)
shapiro.test(df_low_low$Percent.deformed_trans)

mod <- aov(Percent.deformed_trans ~Run, data = df_low_low)
summary(mod)

TukeyHSD(mod)
```


################## 

# Broken down for each experiment 

### F1: Two runs

```{r}

df_2021 <- df%>%
  filter(Year=="2021")

ggplot(data=df_2021, aes(x=Treatment, y=Percent.deformed, fill=Treatment)) +
  geom_boxplot()+  scale_fill_manual(values=c("forestgreen","orange", "purple"))+
  theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title="Percent Deformed (2021: F1 Larvae) ", x ="Treatment", y = "Percent Deformed")+
  facet_wrap("Run")
```


```{r}
df_run1 <- df%>%
  filter(Run=="1")

leveneTest(Percent.deformed_trans ~ Treatment, data=df_run1)
shapiro.test(df_run1$Percent.deformed_trans)

df_run2 <- df%>%
  filter(Run=="2")

leveneTest(Percent.deformed_trans ~ Treatment, data=df_run2)
shapiro.test(df_run2$Percent.deformed_trans)
```


```{r}
dunnTest(Percent.deformed ~ Treatment,
                data = df_run1,
              method="bh")
dunnTest(Percent.deformed ~ Treatment,
                data = df_run2,
              method="bh")
```


```{r}
leveneTest(Percent.deformed_trans ~ Treatment, data=df_2021)
shapiro.test(df_2021$Percent.deformed_trans)
```

### F2 April 2022

```{r}
df_2022 <- df%>%
  filter(Year=="2022")

df_2022_3 <- df_2022%>%
  filter(Run=="3")

ggplot(data=df_2022_3, aes(x=Treatment, y=Percent.deformed, fill=Treatment)) +
  geom_boxplot()+  scale_fill_manual(values=c("forestgreen","orange", "purple"))+
  theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title="Percent Deformed, grouped by parentage (2022: F2 Larvae) ", x ="Treatment", y = "Percent Deformed")+
  facet_wrap("Parent")
```


```{r}
leveneTest(Percent.deformed_trans ~ Treatment, data=df_2022_3)
shapiro.test(df_2022_3$Percent.deformed_trans)
```


```{r}
scheirerRayHare(Percent.deformed ~ Treatment*Parent,
                data = df_2022_3)

dunnTest(Percent.deformed ~ Treatment,
                data = df_2022_3,
              method="bh")

dunnTest(Percent.deformed ~ Parent,
                data = df_2022_3,
              method="bh")
```


### F2 August 2022

```{r}
df_2022 <- df%>%
  filter(Year=="2022")

df_2022_4 <- df_2022%>%
  filter(Run=="4")

ggplot(data=df_2022_4, aes(x=Treatment, y=Percent.deformed, fill=Treatment)) +
  geom_boxplot()+  scale_fill_manual(values=c("forestgreen","orange", "purple"))+
  theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title="Percent Deformed (2022: F2 Larvae) ", x ="Treatment", y = "Percent Deformed")
```

```{r}

leveneTest(Percent.deformed_trans ~ Treatment, data=df_2022_4)
shapiro.test(df_2022_4$Percent.deformed_trans)
```

```{r}

res.kruskal <- kruskal.test(Percent.deformed ~ Treatment, data = df_2022_4)
res.kruskal

```


### F3 Larval April 2023

```{r}
df_6 <- df %>%
  filter(Run=="6")

ggplot(data=df_6, aes(x=Treatment, y=Percent.deformed, fill=Treatment)) +
  geom_boxplot()+  scale_fill_manual(values=c("forestgreen","orange", "purple"))+
  theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title="Percent Deformed, grouped by parentage (2023: F3 Larvae) ", x ="Treatment", y = "Percent Deformed")+
  facet_wrap("Parent")



Run6_mean_deformaties <- df_6 %>% 
                dplyr::select(Treatment,Parent,Percent.deformed) %>% 
                na.omit() %>% 
                dplyr::group_by(Treatment,Parent) %>% 
                dplyr::summarise(mean_Perc_def = mean(Percent.deformed), 
                                 n           = n(),
                                 sd_Perc_def   = sd(Percent.deformed),
                                 se_Perc_def   = sd_Perc_def/(sqrt(n)))

library(forcats)
# relevel Parent so that the facets come in the approporate order - using forcats
Run6_mean_deformaties$Parent <- forcats::fct_relevel(Run6_mean_deformaties$Parent, c("LOW", "MODERATE", "HIGH"))
ggplot(Run6_mean_deformaties) +
  geom_errorbar(aes(x=Treatment, 
                     ymin=mean_Perc_def-se_Perc_def, 
                     ymax=mean_Perc_def+se_Perc_def), 
                 width=0, # removes the horizontal line
                 colour="black", 
                 size=1) +
  geom_bar(aes(x=Treatment, y=mean_Perc_def,fill=factor(Treatment)), 
            stat="identity",
           width = 0.75,
           alpha = 0.5) +
  labs(title="F3 Abnormalities- Full cross exposure", 
      x ="pCO2 Offspring Exposure", 
      y = "Percent Abnormalities") +
   scale_fill_manual(breaks=c("Low OA", "Moderate OA", "High OA"), 
                       values=c("forestgreen","orange", "purple")) +
  theme_classic() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.text=element_text(size=12),
        legend.position="none") +
  facet_wrap(~Parent)

```


```{r, F3 run 6 stats by history}

df_6_HIGH     <- df_6 %>% dplyr::filter(Parent %in% "HIGH")
df_6_MODERATE <- df_6 %>% dplyr::filter(Parent %in% "MODERATE")
df_6_LOW      <- df_6 %>% dplyr::filter(Parent %in% "LOW")


anova_HIGH <- lm(Percent.deformed_trans ~ Treatment, data = df_6_HIGH)
shapiro.test(resid(anova_HIGH)) # 0.7455
leveneTest(anova_HIGH) # 0.865
summary(aov(anova_HIGH))
#             Df   Sum Sq  Mean Sq F value Pr(>F)
# Treatment    2 0.000305 0.000152   0.033  0.968
# Residuals    6 0.027538 0.004590

anova_MODERATE <- lm(Percent.deformed_trans ~ Treatment, data = df_6_MODERATE)
shapiro.test(resid(anova_MODERATE)) # 0.5231
leveneTest(anova_MODERATE) # 0.7784
summary(aov(anova_MODERATE))
#             Df Sum Sq Mean Sq F value Pr(>F)
# Treatment    2 0.1244 0.06218   2.761  0.141
# Residuals    6 0.1351 0.02252

anova_LOW <- lm(Percent.deformed_trans ~ Treatment, data = df_6_LOW)
shapiro.test(resid(anova_LOW)) # 0.9679
leveneTest(anova_LOW) # 0.102
summary(aov(anova_LOW))
#             Df Sum Sq Mean Sq F value Pr(>F)  
# Treatment    2 0.2222 0.11108   5.188 0.0603 .
# Residuals    5 0.1071 0.02141

```


```{r}
leveneTest(Percent.deformed_trans ~ Treatment, data=df_6)
shapiro.test(df_6$Percent.deformed_trans)
```

```{r}
mod <- aov(Percent.deformed_trans ~Treatment * Parent, data = df_6)
summary(mod)
```


```{r}
scheirerRayHare(Percent.deformed ~ Treatment*Parent,
                data = df_6)

dunnTest(Percent.deformed ~ Treatment,
                data = df_6,
              method="bh")

dunnTest(Percent.deformed ~ Parent,
                data = df_6,
              method="bh")
```



### Trimming the F3 to look at differences
```{r}

df_6_trim <- df_6%>%
  filter(Parent %in% c("MODERATE", "LOW"))

ggplot(data=df_6_trim, aes(x=Treatment, y=Percent.deformed, fill=Treatment)) +
  geom_boxplot()+  scale_fill_manual(values=c("forestgreen","orange", "purple"))+
  theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title="Percent Deformed, grouped by parentage (2023: F3 Larvae) ", x ="Treatment", y = "Percent Deformed")+
  facet_wrap("Parent")
```

```{r}
scheirerRayHare(Percent.deformed ~ Treatment*Parent,
                data = df_6_trim)

#dunnTest(Percent.deformed ~ Treatment,
#                data = df_6_trim,
 #             method="bh")

#dunnTest(Percent.deformed ~ Parent,
  #              data = df_6_trim,
   #           method="bh")
```







Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
