---
title: "Larval Growth"
author: "Katie McFarland"
date: "9/29/2023"
output: html_document
---

# 20231217 - started manuscript Github repostiory 'EAD-ASEB-Airradians_Larvae_crossgen_OA'
# 20240616 - RM anova chunks
# SGurr edits
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
#knitr::opts_knit$set(root.dir = "C:/Users/katherine.mcfarland/Documents/GitHub/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis") # Katie's
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis") # Sam's
#knitr::opts_knit$set(root.dir = "C:/Users/samuel.gurr/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis") # Sam's work
```


```{r, echo=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(rcompanion)
library(FSA)
library(car)
library(forcats)
library(kableExtra) # nice Rmd tables install.packages('kableExtra')
library(emmeans)
library(ggpubr)
library(Rmisc)
library(rstatix)
```


# Load Larval Growth Data

  * 'df' -r reformat some variable names and such, not change in data just some housekeeping..
  
  * 'GrowthData' - just the larvae 

```{r load data format it too}

df     <- read.csv("Data/larval_growth.csv", header = T) %>% 
                dplyr::mutate(Lengh = as.numeric(Length)) %>% 
                
                # rename the variables for consistency using 'case_when'
                dplyr::mutate(Parental_OA = case_when(Parental_OA ==  "LOW" ~ "Low pCO2", 
                                                      Parental_OA ==  "MODERATE" ~ "Moderate pCO2", 
                                                      Parental_OA ==  "HIGH" ~ "High pCO2",
                                                      Parental_OA ==  "WILD" ~ "None"),
                              
                              treatment = case_when(treatment ==  "Low OA" ~ "Low pCO2", 
                                                    treatment ==  "Moderate OA" ~ "Moderate pCO2", 
                                                    treatment ==  "High OA" ~ "High pCO2"),
                              
                              Parental_OA = as.factor(forcats::fct_relevel(Parental_OA, 
                                                                           c("Low pCO2", "Moderate pCO2", "High pCO2"))),
                              treatment = as.factor(forcats::fct_relevel(treatment, 
                                                                         c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 

                # rename them
                dplyr::rename(Parent_pCO2 = Parental_OA,
                              Larvae_pCO2  = treatment,
                              Experiment = Run) # done 

GrowthData <- df  %>% filter(Stage=="larvae")

str(GrowthData)
```

# About experiments: 
### Summary: 

  - (1) F1 larvae under three OA levels, *no grow out*

  - (2) F1 larvae under three OA levels, *grown out to adult stage*

  - (3) F2 larvae full cross 3 x 3 parent x offspring OA, *no grow out*

  - (4) F2 larvae under two OA levels, *grown out to adult stage*

  - (5) F2 larvae under 3 x 3 parent x offspring OA, *no grow out*

  - (6) F3 larvae under 3 x 3 parent x offsring OA, *no grow out past juvenile takedown of the experiment*
  

* view replication during the experiments

```{r replication}

# number of runs 
unique(GrowthData$Generation) # 1 and 2
unique(GrowthData$Experiment) # 2 and 4

Rep_Summary_all <- GrowthData %>% dplyr::select(c(Generation, Experiment, Age, Parent_pCO2, Larvae_pCO2)) %>% 
                       dplyr::group_by_all() %>% 
                       dplyr::summarise(n=n())

# print the runs
# summary: 

Rep_Summary_Exp2 <- Rep_Summary_all %>%  dplyr::filter(Experiment %in% 2) 
Rep_Summary_Exp2 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_Exp4 <- Rep_Summary_all %>%  dplyr::filter(Experiment %in% 4) 
Rep_Summary_Exp4 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

```


* data frames for each generation

```{r}
Exp2 <- GrowthData %>%
  filter(Experiment=="2")

Exp4 <- GrowthData %>%
  filter(Experiment=="4")

```



# Plot data as barplots mean +- SE for each run

## Manuscript 1; Expeirments 2 and 4

* length plots 

* specific growth plots - *note* we see here that the speciic growth is wonky (changes with time)
let's just go with size over time for our data report (as discussed with lead PI team on 6/2024)

**output folder** = '1_Survival_Growth'

```{r  Mean table + plots; F1 Experiment #2}

# larvae size 
Exp2_summ.length <- summarySE(Exp2, measurevar="Length", groupvars=c("Age", "Larvae_pCO2"))

Exp2_MeanSE.length <- ggplot(Exp2_summ.length, aes(x=Age, y=Length, color=Larvae_pCO2))+ geom_line()+
                      geom_point()+ 
                      geom_errorbar(aes(ymin=Length-se, ymax=Length+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      labs(title = "Exp #2", 
                           x="Age (days)", 
                           y="Shell Length (μm)") + 
                      scale_x_continuous(breaks = seq(0, 20, by = 2)) +
                      scale_y_continuous(limits=c(60,210))+
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            # axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")

print(Exp2_MeanSE.length)


# output yo stuff homie!
pdf("Output/1_Survival_Growth/size_growth/Experiment2_Mean_Length.pdf", width = 4, height = 4)
print(Exp2_MeanSE.length)
dev.off()



# specific growth rate
Exp2_summ.spGR   <- Exp2 %>%
                    dplyr::select(ID, Larvae_pCO2, Age, spGR) %>% 
                    na.omit() %>% 
                    Rmisc::summarySE(measurevar="spGR", groupvars=c("Age", "Larvae_pCO2"))

Exp2_MeanSE.spGR <- ggplot(Exp2_summ.spGR, aes(x=Age, y=spGR, color=Larvae_pCO2))+ geom_line()+
                      geom_point()+ 
                      geom_errorbar(aes(ymin=spGR-se, ymax=spGR+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      labs(title = "Exp #2", 
                           x="Age (days)", 
                           y="spGR") + 
                      # scale_x_continuous(breaks = seq(0, 20, by = 2)) +
                      # scale_y_continuous(limits=c(60,210))+
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            # axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")

print(Exp2_MeanSE.spGR)

# output yo stuff homie!
pdf("Output/1_Survival_Growth/size_growth/Experiment2_Mean_spGR.pdf", width = 4, height = 4)
print(Exp2_MeanSE.spGR)
dev.off()

```

```{r  Mean table + plots; F2 Experiment #4}
Exp4_summ <- summarySE(Exp4, measurevar="Length", groupvars=c("Age", "Larvae_pCO2"))

Exp4_MeanSE <- ggplot(Exp4_summ, aes(x=Age, y=Length, color=Larvae_pCO2))+ geom_line()+
                      geom_point()+ 
                       geom_errorbar(aes(ymin=Length-se, ymax=Length+se), width=.4,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      labs(title = "Exp #4", 
                           x="Age (days)", 
                           y="Shell Length (μm)") + 
                      scale_x_continuous(breaks = seq(0, 20, by = 2)) +
                      scale_y_continuous(limits=c(60,210))+
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            # axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")

print(Exp4_MeanSE)

# output yo stuff homie!
pdf("Output/1_Survival_Growth/size_growth/Experiment4_Mean_Length.pdf", width = 4, height = 4)
print(Exp4_MeanSE)
dev.off()


# specific growth rate
Exp4_summ.spGR   <- Exp4 %>%
                    dplyr::select(Larvae_pCO2, Age, spGR) %>% 
                    na.omit() %>% 
                    Rmisc::summarySE(measurevar="spGR", groupvars=c("Age", "Larvae_pCO2"))

Exp4_MeanSE.spGR <- ggplot(Exp4_summ.spGR, aes(x=Age, y=spGR, color=Larvae_pCO2))+ geom_line()+
                      geom_point()+ 
                      geom_errorbar(aes(ymin=spGR-se, ymax=spGR+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1"))+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      labs(title = "Exp #4", 
                           x="Age (days)", 
                           y="spGR") + 
                      # scale_x_continuous(breaks = seq(0, 20, by = 2)) +
                      # scale_y_continuous(limits=c(60,210))+
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            # axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")

print(Exp4_MeanSE.spGR)

# output yo stuff homie!
pdf("Output/1_Survival_Growth/size_growth/Experiment4_Mean_spGR.pdf", width = 4, height = 4)
print(Exp4_MeanSE.spGR)
dev.off()
```


# Statistics

## Manuscript 1; Experiments 2 and 4


**Objective:** Run two tests (discussed as team 6/4/2024)
NOTE: we can report differencesin time and size at metamorphosis but *we can not interpret it* because of the 
differences in diet betwen generations. *therefore we can only address the effects of pCO2 WITHIN generations*

  - (1) Repeated measures ANOVA
    
    - F1 for experiment #2 
    - F2 for experiment #4
  
  - (2) Size at metamorphosis 
    
    - Experiment #2 at 10 days, Experiment #4 at 14 days in age
    
**output folder** = '1_Survival_Growth'

```{r  RM Anova length; F1 Experiment #2}

# (1) Repeated measures ANOVA


# (2.1) All treatment Low, moderate, and high
Exp2_RMAnova    <-  Exp2 %>%
                     dplyr::select(ID, Larvae_pCO2, Age,Length) %>% 
                     convert_as_factor(ID, Age, Larvae_pCO2) %>% 
                     na.omit() # prGR is normmalized to the inital measurement, these are NAs, omit them

Exp2_RMAnova$ID <- factor(Exp2_RMAnova$ID)



# * summary stats, QQ
MeanSE_RM.Exp2 <- Exp2_RMAnova %>%  group_by(Age,Larvae_pCO2) %>% get_summary_stats(Length, type = "mean_sd")
qqplot_RM.Exp2 <- ggqqplot(Exp2_RMAnova, "Length", ggtheme = theme_bw()) +
                          facet_grid(Age ~ Larvae_pCO2, labeller = "label_both")

# * normality assumption of the data
Shapiro_RM.Exp2 <- Exp2_RMAnova %>%  group_by(Age, Larvae_pCO2) %>% shapiro_test(Length)

# * visualization as boxplots
bxp_RM.Exp2 <- ggboxplot(
                Exp2_RMAnova, x = "Age", y = "Length",
                color = "Larvae_pCO2", palette = "jco"
                )
# * identify outliers 
outliers_RM.Exp2 <- Exp2_RMAnova %>%
                      group_by(Larvae_pCO2, Age) %>%
                      identify_outliers(Length)

# output all of that..
capture.output(MeanSE_RM.Exp2,file="Output/1_Survival_Growth/size_growth/Experiment2_Length_MeanSE.doc")
capture.output(Shapiro_RM.Exp2,file="Output/1_Survival_Growth/size_growth/Experiment2_Length_ShapiroWilk.doc")
capture.output(outliers_RM.Exp2,file="Output/1_Survival_Growth/size_growth/Experiment2_Length_Outliers.doc")
pdf("Output/1_Survival_Growth/size_growth/Experiment2_Length_Outliers.doc", width =4, height = 10)
qqplot_RM.Exp2
dev.off()
pdf("Output/1_Survival_Growth/size_growth/Experiment2_Length_boxplot.doc", width =4, height = 10)
bxp_RM.Exp2
dev.off()


# Run statistical model 
df_Exp2_res.aov <- anova_test(data = Exp2_RMAnova, 
                              formula = Length ~ Age*Larvae_pCO2 + Error(ID/Age))
get_anova_table(df_Exp2_res.aov)
#            Effect  DFn   DFd       F        p p<.05   ges
# 1     Larvae_pCO2 2.00 10.00 102.294 2.20e-07     * 0.884
# 2             Age 1.44 14.36 220.110 7.52e-11     * 0.933
# 3 Larvae_pCO2:Age 2.87 14.36  10.723 6.45e-04     * 0.574


# Effect of treatment at each time point
# this sees at which time points there is an effect 
Exp2_RMAnova$Length <- as.numeric(Exp2_RMAnova$Length)
Exp2_RMAnova <- ungroup(Exp2_RMAnova)
one.way <- Exp2_RMAnova %>%
  dplyr::filter(!(ID %in% c(1, 6,15))) %>% 
  group_by(Age) %>%
  anova_test(dv = Length, wid = ID, within = Larvae_pCO2) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
one.way

# Pairwise comparisons between treatment groups
# this is alike a posthoc t text within time point
pwc <- Exp2_RMAnova %>%
  dplyr::filter(!(ID %in% c(1, 6,15))) %>% 
  group_by(Age) %>%
  pairwise_t_test(
    Length ~ Larvae_pCO2, paired = TRUE,
    p.adjust.method = "bonferroni"
    )
pwc

```


```{r}

posthoc.Age <- rstatix::tukey_hsd(x = Exp2_RMAnova, formula = Length ~ Age)
# 1	Age	2	4	0	9.161538	-19.23428	37.55735	8.27e-01	ns
# 2	Age	2	8	0	51.994718	24.56176	79.42768	3.60e-05	****
# 3	Age	2	10	0	64.117385	36.68442	91.55035	5.43e-07	****
# 4	Age	4	8	0	42.833179	15.40022	70.26614	7.09e-04	***
# 5	Age	4	10	0	54.955846	27.52288	82.38881	1.32e-05	****
# 6	Age	8	10	0	12.122667	-14.31240	38.55773	6.19e-01	ns

posthoc.pCO2 <- rstatix::tukey_hsd(x = Exp2_RMAnova, formula = Length ~ Larvae_pCO2) 
# 1	Larvae_pCO2	Low pCO2	Moderate pCO2	0	-2.918444	-27.70676	21.86987	9.57e-01	ns
# 2	Larvae_pCO2	Low pCO2	High pCO2	0	-48.418444	-73.20676	-23.63013	5.37e-05	****
# 3	Larvae_pCO2	Moderate pCO2	High pCO2	0	-45.500000	-70.93227	-20.06773	2.05e-04	***

posthoc <- rstatix::tukey_hsd(x = Exp2_RMAnova, formula = Length ~ Age*Larvae_pCO2) %>% 
  dplyr::filter(term %in% 'Age:Larvae_pCO2')



# write output tables and posthoc
capture.output(get_anova_table(df_Exp2_res.aov),
               file="Output/1_Survival_Growth/size_growth/Experiment2_Length_RManova.doc") 
capture.output(posthoc,
               file="Output/1_Survival_Growth/size_growth/Experiment2_Length_RManova_posthoc.doc")


RM.aov <- rstatix::anova_test(data = Exp2_RMAnova, 
                        dv = Length, 
                        wid = ID, 
                        within = c(Larvae_pCO2, Age),
                        type=1)
RM.aov
#            Effect DFn DFd       F        p p<.05   ges
# 1     Larvae_pCO2   2  10 117.735 1.12e-07     * 0.890
# 2             Age   1  10   1.904 1.98e-01       0.062
# 3             Age   3  32 218.495 2.19e-21     * 0.931
# 4 Larvae_pCO2:Age   1  10   3.525 9.00e-02       0.108
# 5 Larvae_pCO2:Age   6  32  10.416 2.13e-06     * 0.561














# (2.2) Truncate just  Low versus Moderate
Exp2_RMAnova <-  Exp2 %>%
                    dplyr::filter(!Larvae_pCO2 %in% 'High pCO2') %>% 
                    dplyr::select(ID, Larvae_pCO2, Age,Length) %>% 
                    convert_as_factor(ID, Age, Larvae_pCO2) %>% 
                    na.omit() # prGR is normmalized to the inital measurement, these are NAs, omit them

# * summary stats
Exp2_RMAnova %>%  group_by(Age,Larvae_pCO2) %>% get_summary_stats(Length, type = "mean_sd")
# mean and standard deviation
# Low pCO2	4	spGR	5	0.054	0.005
# Moderate pCO2	4	spGR	4	0.045	0.007
# High pCO2	4	spGR	4	0.047	0.038
# Low pCO2	8	spGR	5	0.105	0.016
# Moderate pCO2	8	spGR	4	0.110	0.006
# High pCO2	8	spGR	4	0.060	0.048
# Low pCO2	10	spGR	5	0.053	0.026
# Moderate pCO2	10	spGR	5	0.035	0.007
# High pCO2	10	spGR	5	0.027	0.070
ggqqplot(Exp2_RMAnova, "Length", ggtheme = theme_bw()) +
  facet_grid(Age ~ Larvae_pCO2, labeller = "label_both")

# * normality assumption of the data
Exp2_RMAnova %>%  group_by(Age, Larvae_pCO2) %>% shapiro_test(Length)
# Low pCO2	2	Length	0.8431616	0.17385249
# Moderate pCO2	2	Length	0.7850201	0.07795785
# Low pCO2	4	Length	0.8697190	0.26528773
# Moderate pCO2	4	Length	0.7803917	0.07145420
# Low pCO2	8	Length	0.9234469	0.55241544
# Moderate pCO2	8	Length	0.9388713	0.65796946
# Low pCO2	10	Length	0.9314101	0.60601793
# Moderate pCO2	10	Length	0.9142506	0.49355058


# * visualization
bxp <- ggboxplot(
                Exp2_RMAnova, x = "Age", y = "Length",
                color = "Larvae_pCO2", palette = "jco"
                )
# * identify outliers 
Exp2_RMAnova %>%
  group_by(Larvae_pCO2, Age) %>%
  identify_outliers(Length)

# Run statistical model 
df_Exp2_res.aov <- anova_test(data = Exp2_RMAnova, 
                              formula = Length ~ Age*Larvae_pCO2 + Error(ID/Age))
get_anova_table(df_Exp2_res.aov)
#            Effect  DFn   DFd       F        p p<.05   ges
# 1     Larvae_pCO2 1.00  7.00   4.193 8.00e-02       0.206
# 2             Age 1.75 12.24 640.833 6.66e-13     * 0.981
# 3 Larvae_pCO2:Age 1.75 12.24   1.082 3.61e-01       0.081


capture.output(get_anova_table(df_Exp2_res.aov),
               file="Output/1_Survival_Growth/size_growth/Experiment2_Length_RManova_LowvMod.doc") 






# Visualization: box plots with p-values

# Pairwise comparisons between treatment groups
# pwc <- Exp2_RMAnova %>%
#   dplyr::filter(!ID %in% c(6,15)) %>% 
#   group_by(Larvae_pCO2) %>%
#   pairwise_t_test(
#     Length ~ Age, paired = TRUE,
#     p.adjust.method = "bonferroni"
#     )
# pwc
# 
# 
# pwc <- pwc %>% add_xy_position(x = "Age")
# bxp + 
#   stat_pvalue_manual(pwc, tip.length = 0, hide.ns = TRUE) +
#   labs(
#     subtitle = get_test_label(df_Exp2_res.aov, detailed = TRUE),
#     caption = get_pwc_label(pwc)
#   )


```

```{r ANOVA end length; F1 Expeirent #2}
# (2) Size at metamorphosis


# (2.1) All treatment Low, moderate, and high
Exp2_Anova <-  Exp2 %>%
                    dplyr::filter(Age %in% 10) %>% 
                    dplyr::select(ID, Larvae_pCO2, Age,Length) %>% 
                    convert_as_factor(ID, Age, Larvae_pCO2) %>% 
                    na.omit() # prGR is normmalized to the inital measurement, these are NAs, omit them


# * run one way anova 
Exp2_AOV_res.mod <- lm(Length~Larvae_pCO2,data=Exp2_Anova)

# * check model assumptions 
shapiro.test(resid(Exp2_AOV_res.mod)) # 0.9314
leveneTest(Exp2_AOV_res.mod) # 0.1485

# * print the summary aov
summary(aov(Exp2_AOV_res.mod))
#              Df Sum Sq Mean Sq F value   Pr(>F)    
# Larvae_pCO2  2  17180    8590   38.02 6.41e-06 ***
# Residuals   12   2711     226

# Tukey HSD 
Exp2_AOV_res.means <- emmeans::emmeans(object = Exp2_AOV_res.mod, # run tukey
                                    pairwise ~ "Larvae_pCO2",
                                    adjust = "tukey")
Exp2_AOV_res.means <- multcomp::cld(object = Exp2_AOV_res.means$emmeans, Letters = letters) # letter display\
# Larvae_pCO2   emmean   SE df lower.CL upper.CL .group
# High pCO2        105 6.72 12     90.3      120  a    
# Moderate pCO2    169 6.72 12    154.7      184   b   
# Low pCO2         182 6.72 12    167.7      197   b 

capture.output(summary(aov(Exp2_AOV_res.mod)),
               file="Output/1_Survival_Growth/size_growth/Experiment2_Length_metamorphosis_anova.doc") 
capture.output(Exp2_AOV_res.means,
               file="Output/1_Survival_Growth/size_growth/Experiment2_Length_metamorphosis_posthoc.doc") 





# (2.2) Trancate data to just Low and Moderate 
Exp2_Anova <-  Exp2 %>%
                    dplyr::filter(Age %in% 10) %>% 
                    dplyr::filter(!Larvae_pCO2 %in% 'High pCO2') %>% # omit high run for low vs. mod
                    dplyr::select(ID, Larvae_pCO2, Age,Length) %>% 
                    convert_as_factor(ID, Age, Larvae_pCO2) %>% 
                    na.omit() # prGR is normmalized to the inital measurement, these are NAs, omit them


# * run one way anova 
Exp2_AOV_res.mod <- lm(Length~Larvae_pCO2,data=Exp2_Anova)

# * check model assumptions 
shapiro.test(resid(Exp2_AOV_res.mod)) # 0.8102
leveneTest(Exp2_AOV_res.mod) # 0.6524

# * print the summary aov
summary(aov(Exp2_AOV_res.mod))
#              Df Sum Sq Mean Sq F value Pr(>F)  
# Larvae_pCO2  1  421.6   421.6     6.5 0.0342 *
# Residuals    8  518.9    64.9

# Tukey HSD 
Exp2_AOV_res.means <- emmeans::emmeans(object = Exp2_AOV_res.mod, # run tukey
                                    pairwise ~ "Larvae_pCO2",
                                    adjust = "tukey")
Exp2_AOV_res.means <- multcomp::cld(object = Exp2_AOV_res.means$emmeans, Letters = letters) # letter display\
# Larvae_pCO2   emmean  SE df lower.CL upper.CL .group
# Moderate pCO2    169 3.6  8      161      178  a    
# Low pCO2         182 3.6  8      174      191   b


# write outputs
capture.output(summary(aov(Exp2_AOV_res.mod)),
               file="Output/1_Survival_Growth/size_growth/Experiment2_Length_metamorphosis_anova_LowvMod.doc") 
capture.output(Exp2_AOV_res.means,
               file="Output/1_Survival_Growth/size_growth/Experiment2_Length_metamorphosis_posthoc_LowvMod.doc") 
```


```{r  RM Anova length; F1 Experiment #4}

# (1) Repeated measures ANOVA

Exp4_RMAnova <-  Exp4 %>%
                    dplyr::select(ID, Larvae_pCO2, Age,Length) %>% 
                    convert_as_factor(ID, Age, Larvae_pCO2) %>% 
                    na.omit() # prGR is normmalized to the inital measurement, these are NAs, omit them

# * summary stats
Exp4_RMAnova %>%  group_by(Age,Larvae_pCO2) %>% get_summary_stats(Length, type = "mean_sd")


ggqqplot(Exp4_RMAnova, "Length", ggtheme = theme_bw()) +
  facet_grid(Age ~ Larvae_pCO2, labeller = "label_both")

# * normality assumption of the data
Exp4_RMAnova %>%  group_by(Age, Larvae_pCO2) %>% shapiro_test(Length)


# * visualization
bxp <- ggboxplot(
                Exp4_RMAnova, x = "Age", y = "Length",
                color = "Larvae_pCO2", palette = "jco"
                )
# * identify outliers 
Exp4_RMAnova %>%
  group_by(Larvae_pCO2, Age) %>%
  identify_outliers(Length)

# Run statistical model 
df_Exp4_res.aov <- anova_test(data = Exp4_RMAnova, 
                              formula = Length ~ Age*Larvae_pCO2 + Error(ID/Age))
get_anova_table(df_Exp4_res.aov)
#            Effect DFn DFd       F        p p<.05   ges
# 1     Larvae_pCO2   1   4  11.133 2.90e-02     * 0.402
# 2             Age   6  24 545.062 2.01e-24     * 0.990
# 3 Larvae_pCO2:Age   6  24   1.257 3.13e-01       0.193

posthoc <- rstatix::tukey_hsd(x = Exp4_RMAnova, formula = Length ~ Age*Larvae_pCO2) %>% 
  dplyr::filter(term %in% 'Larvae_pCO2')

# write output tables and posthoc
capture.output(get_anova_table(df_Exp4_res.aov),
               file="Output/1_Survival_Growth/size_growth/Experiment4_Length_RManova.doc") 
capture.output(posthoc,
               file="Output/1_Survival_Growth/size_growth/Experiment4_Length_RManova_posthoc.doc")

```

```{r ANOVA end length; F1 Expeirent #4}
# (2) Size at metamorphosis


# All treatment Low and moderate
Exp4_Anova <-  Exp4 %>%
                    dplyr::filter(Age %in% 10) %>% 
                    dplyr::select(ID, Larvae_pCO2, Age,Length) %>% 
                    convert_as_factor(ID, Age, Larvae_pCO2) %>% 
                    na.omit() # prGR is normmalized to the inital measurement, these are NAs, omit them


# * run one way anova 
Exp4_AOV_res.mod <- lm(Length~Larvae_pCO2,data=Exp4_Anova)

# * check model assumptions 
shapiro.test(resid(Exp4_AOV_res.mod)) # 0.2569
leveneTest(Exp4_AOV_res.mod) # 0.7101

# * print the summary aov
summary(aov(Exp4_AOV_res.mod))
#             Df Sum Sq Mean Sq F value Pr(>F)
# Larvae_pCO2  1  110.1  110.08   1.384  0.305
# Residuals    4  318.1   79.52 

capture.output(summary(aov(Exp4_AOV_res.mod)),
               file="Output/1_Survival_Growth/size_growth/Experiment4_Length_metamorphosis_anova.doc") 


```





























```{r}
################## Checking Assumptions #############
#Perform a Shapiro-Wilk test for normality ----- If p > 0.05 then data can be considered normal

library(rstatix)
shapiro.test(df_Exp2$spGR) 

df_Exp2_tests <- df_Exp2 %>%
                    # group_by(Age) %>%
                    select(Larvae_pCO2, Age,spGR) %>% 
                    na.omit()
df_Exp2_tests %>%  group_by(Age) %>% shapiro_test(spGR)
ggqqplot(df_Exp2_tests, "spGR", facet.by = "Age")

df_Exp2_res.aov <- anova_test(data = df_Exp2_tests, dv = spGR, wid = Larvae_pCO2, within = Age)
get_anova_table(res.aov)


shapiro.test(df_Exp4$spGR) 

###################Equal Variances#################
#Levene's test -----Levene's Test for Homogeneity of Variance (center = median)
library(car)
leveneTest(spGR ~ Larvae_pCO2, data=df_Exp2)
leveneTest(spGR ~ Larvae_pCO2, data=df_Exp4)


# ANOVA: The specific growth rate data meet the assumptions of parametric stats
mod_2 <- aov(spGR ~ Larvae_pCO2*Age, data=df_Exp2)
summary(mod_2)

mod_4 <- aov(spGR ~ Larvae_pCO2*Age, data=df_Exp4)
summary(mod_4)

#Mised effects model
#library(lme4)
#library(Matrix


        
Length_mixed = lmer(spGR ~Larvae_pCO2*Age + (1 | Rep), data = df_Exp4)
```


```{r}
df_spGR <- na.omit(df)
Exp_summ <- summarySE(df_spGR, measurevar="spGR", groupvars=c("Age", "Larvae_pCO2", "Run"))


Exp_MeanSE <- ggplot(Exp_summ, aes(x=Age, y=spGR, color=Larvae_pCO2))+ 
                      geom_point()+ 
                       geom_errorbar(aes(ymin=spGR-se, ymax=spGR+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      
                      labs(title = "Left: Exp #2 (F1)                 and         Right: Exp #4 (F2)", 
                           x="Age (days)", 
                           y="Specific Growth Rate (μm/day)") + 
  scale_x_continuous(limits=c(0, 18)) +
  scale_y_continuous(limits=c(0, 0.12))+
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            # axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")+
  facet_wrap(~Run)

print(Exp_MeanSE)
```


```{r F1  Expeirment #2}
Exp2_summ <- summarySE(df_Exp2, measurevar="Length", groupvars=c("Age", "Larvae_pCO2"))

Exp2_MeanSE <- ggplot(Exp2_summ, aes(x=Age, y=Length, color=Larvae_pCO2))+ geom_line()+
                      geom_point()+ 
                       geom_errorbar(aes(ymin=Length-se, ymax=Length+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      
                      labs(title = "Experiment #2 (F1)", 
                           x="Age (days)", 
                           y="Shell Length (μm)") + 
  scale_x_continuous(breaks = seq(0, 10, by = 2)) +
  scale_y_continuous(limits=c(60,200))+
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            # axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")

print(Exp2_MeanSE)
```


```{r F1  Expeirment #2}
Exp2_summ <- summarySE(df_Exp2, measurevar="spGR", groupvars=c("Age", "Larvae_pCO2"))

Exp2_summ_d <- Exp2_summ %>% drop_na()

Exp2_MeanSE <- ggplot(Exp2_summ_d, aes(x=Age, y=spGR, color=Larvae_pCO2))+ 
                      geom_point()+ 
                       geom_errorbar(aes(ymin=spGR-se, ymax=spGR+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
    geom_jitter()+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      
                      labs(title = "Experiment #2 (F1)", 
                           x="Age (days)", 
                           y="Specific Growth Rate (μm / day)") + scale_y_continuous(breaks=seq(0,.2,0.2))+ 
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            # axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")

print(Exp2_MeanSE)
```


```{r F2 Expeirment #4}
Exp4_summ <- summarySE(df_Exp4, measurevar="Length", groupvars=c("Age", "Larvae_pCO2"))

Exp4_MeanSE <- ggplot(Exp4_summ, aes(x=Age, y=Length, color=Larvae_pCO2))+ geom_line()+
                      geom_point()+ 
                       geom_errorbar(aes(ymin=Length-se, ymax=Length+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      
                      labs(title = "Experiment #4 (F2)", 
                           x="Age (days)", 
                           y="Shell Length (μm)") +   
 # scale_x_continuous(breaks = seq(0, 20, by = 2)) +
  scale_y_continuous(limits=c(60,200))+
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")
  

print(Exp4_MeanSE)
```

```{r}
################## Checking Assumptions #############
#Perform a Shapiro-Wilk test for normality ----- If p > 0.05 then data can be considered normal

library(rstatix)
shapiro.test(df_Exp4$Length) # not normal p = 0.00685

###################Equal Variances#################
#Levene's test -----Levene's Test for Homogeneity of Variance (center = median)
library(car)

leveneTest(Length ~ Larvae_pCO2, data=df_Exp4)

#Mised effects model
library(lme4)
Length_mixed = lmer(Length ~Larvae_pCO2*Age + (1 | Rep), data = df_Exp4)
summary(Length_mixed)
```


```{r F2 Expeirment #4}
Exp4_summ <- summarySE(df_Exp4, measurevar="spGR", groupvars=c("Age", "Larvae_pCO2"))

Exp4_MeanSE <- ggplot(Exp4_summ, aes(x=Age, y=spGR, color=Larvae_pCO2))+ 
                      geom_point()+ 
                       geom_errorbar(aes(ymin=spGR-se, ymax=spGR+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      
                      labs(title = "Experiment #4 (F2)", 
                           x="Age (days)", 
                           y="SPecofic Growth Rate (μm/day)")+ 
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")
  

print(Exp4_MeanSE)
```



```{r F3 Expeirment #6}
Exp6_summ <- summarySE(df_Exp6, measurevar="Length", groupvars=c("Age", "Larvae_pCO2", "Parent_pCO2"))

Exp6_MeanSE <- ggplot(Exp6_summ, aes(x=Age, y=Length, color=Larvae_pCO2))+ geom_line()+
                      geom_point()+ 
                       geom_errorbar(aes(ymin=Length-se, ymax=Length+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
  
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      
                      labs(title = "Experiment #6 (F3)", 
                           x="Age (days)", 
                           y="Shell Length (μm)") + #scale_y_continuous(breaks=seq(0,100,20))+ 
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none") + 
                      facet_wrap(~Parent_pCO2)

print(Exp6_MeanSE)
```


```{r arrange plots}

library(ggplot2)
library(gridExtra)
library(grid)

pdf("Output/Shell_length/Shell_length.pdf", width=8, height=6)
grid.newpage()
pushViewport(viewport(layout = grid.layout(2.5, 4))) # 3 rows, 5 columns
print(Exp2_MeanSE, vp = vplayout(1, 1:2))  
print(Exp4_MeanSE, vp = vplayout(1, 3:4))
print(Exp6_MeanSE, vp = vplayout(2:2.5, 1:3))
dev.off()
```


# statistics

## (i) for loop one way effects within each timepoint 

## (ii) repeated measures anova including time


## Experiment 2 F1 - i one ways many times, yippie..
```{r Experiment 2 F1}
df_Exp2

# (1) First, run anova within date for all records (for looped!)
ANOVA_Age_DPF      <- as.data.frame(unique(df_Exp2$Age)) # call a list to loop in 

AOVdf_total       <- data.frame() # start dataframe, this will be the master output
DF_loop           <- data.frame(matrix(nrow = 1, ncol = 12)) # create dataframe to save during for loop
colnames(DF_loop) <- c('Age_DPF', 'Metric', 'ShapiroWilk', 'ResidNorm', 
                       'Levenes', 'HomogVar', 'model', 'DF.num' , 'DF.denom', 
                       'F_val','P_val', 'SigDif') # names for comuns in the for loop

for (i in 1:nrow(ANOVA_Age_DPF)) {
  
  Age_loop     <- as.character(ANOVA_Age_DPF[i,])
  data_loop     <- df_Exp2 %>% # call the dataset above - after 30 dpf and for only the two treatments 
    dplyr::filter(Age %in% Age_loop) %>% 
    dplyr::select(Age, Larvae_pCO2, Length)

    # high cholorphyll model

    DF_loop$Age_DPF     <- Age_loop
    DF_loop$Metric      <- colnames(data_loop[3])
    
    # run both modles
    AOVmod              <- aov(lm(as.numeric(data_loop$Length) ~ as.factor(data_loop$Larvae_pCO2)))
    KWmod               <- kruskal.test(data_loop$Length  ~ as.factor(data_loop$Larvae_pCO2))
    
    # normality tests for the anova model - asign 
    DF_loop$ShapiroWilk <- shapiro.test(resid(AOVmod))[[2]]
    DF_loop$ResidNorm   <- if( shapiro.test(resid(AOVmod))[[2]] > 0.05) {
      'YES'} else {'NO'}
    DF_loop$Levenes     <- leveneTest(AOVmod)[[3]][[1]]
    DF_loop$HomogVar    <- if( leveneTest(AOVmod)[[3]][[1]] > 0.05) {
      'YES'} else {'NO'}
    
    if(shapiro.test(resid(AOVmod))[[2]] > 0.05 & leveneTest(AOVmod)[[3]][[1]] > 0.05) {
        DF_loop$model       <- 'one-way AOV; x ~ treatment'
        DF_loop$DF.num      <- summary(AOVmod)[[1]][["Df"]][1]
        DF_loop$DF.denom    <- summary(AOVmod)[[1]][["Df"]][2]
        DF_loop$F_val       <- summary(AOVmod)[[1]][["F value"]][1]
        DF_loop$P_val       <- summary(AOVmod)[[1]][["Pr(>F)"]][1]
        DF_loop$SigDif      <- if( (summary(AOVmod)[[1]][["Pr(>F)"]][1]) > 0.05) {
          'NO'} else {'YES'}

      } else {
        DF_loop$model       <- 'kruskal-wallis; x ~ treatment'
        DF_loop$DF.num      <- (KWmod)[[2]][["df"]][1]
        DF_loop$DF.denom    <- NA
        DF_loop$F_val       <- NA
        DF_loop$P_val       <- (KWmod)[[3]]
        DF_loop$SigDif      <- if( ((KWmod)[[3]]) > 0.05) {
          'NO'} else {'YES'}
      }
    
    # asign loop and cumulative output table
    df          <- data.frame(DF_loop) # name dataframe for this single row
    AOVdf_total <- rbind(AOVdf_total,DF_loop) #bind to a cumulative list dataframe
    print(AOVdf_total) # print to monitor progress
}
View(AOVdf_total) # view all the anova tests within data 


```



## Experiment 4 F2 - i one ways many times, yippie..
```{r Experiment 4 F2}
df_Exp4

# (1) First, run anova within date for all records (for looped!)
ANOVA_Age_DPF      <- as.data.frame(unique(df_Exp4$Age)) # call a list to loop in 

AOVdf_total       <- data.frame() # start dataframe, this will be the master output
DF_loop           <- data.frame(matrix(nrow = 1, ncol = 12)) # create dataframe to save during for loop
colnames(DF_loop) <- c('Age_DPF', 'Metric', 'ShapiroWilk', 'ResidNorm', 
                       'Levenes', 'HomogVar', 'model', 'DF.num' , 'DF.denom', 
                       'F_val','P_val', 'SigDif') # names for comuns in the for loop

for (i in 1:nrow(ANOVA_Age_DPF)) {
  
  Age_loop     <- as.character(ANOVA_Age_DPF[i,])
  data_loop     <- df_Exp4 %>% # call the dataset above - after 30 dpf and for only the two treatments 
    dplyr::filter(Age %in% Age_loop) %>% 
    dplyr::select(Age, Larvae_pCO2, Length)

    # high cholorphyll model

    DF_loop$Age_DPF     <- Age_loop
    DF_loop$Metric      <- colnames(data_loop[3])
    
    # run both modles
    AOVmod              <- aov(lm(as.numeric(data_loop$Length) ~ as.factor(data_loop$Larvae_pCO2)))
    KWmod               <- kruskal.test(data_loop$Length  ~ as.factor(data_loop$Larvae_pCO2))
    
    # normality tests for the anova model - asign 
    DF_loop$ShapiroWilk <- shapiro.test(resid(AOVmod))[[2]]
    DF_loop$ResidNorm   <- if( shapiro.test(resid(AOVmod))[[2]] > 0.05) {
      'YES'} else {'NO'}
    DF_loop$Levenes     <- leveneTest(AOVmod)[[3]][[1]]
    DF_loop$HomogVar    <- if( leveneTest(AOVmod)[[3]][[1]] > 0.05) {
      'YES'} else {'NO'}
    
    if(shapiro.test(resid(AOVmod))[[2]] > 0.05 & leveneTest(AOVmod)[[3]][[1]] > 0.05) {
        DF_loop$model       <- 'one-way AOV; x ~ treatment'
        DF_loop$DF.num      <- summary(AOVmod)[[1]][["Df"]][1]
        DF_loop$DF.denom    <- summary(AOVmod)[[1]][["Df"]][2]
        DF_loop$F_val       <- summary(AOVmod)[[1]][["F value"]][1]
        DF_loop$P_val       <- summary(AOVmod)[[1]][["Pr(>F)"]][1]
        DF_loop$SigDif      <- if( (summary(AOVmod)[[1]][["Pr(>F)"]][1]) > 0.05) {
          'NO'} else {'YES'}

      } else {
        DF_loop$model       <- 'kruskal-wallis; x ~ treatment'
        DF_loop$DF.num      <- (KWmod)[[2]][["df"]][1]
        DF_loop$DF.denom    <- NA
        DF_loop$F_val       <- NA
        DF_loop$P_val       <- (KWmod)[[3]]
        DF_loop$SigDif      <- if( ((KWmod)[[3]]) > 0.05) {
          'NO'} else {'YES'}
      }
    
    # asign loop and cumulative output table
    df          <- data.frame(DF_loop) # name dataframe for this single row
    AOVdf_total <- rbind(AOVdf_total,DF_loop) #bind to a cumulative list dataframe
    print(AOVdf_total) # print to monitor progress
}
View(AOVdf_total) # view all the anova tests within data 
```



## Experiment 6 F3 - i one ways many times, yippie..
```{r Experiment 6 F3}
df_Exp6

library(rcompanion)
library(FSA)
# (1) First, run anova within date for all records (for looped!)
ANOVA_Age_DPF      <- as.data.frame(unique(df_Exp6$Age)) # call a list to loop in 

AOVdf_total       <- data.frame() # start dataframe, this will be the master output
DF_loop           <- data.frame(matrix(nrow = 1, ncol = 12)) # create dataframe to save during for loop
colnames(DF_loop) <- c('Age_DPF', 'Metric', 'ShapiroWilk', 'ResidNorm', 
                       'Levenes', 'HomogVar', 'model', 'DF.num' , 'DF.denom', 
                       'F_val','P_val', 'SigDif') # names for comuns in the for loop

for (i in 1:nrow(ANOVA_Age_DPF)) {
  
  Age_loop     <- as.character(ANOVA_Age_DPF[1,])
  data_loop     <- df_Exp6 %>% # call the dataset above - after 30 dpf and for only the two treatments 
    dplyr::filter(Age %in% Age_loop) %>% 
    dplyr::select(Age, Larvae_pCO2, Parent_pCO2, Length)

    # high cholorphyll model

    DF_loop$Age_DPF     <- Age_loop
    DF_loop$Metric      <- colnames(data_loop[3])
    
    # run both modles
    AOVmod              <- aov(lm(as.numeric(data_loop$Length) ~ 
                                    as.factor(data_loop$Larvae_pCO2) * as.factor(data_loop$Parent_pCO2)))
    SRHmod              <- scheirerRayHare(as.numeric(data_loop$Length)  ~ 
                                          as.factor(data_loop$Larvae_pCO2) + as.factor(data_loop$Parent_pCO2))
    ?scheirerRayHare
    # normality tests for the anova model - asign 
    DF_loop$ShapiroWilk <- shapiro.test(resid(AOVmod))[[2]]
    DF_loop$ResidNorm   <- if( shapiro.test(resid(AOVmod))[[2]] > 0.05) {
      'YES'} else {'NO'}
    DF_loop$Levenes     <- leveneTest(AOVmod)[[3]][[1]]
    DF_loop$HomogVar    <- if( leveneTest(AOVmod)[[3]][[1]] > 0.05) {
      'YES'} else {'NO'}
    
    if(shapiro.test(resid(AOVmod))[[2]] > 0.05 & leveneTest(AOVmod)[[3]][[1]] > 0.05) {
        DF_loop$model       <- 'two-way AOV; x ~ treatment'
        DF_loop$DF.num      <- summary(AOVmod)[[1]][["Df"]][1]
        DF_loop$DF.denom    <- summary(AOVmod)[[1]][["Df"]][2]
        DF_loop$F_val       <- summary(AOVmod)[[1]][["F value"]][1]
        DF_loop$P_val       <- summary(AOVmod)[[1]][["Pr(>F)"]][1]
        DF_loop$SigDif      <- if( (summary(AOVmod)[[1]][["Pr(>F)"]][1]) > 0.05) {
          'NO'} else {'YES'}

      } else {
        DF_loop$model       <- 'kruskal-wallis; x ~ treatment'
        DF_loop$DF.num      <- (KWmod)[[2]][["df"]][1]
        DF_loop$DF.denom    <- NA
        DF_loop$F_val       <- NA
        DF_loop$P_val       <- (KWmod)[[3]]
        DF_loop$SigDif      <- if( ((KWmod)[[3]]) > 0.05) {
          'NO'} else {'YES'}
      }
    
    # asign loop and cumulative output table
    df          <- data.frame(DF_loop) # name dataframe for this single row
    AOVdf_total <- rbind(AOVdf_total,DF_loop) #bind to a cumulative list dataframe
    print(AOVdf_total) # print to monitor progress
}
View(AOVdf_total) # view all the anova tests within data 



```


