---
title: "Larval Growth"
author: "Katie McFarland"
date: "9/29/2023"
output: html_document
---

# 20231217 - started manuscript Github repostiory 'EAD-ASEB-Airradians_Larvae_crossgen_OA'
# SGurr edits
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
 knitr::opts_knit$set(root.dir = "C:/Users/katherine.mcfarland/Documents/GitHub/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis") # Katie's
#knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis") # Sam's
```


```{r, echo=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(rcompanion)
library(FSA)
library(car)
library(forcats)
library(kableExtra) # nice Rmd tables
library(emmeans)
library(ggpubr)
library(Rmisc)
```


# Load Larval Growth Data

```{r}
df<-read.csv("Data/larval_growth.csv", header = T) %>% 
  dplyr::mutate(Lengh = as.numeric(Length)) %>% 
  
  # rename the variables for consistancy using 'case_when'
  dplyr::mutate(Parental_OA = case_when(Parental_OA ==  "LOW" ~ "Low pCO2", 
                                        Parental_OA ==  "MODERATE" ~ "Moderate pCO2", 
                                        Parental_OA ==  "HIGH" ~ "High pCO2",
                                        Parental_OA ==  "WILD" ~ "None")) %>% 
  dplyr::mutate(treatment = case_when(treatment ==  "Low OA" ~ "Low pCO2", 
                                      treatment ==  "Moderate OA" ~ "Moderate pCO2", 
                                      treatment ==  "High OA" ~ "High pCO2")) %>% 
  
  # change the levels for order when plotting - call as factors too!
  dplyr::mutate(Parental_OA = as.factor(forcats::fct_relevel(Parental_OA, 
                                                             c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  dplyr::mutate(treatment = as.factor(forcats::fct_relevel(treatment, 
                                                           c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  
  # rename them
  dplyr::rename(Parent_pCO2 = Parental_OA) %>% 
  dplyr::rename(Larvae_pCO2  = treatment) # done 

df <- df%>%
  filter(Stage=="larvae")

head(df)
str(df)
```

# About data 
### Summary: 

  - (1) F1 larvae under three OA levels, *no grow out*

  - (2) F1 larvae under three OA levels, *grown out to adult stage*

  - (3) F2 larvae full cross 3 x 3 parent x offspring OA, *no grow out*

  - (4) F2 larvae under two OA levels, *grown out to adult stage*

  - (5) F2 larvae under 3 x 3 parent x offspring OA, *no grow out*

  - (6) F3 larvae under 3 x 3 parent x offsring OA, *no grow out past juvenile takedown of the experiment*

```{r about data}

# number of runs 
unique(df$Generation) # 1 , 2, 3
unique(df$Run) # 2, 4 and 6

Rep_Summary_all <- df %>% dplyr::select(c(Generation, Run, Age, Parent_pCO2, Larvae_pCO2)) %>% 
                       dplyr::group_by_all() %>% 
                       dplyr::summarise(n=n())

# print the runs
# summary: 

Rep_Summary_Exp1 <- Rep_Summary_all %>%  dplyr::filter(Run %in% 2) 
Rep_Summary_Exp1 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_Exp4 <- Rep_Summary_all %>%  dplyr::filter(Run %in% 4) 
Rep_Summary_Exp4 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_Exp6 <- Rep_Summary_all %>%  dplyr::filter(Run %in% 6) 
Rep_Summary_Exp6 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")


```


### make data frames for each generation
```{r}
df_Exp2 <- df%>%
  filter(Run=="2")

df_Exp4 <- df%>%
  filter(Run=="4")

df_Exp6 <- df%>%
  filter(Run=="6")
```

#Plots for all facented by run
```{r}
Exp_summ <- summarySE(df, measurevar="Length", groupvars=c("Age", "Larvae_pCO2", "Run"))

Exp_MeanSE <- ggplot(Exp_summ, aes(x=Age, y=Length, color=Larvae_pCO2))+ geom_line()+
                      geom_point()+ 
                       geom_errorbar(aes(ymin=Length-se, ymax=Length+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      
                      labs(title = "Left: Exp #2 (F1)                 and         Right: Exp #4 (F2)", 
                           x="Age (days)", 
                           y="Shell Length (μm)") + 
  scale_x_continuous(breaks = seq(0, 20, by = 2)) +
  scale_y_continuous(limits=c(60,210))+
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            # axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")+
  facet_wrap(~Run)

print(Exp_MeanSE)
```
```{r}
################## Checking Assumptions #############
#Perform a Shapiro-Wilk test for normality ----- If p > 0.05 then data can be considered normal

library(rstatix)
shapiro.test(df_Exp2$spGR) 
shapiro.test(df_Exp4$spGR) 

###################Equal Variances#################
#Levene's test -----Levene's Test for Homogeneity of Variance (center = median)
library(car)
leveneTest(spGR ~ Larvae_pCO2, data=df_Exp2)
leveneTest(spGR ~ Larvae_pCO2, data=df_Exp4)


# ANOVA: The specific growth rate data meet the assumptions of parametric stats
mod_2 <- aov(spGR ~ Larvae_pCO2*Age, data=df_Exp2)
summary(mod_2)

mod_4 <- aov(spGR ~ Larvae_pCO2*Age, data=df_Exp4)
summary(mod_4)

#Mised effects model
#library(lme4)
#library(Matrix


        
Length_mixed = lmer(spGR ~Larvae_pCO2*Age + (1 | Rep), data = df_Exp4)
```
```{r}
df_spGR <- na.omit(df)
Exp_summ <- summarySE(df_spGR, measurevar="spGR", groupvars=c("Age", "Larvae_pCO2", "Run"))


Exp_MeanSE <- ggplot(Exp_summ, aes(x=Age, y=spGR, color=Larvae_pCO2))+ 
                      geom_point()+ 
                       geom_errorbar(aes(ymin=spGR-se, ymax=spGR+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      
                      labs(title = "Left: Exp #2 (F1)                 and         Right: Exp #4 (F2)", 
                           x="Age (days)", 
                           y="Specific Growth Rate (μm/day)") + 
  scale_x_continuous(limits=c(0, 18)) +
  scale_y_continuous(limits=c(0, 0.12))+
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            # axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")+
  facet_wrap(~Run)

print(Exp_MeanSE)
```


```{r F1  Expeirment #2}
Exp2_summ <- summarySE(df_Exp2, measurevar="Length", groupvars=c("Age", "Larvae_pCO2"))

Exp2_MeanSE <- ggplot(Exp2_summ, aes(x=Age, y=Length, color=Larvae_pCO2))+ geom_line()+
                      geom_point()+ 
                       geom_errorbar(aes(ymin=Length-se, ymax=Length+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      
                      labs(title = "Experiment #2 (F1)", 
                           x="Age (days)", 
                           y="Shell Length (μm)") + 
  scale_x_continuous(breaks = seq(0, 10, by = 2)) +
  scale_y_continuous(limits=c(60,200))+
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            # axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")

print(Exp2_MeanSE)
```


```{r F1  Expeirment #2}
Exp2_summ <- summarySE(df_Exp2, measurevar="spGR", groupvars=c("Age", "Larvae_pCO2"))

Exp2_summ_d <- Exp2_summ %>% drop_na()

Exp2_MeanSE <- ggplot(Exp2_summ_d, aes(x=Age, y=spGR, color=Larvae_pCO2))+ 
                      geom_point()+ 
                       geom_errorbar(aes(ymin=spGR-se, ymax=spGR+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
    geom_jitter()+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      
                      labs(title = "Experiment #2 (F1)", 
                           x="Age (days)", 
                           y="Specific Growth Rate (μm / day)") + scale_y_continuous(breaks=seq(0,.2,0.2))+ 
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            # axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")

print(Exp2_MeanSE)
```


```{r F2 Expeirment #4}
Exp4_summ <- summarySE(df_Exp4, measurevar="Length", groupvars=c("Age", "Larvae_pCO2"))

Exp4_MeanSE <- ggplot(Exp4_summ, aes(x=Age, y=Length, color=Larvae_pCO2))+ geom_line()+
                      geom_point()+ 
                       geom_errorbar(aes(ymin=Length-se, ymax=Length+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      
                      labs(title = "Experiment #4 (F2)", 
                           x="Age (days)", 
                           y="Shell Length (μm)") +   
 # scale_x_continuous(breaks = seq(0, 20, by = 2)) +
  scale_y_continuous(limits=c(60,200))+
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")
  

print(Exp4_MeanSE)
```

```{r}
################## Checking Assumptions #############
#Perform a Shapiro-Wilk test for normality ----- If p > 0.05 then data can be considered normal

library(rstatix)
shapiro.test(df_Exp4$Length) # not normal p = 0.00685

###################Equal Variances#################
#Levene's test -----Levene's Test for Homogeneity of Variance (center = median)
library(car)

leveneTest(Length ~ Larvae_pCO2, data=df_Exp4)

#Mised effects model
library(lme4)
Length_mixed = lmer(Length ~Larvae_pCO2*Age + (1 | Rep), data = df_Exp4)
```


```{r F2 Expeirment #4}
Exp4_summ <- summarySE(df_Exp4, measurevar="spGR", groupvars=c("Age", "Larvae_pCO2"))

Exp4_MeanSE <- ggplot(Exp4_summ, aes(x=Age, y=spGR, color=Larvae_pCO2))+ 
                      geom_point()+ 
                       geom_errorbar(aes(ymin=spGR-se, ymax=spGR+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      
                      labs(title = "Experiment #4 (F2)", 
                           x="Age (days)", 
                           y="SPecofic Growth Rate (μm/day)")+ 
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.x=element_blank(),
                            axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none")
  

print(Exp4_MeanSE)
```



```{r F3 Expeirment #6}
Exp6_summ <- summarySE(df_Exp6, measurevar="Length", groupvars=c("Age", "Larvae_pCO2", "Parent_pCO2"))

Exp6_MeanSE <- ggplot(Exp6_summ, aes(x=Age, y=Length, color=Larvae_pCO2))+ geom_line()+
                      geom_point()+ 
                       geom_errorbar(aes(ymin=Length-se, ymax=Length+se), width=.2,
                                    position=position_dodge(.1)) +
                      scale_color_manual(values=c("green4", "darkorange1", "purple"))+
  
                      theme(legend.position="right", 
                            legend.direction="vertical", 
                            legend.title = element_blank())+ theme_bw() + 
                      
                      labs(title = "Experiment #6 (F3)", 
                           x="Age (days)", 
                           y="Shell Length (μm)") + #scale_y_continuous(breaks=seq(0,100,20))+ 
                      theme_classic() +
                      theme(panel.grid.major = element_blank(), 
                            panel.grid.minor = element_blank(), 
                            axis.title.y=element_blank(),
                            panel.background = element_blank(), 
                            axis.line = element_line(colour = "black"),
                            legend.position="none") + 
                      facet_wrap(~Parent_pCO2)

print(Exp6_MeanSE)
```


```{r arrange plots}

library(ggplot2)
library(gridExtra)
library(grid)

pdf("Output/Shell_length/Shell_length.pdf", width=8, height=6)
grid.newpage()
pushViewport(viewport(layout = grid.layout(2.5, 4))) # 3 rows, 5 columns
print(Exp2_MeanSE, vp = vplayout(1, 1:2))  
print(Exp4_MeanSE, vp = vplayout(1, 3:4))
print(Exp6_MeanSE, vp = vplayout(2:2.5, 1:3))
dev.off()
```


# statistics

## (i) for loop one way effects within each timepoint 

## (ii) repeated measures anova including time


## Experiment 2 F1 - i one ways many times, yippie..
```{r Experiment 2 F1}
df_Exp2

# (1) First, run anova within date for all records (for looped!)
ANOVA_Age_DPF      <- as.data.frame(unique(df_Exp2$Age)) # call a list to loop in 

AOVdf_total       <- data.frame() # start dataframe, this will be the master output
DF_loop           <- data.frame(matrix(nrow = 1, ncol = 12)) # create dataframe to save during for loop
colnames(DF_loop) <- c('Age_DPF', 'Metric', 'ShapiroWilk', 'ResidNorm', 
                       'Levenes', 'HomogVar', 'model', 'DF.num' , 'DF.denom', 
                       'F_val','P_val', 'SigDif') # names for comuns in the for loop

for (i in 1:nrow(ANOVA_Age_DPF)) {
  
  Age_loop     <- as.character(ANOVA_Age_DPF[i,])
  data_loop     <- df_Exp2 %>% # call the dataset above - after 30 dpf and for only the two treatments 
    dplyr::filter(Age %in% Age_loop) %>% 
    dplyr::select(Age, Larvae_pCO2, Length)

    # high cholorphyll model

    DF_loop$Age_DPF     <- Age_loop
    DF_loop$Metric      <- colnames(data_loop[3])
    
    # run both modles
    AOVmod              <- aov(lm(as.numeric(data_loop$Length) ~ as.factor(data_loop$Larvae_pCO2)))
    KWmod               <- kruskal.test(data_loop$Length  ~ as.factor(data_loop$Larvae_pCO2))
    
    # normality tests for the anova model - asign 
    DF_loop$ShapiroWilk <- shapiro.test(resid(AOVmod))[[2]]
    DF_loop$ResidNorm   <- if( shapiro.test(resid(AOVmod))[[2]] > 0.05) {
      'YES'} else {'NO'}
    DF_loop$Levenes     <- leveneTest(AOVmod)[[3]][[1]]
    DF_loop$HomogVar    <- if( leveneTest(AOVmod)[[3]][[1]] > 0.05) {
      'YES'} else {'NO'}
    
    if(shapiro.test(resid(AOVmod))[[2]] > 0.05 & leveneTest(AOVmod)[[3]][[1]] > 0.05) {
        DF_loop$model       <- 'one-way AOV; x ~ treatment'
        DF_loop$DF.num      <- summary(AOVmod)[[1]][["Df"]][1]
        DF_loop$DF.denom    <- summary(AOVmod)[[1]][["Df"]][2]
        DF_loop$F_val       <- summary(AOVmod)[[1]][["F value"]][1]
        DF_loop$P_val       <- summary(AOVmod)[[1]][["Pr(>F)"]][1]
        DF_loop$SigDif      <- if( (summary(AOVmod)[[1]][["Pr(>F)"]][1]) > 0.05) {
          'NO'} else {'YES'}

      } else {
        DF_loop$model       <- 'kruskal-wallis; x ~ treatment'
        DF_loop$DF.num      <- (KWmod)[[2]][["df"]][1]
        DF_loop$DF.denom    <- NA
        DF_loop$F_val       <- NA
        DF_loop$P_val       <- (KWmod)[[3]]
        DF_loop$SigDif      <- if( ((KWmod)[[3]]) > 0.05) {
          'NO'} else {'YES'}
      }
    
    # asign loop and cumulative output table
    df          <- data.frame(DF_loop) # name dataframe for this single row
    AOVdf_total <- rbind(AOVdf_total,DF_loop) #bind to a cumulative list dataframe
    print(AOVdf_total) # print to monitor progress
}
View(AOVdf_total) # view all the anova tests within data 


```



## Experiment 4 F2 - i one ways many times, yippie..
```{r Experiment 4 F2}
df_Exp4

# (1) First, run anova within date for all records (for looped!)
ANOVA_Age_DPF      <- as.data.frame(unique(df_Exp4$Age)) # call a list to loop in 

AOVdf_total       <- data.frame() # start dataframe, this will be the master output
DF_loop           <- data.frame(matrix(nrow = 1, ncol = 12)) # create dataframe to save during for loop
colnames(DF_loop) <- c('Age_DPF', 'Metric', 'ShapiroWilk', 'ResidNorm', 
                       'Levenes', 'HomogVar', 'model', 'DF.num' , 'DF.denom', 
                       'F_val','P_val', 'SigDif') # names for comuns in the for loop

for (i in 1:nrow(ANOVA_Age_DPF)) {
  
  Age_loop     <- as.character(ANOVA_Age_DPF[i,])
  data_loop     <- df_Exp4 %>% # call the dataset above - after 30 dpf and for only the two treatments 
    dplyr::filter(Age %in% Age_loop) %>% 
    dplyr::select(Age, Larvae_pCO2, Length)

    # high cholorphyll model

    DF_loop$Age_DPF     <- Age_loop
    DF_loop$Metric      <- colnames(data_loop[3])
    
    # run both modles
    AOVmod              <- aov(lm(as.numeric(data_loop$Length) ~ as.factor(data_loop$Larvae_pCO2)))
    KWmod               <- kruskal.test(data_loop$Length  ~ as.factor(data_loop$Larvae_pCO2))
    
    # normality tests for the anova model - asign 
    DF_loop$ShapiroWilk <- shapiro.test(resid(AOVmod))[[2]]
    DF_loop$ResidNorm   <- if( shapiro.test(resid(AOVmod))[[2]] > 0.05) {
      'YES'} else {'NO'}
    DF_loop$Levenes     <- leveneTest(AOVmod)[[3]][[1]]
    DF_loop$HomogVar    <- if( leveneTest(AOVmod)[[3]][[1]] > 0.05) {
      'YES'} else {'NO'}
    
    if(shapiro.test(resid(AOVmod))[[2]] > 0.05 & leveneTest(AOVmod)[[3]][[1]] > 0.05) {
        DF_loop$model       <- 'one-way AOV; x ~ treatment'
        DF_loop$DF.num      <- summary(AOVmod)[[1]][["Df"]][1]
        DF_loop$DF.denom    <- summary(AOVmod)[[1]][["Df"]][2]
        DF_loop$F_val       <- summary(AOVmod)[[1]][["F value"]][1]
        DF_loop$P_val       <- summary(AOVmod)[[1]][["Pr(>F)"]][1]
        DF_loop$SigDif      <- if( (summary(AOVmod)[[1]][["Pr(>F)"]][1]) > 0.05) {
          'NO'} else {'YES'}

      } else {
        DF_loop$model       <- 'kruskal-wallis; x ~ treatment'
        DF_loop$DF.num      <- (KWmod)[[2]][["df"]][1]
        DF_loop$DF.denom    <- NA
        DF_loop$F_val       <- NA
        DF_loop$P_val       <- (KWmod)[[3]]
        DF_loop$SigDif      <- if( ((KWmod)[[3]]) > 0.05) {
          'NO'} else {'YES'}
      }
    
    # asign loop and cumulative output table
    df          <- data.frame(DF_loop) # name dataframe for this single row
    AOVdf_total <- rbind(AOVdf_total,DF_loop) #bind to a cumulative list dataframe
    print(AOVdf_total) # print to monitor progress
}
View(AOVdf_total) # view all the anova tests within data 
```



## Experiment 6 F3 - i one ways many times, yippie..
```{r Experiment 6 F3}
df_Exp6

library(rcompanion)
library(FSA)
# (1) First, run anova within date for all records (for looped!)
ANOVA_Age_DPF      <- as.data.frame(unique(df_Exp6$Age)) # call a list to loop in 

AOVdf_total       <- data.frame() # start dataframe, this will be the master output
DF_loop           <- data.frame(matrix(nrow = 1, ncol = 12)) # create dataframe to save during for loop
colnames(DF_loop) <- c('Age_DPF', 'Metric', 'ShapiroWilk', 'ResidNorm', 
                       'Levenes', 'HomogVar', 'model', 'DF.num' , 'DF.denom', 
                       'F_val','P_val', 'SigDif') # names for comuns in the for loop

for (i in 1:nrow(ANOVA_Age_DPF)) {
  
  Age_loop     <- as.character(ANOVA_Age_DPF[1,])
  data_loop     <- df_Exp6 %>% # call the dataset above - after 30 dpf and for only the two treatments 
    dplyr::filter(Age %in% Age_loop) %>% 
    dplyr::select(Age, Larvae_pCO2, Parent_pCO2, Length)

    # high cholorphyll model

    DF_loop$Age_DPF     <- Age_loop
    DF_loop$Metric      <- colnames(data_loop[3])
    
    # run both modles
    AOVmod              <- aov(lm(as.numeric(data_loop$Length) ~ 
                                    as.factor(data_loop$Larvae_pCO2) * as.factor(data_loop$Parent_pCO2)))
    SRHmod              <- scheirerRayHare(as.numeric(data_loop$Length)  ~ 
                                          as.factor(data_loop$Larvae_pCO2) + as.factor(data_loop$Parent_pCO2))
    ?scheirerRayHare
    # normality tests for the anova model - asign 
    DF_loop$ShapiroWilk <- shapiro.test(resid(AOVmod))[[2]]
    DF_loop$ResidNorm   <- if( shapiro.test(resid(AOVmod))[[2]] > 0.05) {
      'YES'} else {'NO'}
    DF_loop$Levenes     <- leveneTest(AOVmod)[[3]][[1]]
    DF_loop$HomogVar    <- if( leveneTest(AOVmod)[[3]][[1]] > 0.05) {
      'YES'} else {'NO'}
    
    if(shapiro.test(resid(AOVmod))[[2]] > 0.05 & leveneTest(AOVmod)[[3]][[1]] > 0.05) {
        DF_loop$model       <- 'two-way AOV; x ~ treatment'
        DF_loop$DF.num      <- summary(AOVmod)[[1]][["Df"]][1]
        DF_loop$DF.denom    <- summary(AOVmod)[[1]][["Df"]][2]
        DF_loop$F_val       <- summary(AOVmod)[[1]][["F value"]][1]
        DF_loop$P_val       <- summary(AOVmod)[[1]][["Pr(>F)"]][1]
        DF_loop$SigDif      <- if( (summary(AOVmod)[[1]][["Pr(>F)"]][1]) > 0.05) {
          'NO'} else {'YES'}

      } else {
        DF_loop$model       <- 'kruskal-wallis; x ~ treatment'
        DF_loop$DF.num      <- (KWmod)[[2]][["df"]][1]
        DF_loop$DF.denom    <- NA
        DF_loop$F_val       <- NA
        DF_loop$P_val       <- (KWmod)[[3]]
        DF_loop$SigDif      <- if( ((KWmod)[[3]]) > 0.05) {
          'NO'} else {'YES'}
      }
    
    # asign loop and cumulative output table
    df          <- data.frame(DF_loop) # name dataframe for this single row
    AOVdf_total <- rbind(AOVdf_total,DF_loop) #bind to a cumulative list dataframe
    print(AOVdf_total) # print to monitor progress
}
View(AOVdf_total) # view all the anova tests within data 



```


