---
title: "Seawater Chemistry"
author: "Samuel Gurr"
date: "8/7/2024"
output: html_document
---

### SET UP
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 

#knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis") # Sam's
knitr::opts_knit$set(root.dir = 
                       "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/") # Sam's work


```





```{r load data}
# install.packages('gcookbook')
# LOAD PACKAGES :::::::::::::::::::::::::::::::::::::::::::::::::::::::
library(dplyr)
library(ggplot2)
library(reshape2)
library(knitr)
library(kableExtra)
library(car)
library(gcookbook)
library(ggpubr)



# LOAD DATA :::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# 8.16/2025 shannon assembled a means and SD dataset for the DIC+TA to calculate 
CarbChem_Meseck_menas <- read.csv(file="Data/Seawater_chemistry/raw/Seawater_Chemsitry_means_Meseck_20240816.csv", 
                                  header=TRUE) # load the chem data
# This is all the data, we can back calculate to get to what Shannon assembled or just use the one above..
# if we come toa conclusion that TA+pH or pH+DIC is desired than we can do that
CarbChem_all <- read.csv(file="Data/Seawater_chemistry/raw/Seawater_Chemsitry_raw.csv", 
                         header=TRUE) # load the chem data

```


## Objective 

* For the F1 F2 Larvae paper we want two 8-panel figures (one per generation) containing the following: 

  - Salinity 
  - temperature bucket
  - total alkalinity
  - total dissolved carbon 
  - pH measured
  - pH out calculated
  - pCO2 out
  - omega aragonite


## Master files 

CarbChem_all - contains the chemistry for all larval periods, note the generation and the run #s 
the manuscript on F1 and F2 larvae used experiment #s 2 and 4

```{r master files}

# Filter for experiment 2 and 4
CarbChem_Exp2_Exp4 <- CarbChem_all %>% dplyr::filter(Run == c(2,4))

# call columns with TADIC in the column name - these are the CO2sys calculated values 
CarbChem_Exp2_Exp4_select <- cbind( 
                                  (CarbChem_Exp2_Exp4[,c(1,3,7:11)]), # columns, core measured data  and factors
                                  (CarbChem_Exp2_Exp4[ ,grepl( ("TADIC"), names(CarbChem_Exp2_Exp4) ) ]) # TADIC colmns
                                )

# long format by group 
CarbChem_Exp2_Exp4_long <- CarbChem_Exp2_Exp4_select %>% 
                          melt(id.vars=c('Date', 'Generation', 'Larvae_OA')) %>% 
                          na.omit() %>% 
                          dplyr::mutate(Date = 
                                          as.POSIXct(Date, format = "%m/%d/%Y"),
                                        Month = months(Date),
                                        Year  = lubridate::year(Date)) 

CarbChem_Exp2_Exp4_long$value <- as.numeric(CarbChem_Exp2_Exp4_long$value) # make numeric

# MEANS PER SAMPLING DAY (reducing replicates within treatment to one mean) 
# aquire means by group for mean SE plotting by Date and treamtent
CarbChem_Exp2_Exp4_means.by.day <- CarbChem_Exp2_Exp4_long %>% # calc means and standard error
                           dplyr::group_by(Date, Generation, Larvae_OA, variable) %>% 
                           dplyr::summarise(mean = mean(value),
                                            sd   = sd(value),
                                            se   = sd/(sqrt(n())), 
                                            n = n()) %>% 
                            dplyr::rename(pCO2_treatment = Larvae_OA) %>% 
                            dplyr::mutate(pCO2_treatment = 
                                            case_when(
                                                      pCO2_treatment %in% 'Mid' ~ 'Moderate',
                                                      TRUE ~ as.factor(pCO2_treatment)
                                            ),
                                          
                            )
CarbChem_Exp2_Exp4_means.by.day$pCO2_treatment <-
  forcats::fct_relevel(CarbChem_Exp2_Exp4_means.by.day$pCO2_treatment,c("Low","Moderate","High"))                           
# Note there are quite a few dates that have only a signle same for that day...


```

```{r SmoothRibbon function with gam}
# how do we get smooth geom ribbon figures
# in summary run gam and use a fine scale for x axis 
# https://stackoverflow.com/questions/61029929/smoothing-geom-ribbon
# A principled way to achieve what you want is to fit a GAM model to your data using the gam() function in mgcv and then apply the predict() function to that model over a finer grid of values for your predictor variable. The grid can cover the span defined by the range of observed values for your predictor variable.
# load R packages
library(ggplot2)
library(mgcv)


SmoothRibbon_twoOA <- function(datafilename, col_name) {

data <- as.data.frame(
  # carbchem.MEANS.by.day %>% 
                datafilename %>% 
                na.omit() %>% 
                # dplyr::filter(!Type %in% 'Larvae') %>carbchem.MASTER_MEANS.MONTLY% 
                dplyr::filter(pCO2_treatment %in% c("Low", "Moderate")) %>% 
                dplyr::filter(variable %in% (
                                               # 'Temperature_bucket_C',
                                               # 'Salinity',
                                               # 'DO_mg_l',
                                               # 'pH',
                                               # 'pH_out'
                                               col_name
                                               # 'TCO2_mmol_kgSW'
                                               # 'TA_mmol_kgSW',
                                               # 'pCO2_out_matm'
                                               #'War_out_DIC_pH'
                                               # 'Wca_out'
                                               ))
                                          ) %>% 
                 mutate(Date_asnum = dense_rank(Date))


# calc the num days because the Date_asnumin ascending order dot not 
# represent the elapsed time properly as an x axis
data$NumDays  <- as.numeric(
                            difftime(as.Date(data$Date),
                            as.Date(data$Date[1]),
                            units="days")
                            )

data.low      <- data %>% dplyr::filter(pCO2_treatment %in% 'Low')
data.moderate <- data %>% dplyr::filter(pCO2_treatment %in% 'Moderate')
data.all.ordered  <- rbind(data.low, data.moderate)


# fit GAM model
m.low      <- gam(mean ~ s(NumDays, k = length(table(data.low$mean)) ), data = data.low) 
m.moderate <- gam(mean ~ s(NumDays, k = length(table(data.moderate$mean)) ), data = data.moderate) 

# define finer grid of predictor values
xnew.low      <- seq(min(data.low$NumDays), 
                     max(data.low$NumDays), by = 0.1) 
xnew.moderate <- seq(min(data.moderate$NumDays), 
                     max(data.moderate$NumDays), by = 0.1) 

# apply predict() function to the fitted GAM model
# using the finer grid of x values
p.low      <- predict(m.low, newdata = data.frame(NumDays = xnew.low), se = TRUE) 
p.moderate <- predict(m.moderate, newdata = data.frame(NumDays = xnew.moderate), se = TRUE) 

# plot the estimated mean values of y (fit) at given x values
# over the finer grid of x values;
# superimpose approximate 95% confidence band for the true 
# mean values of y at given x values in the finer grid
g.low <- data.frame(NumDays = xnew.low, 
            fit = p.low$fit,
            lwr = p.low$fit - 1.96*p.low$se.fit, # 1.96 * SE = 95% CI
            upr = p.low$fit + 1.96*p.low$se.fit  # 1.96 * SE = 95% CI
            ) %>% 
         dplyr::mutate(pCO2_treatment = "Low")

g.moderate <- data.frame(NumDays = xnew.moderate, 
            fit = p.moderate$fit,
            lwr = p.moderate$fit - 1.96*p.moderate$se.fit, 
            upr = p.moderate$fit + 1.96*p.moderate$se.fit) %>% 
         dplyr::mutate(pCO2_treatment = "Moderate")

# r bind the data sets
# use the first date of data and add the elapsed numberic time, not that 1 = seconds, we 
# convert to days by mutliplying by 86400 (seconds in a day)
g.master <- rbind(g.low, g.moderate) %>% 
                  dplyr::mutate(Date = data$Date[1] + NumDays*86400)

theme_set(theme_classic())

p <- ggplot(data = g.master, aes(Date, fit)) + 
  geom_ribbon(aes(ymin = lwr, ymax = upr, fill=factor(pCO2_treatment)), alpha = 0.2) + 
  geom_line(aes(y = lwr,
                colour=factor(pCO2_treatment)),
            linetype = "solid") +
  geom_line(aes(y = upr,
                colour=factor(pCO2_treatment)),
            linetype = "solid") +
  geom_point(data = data.all.ordered, 
             aes(Date, mean,
                 colour=factor(pCO2_treatment)), 
             shape = 3, size =1) +
  ylab(col_name) +
  theme(legend.position="none") +
  scale_fill_manual(values=c("forestgreen", "orange"), name="fill") +
  scale_colour_manual(values=c("forestgreen", "orange"), name="fill") 

print(p)
}




SmoothRibbon_threeOA <- function(datafilename, col_name) {

data <- as.data.frame(
  #carbchem.MEANS.by.day.F1 %>% 
                datafilename %>% 
                na.omit() %>% 
                # dplyr::filter(!Type %in% 'Larvae') %>carbchem.MASTER_MEANS.MONTLY% 
                # dplyr::filter(pCO2_treatment %in% c("Low", "Moderate", "High")) %>% 
                dplyr::filter(variable %in% (
                                               #'TADIC_WAr_out_in_situ'
                                               # 'Salinity',
                                               # 'DO_mg_l',
                                               # 'pH',
                                               # 'pH_out'
                                               col_name
                                               # 'TCO2_mmol_kgSW'
                                               # 'TA_mmol_kgSW',
                                               # 'pCO2_out_matm'
                                               #'War_out_DIC_pH'
                                               # 'Wca_out'
                                               ))
                                          ) %>% 
                 mutate(Date_asnum = dense_rank(Date))


# calc the num days because the Date_asnumin ascending order dot not 
# represent the elapsed time properly as an x axis
data$NumDays  <- as.numeric(
                            difftime(as.Date(data$Date),
                            as.Date(data$Date[1]),
                            units="days")
                            )

data.low      <- data %>% dplyr::filter(pCO2_treatment %in% 'Low')
data.moderate <- data %>% dplyr::filter(pCO2_treatment %in% 'Moderate')
data.high     <- data %>% dplyr::filter(pCO2_treatment %in% 'High')
data.all.ordered  <- rbind(data.low, data.moderate, data.high)


# fit GAM model
m.low      <- gam(mean ~ s(NumDays, k = length(table(data.low$mean)) ), data = data.low) 
m.moderate <- gam(mean ~ s(NumDays, k = length(table(data.moderate$mean)) ), data = data.moderate) 
m.high     <- gam(mean ~ s(NumDays, k = length(table(data.high$mean)) ), data = data.high) 

# define finer grid of predictor values
xnew.low      <- seq(min(data.low$NumDays), 
                     max(data.low$NumDays), by = 0.1) 
xnew.moderate <- seq(min(data.moderate$NumDays), 
                     max(data.moderate$NumDays), by = 0.1) 
xnew.high     <- seq(min(data.high$NumDays), 
                     max(data.high$NumDays), by = 0.1) 
# apply predict() function to the fitted GAM model
# using the finer grid of x values
p.low      <- predict(m.low, newdata = data.frame(NumDays = xnew.low), se = TRUE) 
p.moderate <- predict(m.moderate, newdata = data.frame(NumDays = xnew.moderate), se = TRUE) 
p.high     <- predict(m.high, newdata = data.frame(NumDays = xnew.high), se = TRUE) 

# plot the estimated mean values of y (fit) at given x values
# over the finer grid of x values;
# superimpose approximate 95% confidence band for the true 
# mean values of y at given x values in the finer grid
g.low <- data.frame(NumDays = xnew.low, 
            fit = p.low$fit,
            lwr = p.low$fit - 1.96*p.low$se.fit, # 1.96 * SE = 95% CI
            upr = p.low$fit + 1.96*p.low$se.fit  # 1.96 * SE = 95% CI
            ) %>% 
         dplyr::mutate(pCO2_treatment = "Low")

g.moderate <- data.frame(NumDays = xnew.moderate, 
            fit = p.moderate$fit,
            lwr = p.moderate$fit - 1.96*p.moderate$se.fit, 
            upr = p.moderate$fit + 1.96*p.moderate$se.fit) %>% 
         dplyr::mutate(pCO2_treatment = "Moderate")

g.high <- data.frame(NumDays = xnew.high, 
            fit = p.high$fit,
            lwr = p.high$fit - 1.96*p.high$se.fit, 
            upr = p.high$fit + 1.96*p.high$se.fit) %>% 
         dplyr::mutate(pCO2_treatment = "High")
# r bind the data sets
# use the first date of data and add the elapsed numberic time, not that 1 = seconds, we 
# convert to days by mutliplying by 86400 (seconds in a day)
g.master <- rbind(g.low, g.moderate, g.high) %>% 
                  dplyr::mutate(Date = data$Date[1] + NumDays*86400)

theme_set(theme_classic())

levels(g.master$pCO2_treatment) = c("Low", "Moderate", "High")

p <- ggplot(data = g.master, aes(Date, fit)) + 
  geom_ribbon(aes(ymin = lwr, ymax = upr, 
                  fill=(forcats::fct_relevel(pCO2_treatment,c("Low", "Moderate", "High")))), 
                  alpha = 0.2) + 
  geom_line(aes(y = lwr,
                colour=(pCO2_treatment)),
            linetype = "solid") +
  geom_line(aes(y = upr,
                colour=(pCO2_treatment)),
            linetype = "solid") +
  geom_point(data = data.all.ordered, 
             aes(Date, mean,
                 colour=(pCO2_treatment)), 
             shape = 3, size =1) +
  ylab(col_name) +
  theme(legend.position="none") +
  scale_fill_manual(values=c("forestgreen", "orange","purple"), name="fill") +
  scale_colour_manual(values=c("forestgreen","orange", "purple"), name="fill") 

print(p)
}
```

* USe Smooth Ribbon and GAM to make some narly plots!
```{r F1 three OA treamtents - Nice plots!}
library(ggpubr)
carbchem.MEANS.by.day.F1 <- CarbChem_Exp2_Exp4_means.by.day %>% filter(Generation %in% 'F1') 
# pdf("C:/Users/samjg/Documents/Github_repositories/Airradians_multigen_OA/RAnalysis/Output/SeawaterChemistry/F1/plots/F1_Chem_95CI.pdf", width=8, height=11)
pdf("Output/1_Survival_Growth/Seawater_chemistry/F1_Chem_95CI.pdf", width=8, height=11)
    ggarrange(
          SmoothRibbon_threeOA(carbchem.MEANS.by.day.F1, "Temperature_bucket_C"),
          SmoothRibbon_threeOA(carbchem.MEANS.by.day.F1, "Salinity"),
          SmoothRibbon_threeOA(carbchem.MEANS.by.day.F1, "TA_mmol_kgSW"),
          SmoothRibbon_threeOA(carbchem.MEANS.by.day.F1, "TCO2_mmol_kgSW"),
          # SmoothRibbon_threeOA(carbchem.MEANS.by.day.F1, ""),
          SmoothRibbon_threeOA(carbchem.MEANS.by.day.F1, "TADIC_pH_out_in_situ"),          
          SmoothRibbon_threeOA(carbchem.MEANS.by.day.F1, "TADIC_pCO2_out_.matm._in_situ"),          
          SmoothRibbon_threeOA(carbchem.MEANS.by.day.F1, "TADIC_WAr_out_in_situ"),          
          nrow = 4,
          ncol = 2
    )
graphics.off()


```


```{r F2 two OA treamtents - Nice plots!}

carbchem.MEANS.by.day.F2 <- CarbChem_Exp2_Exp4_means.by.day %>% filter(Generation %in% 'F2')

# pdf("C:/Users/samjg/Documents/Github_repositories/Airradians_multigen_OA/RAnalysis/Output/SeawaterChemistry/F2/plots/F2_Chem_95CI.pdf", width=8, height=11)
pdf("Output/1_Survival_Growth/Seawater_chemistry/F2_Chem_95CI.pdf", width=8, height=11)
    ggarrange(
          SmoothRibbon_twoOA(carbchem.MEANS.by.day.F1, "Temperature_bucket_C"),
          SmoothRibbon_twoOA(carbchem.MEANS.by.day.F1, "Salinity"),
          SmoothRibbon_twoOA(carbchem.MEANS.by.day.F1, "TA_mmol_kgSW"),
          SmoothRibbon_twoOA(carbchem.MEANS.by.day.F1, "TCO2_mmol_kgSW"),
          # SmoothRibbon_twoOA(carbchem.MEANS.by.day.F1, ""),
          SmoothRibbon_twoOA(carbchem.MEANS.by.day.F1, "TADIC_pH_out_in_situ"),          
          SmoothRibbon_twoOA(carbchem.MEANS.by.day.F1, "TADIC_pCO2_out_.matm._in_situ"),          
          SmoothRibbon_twoOA(carbchem.MEANS.by.day.F1, "TADIC_WAr_out_in_situ"), 
          nrow = 4,
          ncol = 2
    )
graphics.off()

```



# TABLE (all data)

```{r use dplyr to build chemistry table, include=FALSE}
carbchem.MASTER_MEANS <- carbchem.MASTER_long %>% # calc means and standard error
   na.omit() %>% 
   dplyr::group_by(Date, Type, pCO2_treatment, variable) %>% 
   dplyr::summarise(mean = mean(value),
                    sd   = sd(value),
                    se   = sd/(sqrt(n())) )

chem_meanDate <- carbchem.MASTER_MEANS %>% 
     na.omit() %>% 
   dplyr::group_by(pCO2_treatment, variable) %>% 
   dplyr::summarise(mean2 = mean(mean),
                    sd   = sd(mean),
                    se   = sd/(sqrt(n())) ,
                    n = n())

carbchem.MASTER_wideMEANS <- dcast(chem_meanDate, pCO2_treatment ~ variable, value.var="mean2")
carbchem.MASTER_wideStErr <- dcast(chem_meanDate, pCO2_treatment ~ variable, value.var="se")
carbchem.MASTER_wideN     <- dcast(chem_meanDate, pCO2_treatment ~ variable, value.var="n")

# final table

colnames(carbchem.MASTER_wideMEANS)
FINAL_TABLE                <- data.frame(matrix(nrow = nrow(carbchem.MASTER_wideMEANS), ncol = 1))

FINAL_TABLE$pCO2_Treatment <- carbchem.MASTER_wideMEANS$pCO2_treatment

FINAL_TABLE$Salinity       <-  paste( (paste( (signif(carbchem.MASTER_wideMEANS$Salinity, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$Salinity, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$Salinity,
                                     ")",
                                     sep = "")

FINAL_TABLE$pCO2           <-  paste( (paste( (signif(carbchem.MASTER_wideMEANS$xCO2_out._dry_at._1_atm_.ppm_.DIC_pH, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$xCO2_out._dry_at._1_atm_.ppm_.DIC_pH, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$xCO2_out._dry_at._1_atm_.ppm_.DIC_pH,
                                     ")",
                                     sep = "")

FINAL_TABLE$Temperature    <- paste( (paste( (signif(carbchem.MASTER_wideMEANS$Temperature_bucket_C, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$Temperature_bucket_C, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$Temperature_bucket_C,
                                     ")",
                                     sep = "")

FINAL_TABLE$pH             <- paste( (paste( (signif(carbchem.MASTER_wideMEANS$pH_out, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$pH_out, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$pH_out,
                                     ")",
                                     sep = "")

FINAL_TABLE$HCO3           <-  paste( (paste( (signif(carbchem.MASTER_wideMEANS$HCO3_out_mmol_kgSW, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$HCO3_out_mmol_kgSW, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$HCO3_out_mmol_kgSW,
                                     ")",
                                     sep = "")

FINAL_TABLE$CO3           <-  paste( (paste( (signif(carbchem.MASTER_wideMEANS$CO3_out_mmol_kgSW, digits=3)),
                                             "±",
                                     (signif(carbchem.MASTER_wideStErr$CO3_out_mmol_kgSW, digits=3)), sep="  ")),
                                     " (",
                                     carbchem.MASTER_wideN$CO3_out_mmol_kgSW,
                                     ")",
                                     sep = "")

FINAL_TABLE$TA             <- paste( (paste( (signif(carbchem.MASTER_wideMEANS$TA_mmol_kgSW, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$TA_mmol_kgSW, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$TA_mmol_kgSW,
                                     ")",
                                     sep = "")

FINAL_TABLE$Aragonite.Sat  <-  paste( (paste( (signif(carbchem.MASTER_wideMEANS$War_out_DIC_pH, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$War_out_DIC_pH, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$War_out_DIC_pH,
                                     ")",
                                     sep = "")

FINAL_TABLE                <- FINAL_TABLE[,-1] # view table
```


## knitR to print the table 
```{r lable output, echo=TRUE}
FINAL_TABLE %>%
  kbl(caption = "Table 1. Seawater chemistry") %>%
  kable_classic(full_width = T, html_font = "Cambria")
```


#### save data
```{r save tables, echo=TRUE}
# save output table
# write.table(chem_meanDate,"C:/Users/samjg/Documents/Github_repositories/Cvirginica_multistressor/RAnalysis/Output/Water_Chem/Chem.Table_sep.csv",sep=",", row.names=FALSE)  # write table to 
# write.table(FINAL_TABLE,"C:/Users/samjg/Documents/Github_repositories/Cvirginica_multistressor/RAnalysis/Output/Water_Chem/Chem.Table.csv",sep=",", row.names=FALSE)  # write table to 

```










# TABLE (all data)

```{r use dplyr to build chemistry table, include=FALSE}

# change to long format to properly facet
carbchem.MASTER_long <- carbchem.MASTER %>% 
  dplyr::select(!c(Treatment,Replicate)) %>% 
  melt(id.vars=c('Date', 'Type', 'pCO2_treatment'))

carbchem.MASTER_long$value <- as.numeric(carbchem.MASTER_long$value) # make numeric

carbchem.MASTER_MEANS <- carbchem.MASTER_long %>% # calc means and standard error
   na.omit() %>% 
   dplyr::group_by(Date, Type, pCO2_treatment, variable) %>% 
   dplyr::summarise(mean = mean(value),
                    sd   = sd(value),
                    se   = sd/(sqrt(n())) )

chem_meanDate <- carbchem.MASTER_MEANS %>% 
     na.omit() %>% 
   dplyr::group_by(pCO2_treatment, variable) %>% 
   dplyr::summarise(mean2 = mean(mean),
                    sd   = sd(mean),
                    se   = sd/(sqrt(n())) ,
                    n = n())

carbchem.MASTER_wideMEANS <- dcast(chem_meanDate, pCO2_treatment ~ variable, value.var="mean2")
carbchem.MASTER_wideStErr <- dcast(chem_meanDate, pCO2_treatment ~ variable, value.var="se")
carbchem.MASTER_wideN     <- dcast(chem_meanDate, pCO2_treatment ~ variable, value.var="n")

# final table

colnames(carbchem.MASTER_wideMEANS)
FINAL_TABLE                <- data.frame(matrix(nrow = nrow(carbchem.MASTER_wideMEANS), ncol = 1))

FINAL_TABLE$pCO2_Treatment <- carbchem.MASTER_wideMEANS$pCO2_treatment

FINAL_TABLE$Salinity       <-  paste( (paste( (signif(carbchem.MASTER_wideMEANS$Salinity, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$Salinity, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$Salinity,
                                     ")",
                                     sep = "")

FINAL_TABLE$pCO2           <-  paste( (paste( (signif(carbchem.MASTER_wideMEANS$xCO2_out._dry_at._1_atm_.ppm_.DIC_pH, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$xCO2_out._dry_at._1_atm_.ppm_.DIC_pH, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$xCO2_out._dry_at._1_atm_.ppm_.DIC_pH,
                                     ")",
                                     sep = "")

FINAL_TABLE$Temperature    <- paste( (paste( (signif(carbchem.MASTER_wideMEANS$Temperature_bucket_C, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$Temperature_bucket_C, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$Temperature_bucket_C,
                                     ")",
                                     sep = "")

FINAL_TABLE$pH             <- paste( (paste( (signif(carbchem.MASTER_wideMEANS$pH_out, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$pH_out, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$pH_out,
                                     ")",
                                     sep = "")

FINAL_TABLE$HCO3           <-  paste( (paste( (signif(carbchem.MASTER_wideMEANS$HCO3_out_mmol_kgSW, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$HCO3_out_mmol_kgSW, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$HCO3_out_mmol_kgSW,
                                     ")",
                                     sep = "")

FINAL_TABLE$CO3           <-  paste( (paste( (signif(carbchem.MASTER_wideMEANS$CO3_out_mmol_kgSW, digits=3)),
                                             "±",
                                     (signif(carbchem.MASTER_wideStErr$CO3_out_mmol_kgSW, digits=3)), sep="  ")),
                                     " (",
                                     carbchem.MASTER_wideN$CO3_out_mmol_kgSW,
                                     ")",
                                     sep = "")

FINAL_TABLE$TA             <- paste( (paste( (signif(carbchem.MASTER_wideMEANS$TA_mmol_kgSW, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$TA_mmol_kgSW, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$TA_mmol_kgSW,
                                     ")",
                                     sep = "")

FINAL_TABLE$Aragonite.Sat  <-  paste( (paste( (signif(carbchem.MASTER_wideMEANS$War_out_DIC_pH, digits=3)),
                                     (signif(carbchem.MASTER_wideStErr$War_out_DIC_pH, digits=3)), sep=" ± ")),
                                     " (",
                                     carbchem.MASTER_wideN$War_out_DIC_pH,
                                     ")",
                                     sep = "")

FINAL_TABLE                <- FINAL_TABLE[,-1] # view table
```


## knitR to print the table 
```{r lable output, echo=TRUE}
FINAL_TABLE %>%
  kbl(caption = "Table 1. Seawater chemistry") %>%
  kable_classic(full_width = T, html_font = "Cambria")
```


#### save data (food x OA manuscript)
```{r save tables, echo=TRUE}
# save output table
# write.table(chem_meanDate,"C:/Users/samjg/Documents/Github_repositories/Cvirginica_multistressor/RAnalysis/Output/Water_Chem/Chem.Table_sep.csv",sep=",", row.names=FALSE)  # write table to 
# write.table(FINAL_TABLE,"C:/Users/samjg/Documents/Github_repositories/Cvirginica_multistressor/RAnalysis/Output/Water_Chem/Chem.Table.csv",sep=",", row.names=FALSE)  # write table to 

```
