---
title: "Larval Growth and Survival"
author: "Katie McFarland"
date: "9/29/2023"
output: html_document
---

# 20231217 - started manuscript Github repostiory 'EAD-ASEB-Airradians_Larvae_crossgen_OA'
# SGurr edits
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
# knitr::opts_knit$set(root.dir = "C:/Users/katherine.mcfarland/Documents/GitHub/EAD-ASEB-Airradians_multigen_OA/larvae") # Katie's
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis") # Sam's
```


```{r, echo=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(rcompanion)
library(FSA)
library(car)
library(forcats)
library(kableExtra) # nice Rmd tables
library(emmeans)
library(ggpubr)
library(survival)
library(Rmisc)
```


#Larval Survival
```{r}
df<-read.csv("Data/larval_survival.csv", header = T) %>% 
  dplyr::mutate(Percent_Survival = as.numeric(Percent_Survival)) %>% 
  dplyr::select(-treatment) %>%  # redundant column with Exposure_OA
  
  # rename the variables for consistency using 'case_when'
  dplyr::mutate(Parental_OA = case_when(Parental_OA ==  "LOW" ~ "Low pCO2", 
                                        Parental_OA ==  "MODERATE" ~ "Moderate pCO2", 
                                        Parental_OA ==  "HIGH" ~ "High pCO2",
                                        Parental_OA ==  "WILD" ~ "None")) %>% 
  dplyr::mutate(Exposure_OA = case_when(Exposure_OA ==  "LOW" ~ "Low pCO2", 
                                        Exposure_OA ==  "MODERATE" ~ "Moderate pCO2", 
                                        Exposure_OA ==  "HIGH" ~ "High pCO2")) %>% 
  
  # change the levels for order when plotting - call as factors too!
  dplyr::mutate(Parental_OA = as.factor(forcats::fct_relevel(Parental_OA, c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  dplyr::mutate(Exposure_OA = as.factor(forcats::fct_relevel(Exposure_OA, c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  
  # rename them
  dplyr::rename(Parent_pCO2 = Parental_OA) %>% 
  dplyr::rename(Larvae_pCO2  = Exposure_OA) # done 
  
  
df$Percent_Survival_trans <- asin(sqrt(df$Percent_Survival))

head(df)
str(df)
```



```{r about data}

# number of runs 
unique(df$Generation) # 1, 2, and 3

Rep_Summary_all <- df %>% dplyr::select(c(Generation, Age, Parent_pCO2, Larvae_pCO2)) %>% 
                       dplyr::group_by_all() %>% 
                       dplyr::summarise(n=n())

# print the runs
# summary: 

Rep_Summary_G1 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 1) # 
Rep_Summary_G1 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_G2 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 2) #
Rep_Summary_G2 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_G3 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 3) # 
Rep_Summary_G3 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

```


### make data frames for each generation
```{r}
df_F1 <- df%>%
  filter(Generation=="1")

df_F2 <- df%>%
  filter(Generation=="2")

df_F3 <- df%>%
  filter(Generation=="3")
```

# F1 Larval Survival
```{r}

stL <- summarySE(df_F1, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2"))
stL 

ggplot(stL, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
  geom_point()+ 
   geom_errorbar(aes(ymin=Percent_Survival-se, ymax=Percent_Survival+se), width=.2,
                position=position_dodge(.1)) +
  scale_color_manual(values=c("green4", "darkorange1", "purple"))+
  theme(legend.position="right", legend.direction="vertical", legend.title = element_blank())+ theme_bw() + 
  labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))
 
```

# F2 Larval Survival
```{r}

stL <- summarySE(df_F2, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2"))
stL 

ggplot(stL, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2))+ geom_line()+
  geom_point()+ 
   geom_errorbar(aes(ymin=Percent_Survival-se, ymax=Percent_Survival+se), width=.2,
                position=position_dodge(.1)) +
  scale_color_manual(values=c("green4", "darkorange1", "purple"))+
  theme(legend.position="right", legend.direction="vertical", legend.title = element_blank())+ theme_bw() + 
  labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))
 
```
# F3 Larval Survival
```{r}

stL <- summarySE(df_F3, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2", "Parent_pCO2"))
stL 

Gen3_PlotMeanSE_All <- ggplot(stL, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2))+ 
                        geom_line()+
                        geom_point()+ 
                        geom_errorbar(aes(ymin=Percent_Survival-se, ymax=Percent_Survival+se), width=.2,
                                      position=position_dodge(.1)) +
                        scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                        theme(legend.position="right", legend.direction="vertical", 
                              legend.title = element_blank())+ theme_bw() + 
                        labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              panel.background = element_blank(), 
                              axis.line = element_line(colour = "black"))+
                        facet_wrap("Parent_pCO2")



# Summarise Percent Deformities for plotting 
Gen3_mean_survival <- df_F3 %>% 
                          dplyr::select(Parent_pCO2,Larvae_pCO2,Age, Percent_Survival) %>% 
                          na.omit() %>% 
                          dplyr::group_by(Age, Parent_pCO2,Larvae_pCO2) %>% 
                          dplyr::summarise(mean_Perc_surv = mean(Percent_Survival), 
                                           n           = n(),
                                           sd_Perc_surv   = sd(Percent_Survival),
                                           se_Perc_surv   = sd_Perc_surv/(sqrt(n)))

# Barplots 
unique(Gen3_mean_survival$Age)
Gen3_barplots_Age2 <- Gen3_mean_survival %>% dplyr::filter(Age ==2) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 2 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)
Gen3_barplots_Age5 <- Gen3_mean_survival %>% dplyr::filter(Age ==5) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 5 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

Gen3_barplots_Age7 <- Gen3_mean_survival %>% dplyr::filter(Age ==7) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 7 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

Gen3_barplots_Age12 <- Gen3_mean_survival %>% dplyr::filter(Age ==12) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 12 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)
Gen3_barplots_Age16 <- Gen3_mean_survival %>% dplyr::filter(Age ==16) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 16 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

# print the plot

pdf("Output/Survival/PercSurvival_Gen3_All.pdf", width=10, height=5)
print(Gen3_PlotMeanSE_All)
dev.off()


pdf("Output/Survival/PercSurvival_Gen3_2DPF.pdf", width=5, height=5)
print(Gen3_barplots_Age2)
dev.off()

pdf("Output/Survival/PercSurvival_Gen3_5DPF.pdf", width=5, height=5)
print(Gen3_barplots_Age5)
dev.off()

pdf("Output/Survival/PercSurvival_Gen3_7DPF.pdf", width=5, height=5)
print(Gen3_barplots_Age7)
dev.off()

pdf("Output/Survival/PercSurvival_Gen3_12DPF.pdf", width=5, height=5)
print(Gen3_barplots_Age12)
dev.off()

pdf("Output/Survival/PercSurvival_Gen3_16DPF.pdf", width=5, height=5)
print(Gen3_barplots_Age16)
dev.off()

```


# Survival package  stats

create surv object
```{r}
#make a surv object to pass into coxme
df_F3 <- df_F3 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(Count_dead = 
                          (1/as.numeric(proportion))*as.numeric(Count_alive)) # proportion is that of alive

# Call the cumulative dataframe that we will write to in the for loop below
binary_df_F3_all     <- data.frame() # start dataframe 
binary_df_F3_high    <- data.frame() # start dataframe 
binary_df_F3_mod     <- data.frame() # start dataframe 
binary_df_F3_low     <- data.frame() # start dataframe 

for (i in 1:nrow(df_F3)) {
   count_alive <- round(df_F3[i,]$Count_alive/1000) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F3[i,]$Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_F3[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_F3[i,]$Larvae_pCO2,
                            Generation  = df_F3[i,]$Generation,
                            Tank        = df_F3[i,]$Tank,  
                            Age         = df_F3[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF       <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F3_all <- rbind(binary_df_F3_all,loopDF) #bind to a cumulative list dataframe
  # print(binary_df_F3) # print to monitor progress

     
  if (loopDF$Parent_pCO2[1] == 'High pCO2') {
      binary_df_F3_high <- rbind(binary_df_F3_high,loopDF) #bind to a cumulative list dataframe
  } else if (loopDF$Parent_pCO2[1] == 'Moderate pCO2') {
      binary_df_F3_mod <- rbind(binary_df_F3_mod,loopDF) #bind to a cumulative list dataframe
  } else {
      binary_df_F3_low <- rbind(binary_df_F3_low,loopDF) #bind to a cumulative list dataframe
    }
      
  
}

binary_df_F3$All_pCO2 <- as.factor(paste0(binary_df_F3$Parent_pCO2, ' x ', binary_df_F3$Larvae_pCO2))


```


```{r}

surv_F3_all  <- Surv(time = as.numeric(binary_df_F3_all$Age), 
                     event = as.numeric(binary_df_F3_all$Count_dead), type = "right")

surv_F3_high <- Surv(time = as.numeric(binary_df_F3_high$Age), 
                     event = as.numeric(binary_df_F3_high$Count_dead), type = "right")

surv_F3_mod  <- Surv(time = as.numeric(binary_df_F3_mod$Age), 
                     event = as.numeric(binary_df_F3_mod$Count_dead), type = "right")

surv_F3_low  <- Surv(time = as.numeric(binary_df_F3_low$Age), 
                     event = as.numeric(binary_df_F3_low$Count_dead), type = "right")


library(survminer)



ggsurvplot(survfit(surv_F3_high ~ Larvae_pCO2, binary_df_F3_high), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange", "purple")),
           #                 c(0.7,0.5, 0.3, 0.1)), 
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")

ggsurvplot(survfit(surv_F3_mod ~ Larvae_pCO2, binary_df_F3_mod), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange", "purple")),
           #                 c(0.7,0.5, 0.3, 0.1)), 
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")

ggsurvplot(survfit(surv_F3_low ~ Larvae_pCO2, binary_df_F3_low), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange", "purple")),
           #                 c(0.7,0.5, 0.3, 0.1)), 
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")
```
run cox propotional hazard test
```{r}
survFit1 <- coxph(surv_F3 ~ factor(All_pCO2), binary_df_F3 )
survFit1 <- coxph(surv_F3 ~ Parent_pCO2 *Larvae_pCO2 , binary_df_F3 )
summary(survFit1)
View(binary_df_F3)

survFit_all <- coxph(surv_F3_all ~ Parent_pCO2 *Larvae_pCO2 , binary_df_F3_all)
summary(survFit_all)

survFit_high <- coxph(surv_F3_high ~ Larvae_pCO2 , binary_df_F3_high)
summary(survFit_high)

survFit_low <- coxph(surv_F3_low ~ Larvae_pCO2 , binary_df_F3_low)
summary(survFit_low)

survFit_mod <- coxph(surv_F3_mod ~ Larvae_pCO2 , binary_df_F3_mod)
summary(survFit_mod)
```

include tank as a random effect
```{r}

survFitME <- coxme(surv_F3 ~ Parent_pCO2 *Larvae_pCO2 + (1|Tank), binary_df_F3)
summary(survFitME)
#The standard deviation of the tank random effect is very small (0.02) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (15.4), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 
```

calculate coefficient of determination (r^2)
```{r}
library(yardstick)
rsq(survFit1, sigD=2)

```



```{r}


#pairwise comparisons

library(multcomp)

glht(survFitME, linfct = mcp(pH = "Dunnett"),alternative = "greater")

```




# Stats
Lenght data do not meet the assumptions of noramlity even with an asin(sqrt()) transformation
```{r}
leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F1)
shapiro.test(df_F1$Percent_Survival_trans)

leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F2)
shapiro.test(df_F2$Percent_Survival_trans)

leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F3)
shapiro.test(df_F3$Percent_Survival_trans)
```

#Non parametric Stats
### F1 Survival
```{r}
scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Age,
                data = df_F1)

dunnTest(Percent_Survival ~ Larvae_pCO2,
                data = df_F1,
              method="bh")

```

### F2 Survival
```{r}
scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Age,
                data = df_F2)

res.kruskal <- kruskal.test(Percent_Survival ~ Larvae_pCO2, data = df_F2)
res.kruskal

```

### F3 Survival
```{r}


scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Parent_pCO2,
                data = df_F3)

dunnTest(Percent_Survival ~ Larvae_pCO2,
                data = df_F3,
              method="bh")

dunnTest(Percent_Survival ~ Parent_pCO2,
                data = df_F3,
              method="bh")
```