---
title: "Larval Growth and Survival"
author: "Katie McFarland"
date: "9/29/2023"
output:
  pdf_document: default
  html_document: default
---

# 20231217 - started manuscript Github repostiory 'EAD-ASEB-Airradians_Larvae_crossgen_OA'
# SGurr edits
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
# knitr::opts_knit$set(root.dir = "C:/Users/katherine.mcfarland/Documents/GitHub/EAD-ASEB-Airradians_multigen_OA/larvae") # Katie's
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis") # Sam's
```


```{r, echo=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(rcompanion)
library(FSA)
library(car)
library(forcats)
library(kableExtra) # nice Rmd tables
library(emmeans)
library(ggpubr)
library(survival)
library(Rmisc)
library(coxme)
library(survminer)
library(ggsurvfit) # survfit2
library(gtsummary) # tbl_survfit
```

# Load data 
```{r, echo=FALSE}
df <- read.csv("Data/larval_survival.csv", header = T) %>% 
  dplyr::mutate(Percent_Survival = as.numeric(Percent_Survival)) %>% 
  # dplyr::select(-treatment) %>%  # redundant column with Exposure_OA
  
  # rename the variables for consistency using 'case_when'
  dplyr::mutate(Parental_OA = case_when(Parental_OA ==  "LOW" ~ "Low pCO2", 
                                        Parental_OA ==  "MODERATE" ~ "Moderate pCO2", 
                                        Parental_OA ==  "HIGH" ~ "High pCO2",
                                        Parental_OA ==  "WILD" ~ "None")) %>% 
  dplyr::mutate(Exposure_OA = case_when(Exposure_OA ==  "LOW" ~ "Low pCO2", 
                                        Exposure_OA ==  "MODERATE" ~ "Moderate pCO2", 
                                        Exposure_OA ==  "HIGH" ~ "High pCO2")) %>% 
  
  # change the levels for order when plotting - call as factors too!
  dplyr::mutate(Parental_OA = as.factor(forcats::fct_relevel(Parental_OA, 
                                                             c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  dplyr::mutate(Exposure_OA = as.factor(forcats::fct_relevel(Exposure_OA, 
                                                             c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  
  # rename them
  dplyr::rename(Parent_pCO2 = Parental_OA) %>% 
  dplyr::rename(Larvae_pCO2  = Exposure_OA) # done 
  
  
df$Percent_Survival_trans <- asin(sqrt(df$Percent_Survival))

# head(df)
# str(df)
```

## About data 

* output tables for each generation where 'n' is the number of tanks!

```{r about data}

# number of runs 
unique(df$Generation) # 1, 2, and 3

Rep_Summary_all <- df %>% dplyr::select(c(Generation, Age, Parent_pCO2, Larvae_pCO2)) %>% 
                       dplyr::group_by_all() %>% 
                       dplyr::summarise(n=n())

# print the runs
# summary: 

Rep_Summary_G1 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 1) # 
Rep_Summary_G1 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_G2 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 2) #
Rep_Summary_G2 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_G3 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 3) # 
Rep_Summary_G3 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

```


### Build data frames for each generation
```{r}
df_F1 <- df%>%
  filter(Generation=="1")

df_F2 <- df%>%
  filter(Generation=="2")

df_F3 <- df%>%
  filter(Generation=="3")
```

# F1 & F2 Larval Survival

* note that metamorphosis was observed at 10 dpf in F1s and 14 dpf in F2s, therfore 
larval survival comapred across gnerations should only show < 10 dpf survival 

## F1s: Plot all raw count live data  
```{r F1 summary and plot ALL}

stL_F1 <- summarySE(df_F1, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2"))
stL_F1 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")


F1_PlotMeanSE <-  ggplot(stL_F1, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                          geom_point()+ 
                           geom_errorbar(aes(ymin=Percent_Survival-se, 
                                             ymax=Percent_Survival+se), width=.2,
                                        position=position_dodge(.1)) +
                          scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                          theme(legend.position="right", legend.direction="vertical", 
                                legend.title = element_blank())+ theme_bw() + 
                          labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                          ggtitle("F1 larvae + post-metamorphosis survival (Mean +- SE)") + 
                          theme(panel.grid.major = element_blank(), 
                                panel.grid.minor = element_blank(), 
                                panel.background = element_blank(), 
                                axis.line = element_line(colour = "black"))
 
F1_PlotMeanSE

pdf("Output/Survival/F1_survival_all_meanSE.pdf", width=10, height=5)
print(F1_PlotMeanSE)
dev.off()

```

## F1s: Plot  raw count live data before 10 dpf 
```{r F1 summary and plot JUST LARVAE}

stL_F1 <- summarySE(df_F1, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2"))
stL_F1 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")


F1_larvae_PlotMeanSE <-  stL_F1 %>%  dplyr::filter(Age <= 10) %>% 
                                ggplot(aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                                        geom_point()+ 
                                         geom_errorbar(aes(ymin=Percent_Survival-se, 
                                                           ymax=Percent_Survival+se), width=.2,
                                                      position=position_dodge(.1)) +
                                        scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                                        theme(legend.position="right", legend.direction="vertical", 
                                              legend.title = element_blank())+ theme_bw() + 
                                        labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                                        ggtitle("F1 larvae survival (Mean +- SE)") + 
                                        theme(panel.grid.major = element_blank(), 
                                              panel.grid.minor = element_blank(), 
                                              panel.background = element_blank(), 
                                              axis.line = element_line(colour = "black"))
 
F1_PlotMeanSE

pdf("Output/Survival/F1_survival_larvae_meanSE.pdf", width=10, height=5)
print(F1_larvae_PlotMeanSE)
dev.off()

```

## F2s: Plot all raw count live data  
```{r F2 summary and plot ALL}

stL_F2 <- summarySE(df_F2, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2"))
stL_F2 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

F2_PlotMeanSE <- ggplot(stL_F2, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2))+ geom_line()+
                            geom_point()+ 
                             geom_errorbar(aes(ymin=Percent_Survival-se, ymax=Percent_Survival+se), width=.2,
                                          position=position_dodge(.1)) +
                            scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                            theme(legend.position="right", 
                                  legend.direction="vertical", 
                                  legend.title = element_blank())+ theme_bw() + 
                            labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                            ggtitle("F2 larvae survival (Mean +- SE)") + 
                            theme(panel.grid.major = element_blank(), 
                                  panel.grid.minor = element_blank(), 
                                  panel.background = element_blank(), 
                                  axis.line = element_line(colour = "black"))
 
F2_PlotMeanSE

pdf("Output/Survival/F2_survival__all_meanSE.pdf", width=10, height=5)
print(F2_PlotMeanSE)
dev.off()

```

## F2s: Plot  raw count live data before 10 dpf 
```{r F2 summary and plot JUST LARVAE}

stL_F2 <- summarySE(df_F2, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2"))
stL_F2 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")


F2_larvae_PlotMeanSE <-  stL_F2 %>%  dplyr::filter(Age <= 10) %>% 
                                ggplot(aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                                        geom_point()+ 
                                         geom_errorbar(aes(ymin=Percent_Survival-se, 
                                                           ymax=Percent_Survival+se), width=.2,
                                                      position=position_dodge(.1)) +
                                        scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                                        theme(legend.position="right", legend.direction="vertical", 
                                              legend.title = element_blank())+ theme_bw() + 
                                        labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                                        ggtitle("F2 larvae survival (Mean +- SE)") + 
                                        theme(panel.grid.major = element_blank(), 
                                              panel.grid.minor = element_blank(), 
                                              panel.background = element_blank(), 
                                              axis.line = element_line(colour = "black"))
 
F2_larvae_PlotMeanSE

pdf("Output/Survival/F2_survival_larvae_meanSE.pdf", width=10, height=5)
print(F2_larvae_PlotMeanSE)
dev.off()

```

# Kaplan-Meier plots

* We will use the {ggsurvfit} package to generate Kaplan-Meier plots. This package aims to ease plotting of time-to-event endpoints using the power of the {ggplot2} package. See http://www.danieldsjoberg.com/ggsurvfit/index.html for details.

## Build 'surv' object for statistical analysis 

* useful links [https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html#Comparing_survival_times_between_groups]

* What the hell dpes a surv object mean?? 
  
  - a 'surv' object reads ALL events as that of a dead (1) or alive (0) individual
  
  - but but but... we have volumetric count data of live and dead! Does this not throw a wrench into this type of analysis? 
  
    - Nope! Below we use logic to create a GIGANTIC table formatted correctly for this purpose!

```{r F1 & F2 edit}
# create new columns for live and dead based on the count and proportion (of alive) tobackcalculate dead - these are integers! 
df_F1 <- df_F1 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(Count_dead = case_when(proportion == 0 ~ 0,
                                               proportion >0 ~
                                                 (1/as.numeric(proportion))*
                                                 as.numeric(Count_alive))) # proportion is that of alive

df_F2 <- df_F2 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(Count_dead = case_when(proportion == 0 ~ 0,
                                               proportion >0 ~ (1/as.numeric(proportion))*
                                                 as.numeric(Count_alive))) # proportion is that of alive

```

# Run two for loops 

 * (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting
 
 * (2) Mean counts alive and dead by tank - better representated for plots!

### F1s binary loop 1
```{r F1 build binary live and dead column - 1}
# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting

# Call the cumulative dataframe that we will write to in the for loop below
binary_df_F1_all     <- data.frame() # start dataframe  - all
# run it 
for (i in 1:nrow(df_F1)) {
   count_alive <- round(df_F1[i,]$Count_alive/1000) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F1[i,]$Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   #count_alive <- round(df_F1[i,]$Count_alive) # raw counts
   #count_dead  <- round(df_F1[i,]$Count_dead)  # raw counts
   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_F1[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_F1[i,]$Larvae_pCO2,
                            Generation  = df_F1[i,]$Generation,
                            Tank        = df_F1[i,]$Tank,  
                            Age         = df_F1[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF           <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F1_all <- rbind(binary_df_F1_all,loopDF) #bind to a cumulative list dataframe
}
# To address tank as a random factor it MUST be numeric, change this here
unique(binary_df_F1_all$Tank) # "A" "B" "C" "D" "E" NA
binary_df_F1_all <- binary_df_F1_all %>% 
                    dplyr::mutate(Generation = 1,
                                  Tank = case_when(Tank== 'A'~1,
                                                   Tank== 'B'~2,
                                                   Tank== 'C'~3,
                                                   Tank== 'D'~4,
                                                   Tank== 'E'~4)) %>% 
                    na.omit()
unique(binary_df_F1_all$Tank)  #1 2 3 4
```

### F1s binary loop 2
```{r F1 build binary live and dead column - 2}
# (2) Mean counts alive and dead by tank - better representated for plots!

# for loop for plotting!! 
# difference here is we take a mean for counts alive and dead within treatment, no tank included and a singluar line for plotting 
df_F1_MEANS <- df_F1 %>% 
              dplyr::select(-c(Generation, Tank, Date, Percent_Survival, proportion)) %>% 
              dplyr::group_by(Larvae_pCO2, Age) %>% 
              dplyr::summarise(mean_Count_alive = mean(Count_alive), 
                               mean_Count_dead = mean(Count_dead))

# call datafarme binary for plotting
binary_df_F1_MEANS       <- data.frame() # start dataframe 

# run it 
for (i in 1:nrow(df_F1_MEANS)) {
   count_alive <- round(df_F1_MEANS[i,]$mean_Count_alive/1000) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F1_MEANS[i,]$mean_Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   loopDF           <- data.frame(matrix(0,ncol = 3, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Larvae_pCO2','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Larvae_pCO2 = df_F1_MEANS[i,]$Larvae_pCO2,
                            Age         = df_F1_MEANS[i,]$Age
                          ) 
  loopDF[c(1:count_dead),3] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),3] = 0 # 0 for alive

  loopDF             <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F1_MEANS <- rbind(binary_df_F1_MEANS,loopDF) #bind to a cumulative list dataframe
  
}
binary_df_F1_MEANS <- binary_df_F1_MEANS %>% 
                      dplyr::mutate(Generation = 1) %>% 
                      na.omit()
```

### F2s binary loop 1
```{r F2 build binary live and dead column - 1}
# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting

# keep tank to use as a random effect in the cox model!
# Call the cumulative dataframe that we will write to in the for loop below
binary_df_F2_all     <- data.frame() # start dataframe  - all

# run it 
for (i in 1:nrow(df_F2)) {
   count_alive <- round(df_F2[i,]$Count_alive/1000) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F2[i,]$Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_F2[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_F2[i,]$Larvae_pCO2,
                            Generation  = df_F2[i,]$Generation,
                            Tank        = df_F2[i,]$Tank,  
                            Age         = df_F2[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF           <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F2_all <- rbind(binary_df_F2_all,loopDF) #bind to a cumulative list dataframe
}
# To address tank as a random factor it MUST be numeric, change this here
unique(binary_df_F2_all$Tank) # "A" "B" "C" NA 
binary_df_F2_all <- binary_df_F2_all %>% 
                    dplyr::mutate(Generation = 2,
                                  Tank = case_when(Tank== 'A'~1,
                                                   Tank== 'B'~2,
                                                   Tank== 'C'~3)) %>% 
                    na.omit()
unique(binary_df_F2_all$Tank)  #1 2 3 

```

### F2s binary loop 2
```{r F2 build binary live and dead column - 2}
# (2) Mean counts alive and dead by tank - better presentative for plots!

# for loop for plotting!! 
# difference here is we take a mean for counts alive and dead within treatment, no tank included and a singluar line for plotting 
df_F2_MEANS <- df_F2 %>% 
              dplyr::select(-c(Generation, Tank, Date, Percent_Survival, proportion)) %>% 
              dplyr::group_by(Larvae_pCO2, Age) %>% 
              dplyr::summarise(mean_Count_alive = mean(Count_alive), 
                               mean_Count_dead = mean(Count_dead))
# call datafarme binary for plotting
binary_df_F2_MEANS       <- data.frame() # start dataframe 

# run it 
for (i in 1:nrow(df_F2_MEANS)) {
   count_alive <- round(df_F2_MEANS[i,]$mean_Count_alive/1000) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F2_MEANS[i,]$mean_Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   loopDF           <- data.frame(matrix(0,ncol = 3, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Larvae_pCO2','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Larvae_pCO2 = df_F2_MEANS[i,]$Larvae_pCO2,
                            Age         = df_F2_MEANS[i,]$Age
                          ) 
  loopDF[c(1:count_dead),3] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),3] = 0 # 0 for alive

  loopDF             <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F2_MEANS <- rbind(binary_df_F2_MEANS,loopDF) #bind to a cumulative list dataframe
  
}
binary_df_F2_MEANS <- binary_df_F2_MEANS %>% 
                      dplyr::mutate(Generation = 2) %>% 
                      na.omit()
```

### Rbind data for F1_F2 comparisons
* NOTE: omit the 'high' treamtent in F1 as tihs was not present in F2 larvae
```{r merge datasets}
# all - with tank replicate included
binary_df_F1_F2_all             <- rbind(binary_df_F1_all, binary_df_F2_all) %>% 
                                         dplyr::filter(!(Larvae_pCO2 == 'High pCO2')) %>% 
                                         dplyr::mutate(Prop_time_metamophosis = 
                                                            case_when(Generation == 1 ~ (Age / 10),
                                                                      Generation == 2 ~ (Age / 14))) %>% 
                                         # omit timpoints after 10 days for Gen 1 and after 14 days for Gen 2
                                         dplyr::filter(!Prop_time_metamophosis >1) 
binary_df_F1_F2_all$Larvae_pCO2 <- factor(binary_df_F1_F2_all$Larvae_pCO2, exclude = 'High pCO2')

#  by mean tank
binary_df_F1_F2_MEANS             <- rbind(binary_df_F1_MEANS, binary_df_F2_MEANS) %>% 
                                           dplyr::filter(!(Larvae_pCO2 == 'High pCO2')) %>% 
                                           dplyr::mutate(Prop_time_metamophosis = 
                                                            case_when(Generation == 1 ~ (Age / 10),
                                                                      Generation == 2 ~ (Age / 14))) %>% 
                                         # omit timpoints after 10 days for Gen 1 and after 14 days for Gen 2
                                         dplyr::filter(!Prop_time_metamophosis >1) 
binary_df_F1_F2_MEANS$Larvae_pCO2 <- factor(binary_df_F1_F2_MEANS$Larvae_pCO2, exclude = 'High pCO2')


```
## Truncate binary master files by the larval period! 

* note that metamorphosis was observed at 10 dpf in F1s and 14 dpf in F2s, therfore 
larval survival comapred across gnerations should only show < 10 dpf survival 

```{r IMPORTNAT truncate <10DPF}
# all
df_F1_all_larvae    <- binary_df_F1_all %>% filter(Age <=10) # F1 - aligned timepoint prior to F1 metamorph for Gen 1 and 2 comparisons
df_F2_all_larvae    <- binary_df_F2_all %>% filter(Age <=10) # F2- aligned timepoint prior to F1 metamorph for Gen 1 and 2 comparisons
df_F1_F2_all_larvae <- binary_df_F1_F2_all  %>% filter(!Prop_time_metamophosis >1) # omit timpoints after 10 days for Gen 1 and after 14 days for Gen 2
df_F1_F2_all_larvaeLOW <- df_F1_F2_all_larvae %>% filter(Larvae_pCO2 %in% "Low pCO2")
df_F1_F2_all_larvaeMOD <- df_F1_F2_all_larvae %>% filter(Larvae_pCO2 %in% "Moderate pCO2")
# mean data by tank replicate within treatment
df_F1_MEANS_larvae    <- binary_df_F1_MEANS %>% filter(Age <=10) # F1 - aligned timepoint prior to F1 metamorph for Gen 1 and 2 comparisons 
df_F2_MEANS_larvae    <- binary_df_F2_MEANS %>% filter(Age <=10) # F2- aligned timepoint prior to F1 metamorph for Gen 1 and 2 comparisons
df_F1_F2_MEANS_larvae <- binary_df_F1_F2_MEANS  %>% filter(!Prop_time_metamophosis >1) # omit timpoints after 10 days for Gen 1 and after 14 days for Gen 2
df_F1_F2_MEANS_larvaeLOW <- df_F1_F2_MEANS_larvae %>% filter(Larvae_pCO2 %in% "Low pCO2") # comparisons to ACTUAL developmental timing as proportion time
df_F1_F2_MEANS_larvaeMOD <- df_F1_F2_MEANS_larvae %>% filter(Larvae_pCO2 %in% "Moderate pCO2") # comparisons to ACTUAL developmental timing as proportion time
# what days are included?
unique(df_F1_all_larvae$Age) # 2  4  8 10
unique(df_F2_all_larvae$Age) # 2  6  8 10
unique(df_F1_F2_all_larvae$Age) #  2  4  8 10  6 14

unique(df_F1_MEANS_larvae$Age) # 2  4  8 10
unique(df_F2_MEANS_larvae$Age) # 2  6  8 10
unique(df_F1_F2_MEANS_larvae$Age) #  2  4  8 10  6 14
```

## Surv object with all data v. mean data

  * note you can see the CI for the 'all' data surv object is small - these are 0s and 1s for each

  * note there is a large CI for the 'mean' data 
  
```{r F1 & F2 create surv object}

s1_all <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS)


# convert Age to the proportion of time to metamorphosis 
df_F1_MEANS_larvae$Prop_time_metamophosis <- (df_F1_MEANS_larvae$Age / 10)
df_F2_MEANS_larvae$Prop_time_metamophosis <- (df_F2_MEANS_larvae$Age / 14)



s1          <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = df_F1_MEANS_larvae)
s1_proptime <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Larvae_pCO2, data = df_F1_MEANS_larvae)


s2          <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = df_F2_MEANS_larvae)
s2_proptime <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Larvae_pCO2, data = df_F2_MEANS_larvae)


s1_2          <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2+Generation, data = df_F1_F2_MEANS_larvae)
s1_2_proptime <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Larvae_pCO2+Generation, data = df_F1_F2_MEANS_larvae)




s1_2LOW          <- survfit2(Surv(Age, Count_dead) ~ Generation, data = df_F1_F2_MEANS_larvaeLOW)
s1_2proptimeLOW  <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Generation, data = df_F1_F2_MEANS_larvaeLOW)

s1_2MOD          <- survfit2(Surv(Age, Count_dead) ~ Generation, data = df_F1_F2_MEANS_larvaeMOD)
s1_2proptimeMOD  <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Generation, data = df_F1_F2_MEANS_larvaeMOD)
```
# Kaplan Meier plots 
* F1 - all treatments and as age & proportion time to metamorphosis
```{r F1 plot}

F1_larvaeKaplanMeier <- s1 %>% 
                        ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                        scale_color_manual(values = c("forestgreen","orange","purple")) +
                        scale_fill_manual(values = c("forestgreen","orange","purple")) +
                        labs(
                          x = "Days",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        add_risktable() +
                        add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:10)) #+
                       # add_pvalue(location  = "annotation", x = 8.5)

F1_larvaeKaplanMeier <- s1_proptime %>% 
                        ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                        scale_color_manual(values = c("forestgreen","orange","purple")) +
                        scale_fill_manual(values = c("forestgreen","orange","purple")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        add_risktable() +
                        add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                       # add_pvalue(location  = "annotation", x = 8.5)
```
* F2 - all treatments and as age &  proportion time to metamorphosis
```{r F2 plot}

F2_larvaeKaplanMeier <- s2 %>% 
                        ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                        scale_color_manual(values = c("forestgreen","orange")) +
                        scale_fill_manual(values = c("forestgreen","orange")) +
                        labs(
                          x = "Days",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        add_risktable() +
                        add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:10)) #+
                       # add_pvalue(location  = "annotation", x = 8.5)


F2_larvaeKaplanMeier <- s2_proptime %>% 
                        ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                        scale_color_manual(values = c("forestgreen","orange")) +
                        scale_fill_manual(values = c("forestgreen","orange")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        add_risktable() +
                        add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:10)) #+
                       # add_pvalue(location  = "annotation", x = 8.5)
```


* F1 & F2 - low vs, mod and ONLY as proportion time to metamorphosis
```{r F1 & F2 plot low v. mod treatments as prop time to metamorphosis }

F1_F2_larvaeKaplanMeier <- s1_2_proptime %>% 
                        ggsurvfit(linetype_aes = FALSE, linewidth = 1) +
                        scale_color_manual(values = c("green","forestgreen",
                                                      "orange","darkorange4")) +
                        scale_fill_manual(values = c("green","forestgreen",
                                                     "orange","darkorange4")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                     #   add_pvalue(location  = "annotation", x = 8.5)


# output plot 
pdf(paste0("Output/Survival/F1_F2_Mod_Low_Survival_ggsurvfit.pdf"),width=8, height=4)
print(F1_F2_larvaeKaplanMeier)
dev.off()
```

* F1 & F2 - Explore generation effect within treatment - low ONLY as proportion time to metamorphosis
```{r F1 & F2 plot low ONLY as prop time to metamorphosis }

F1_F2_larvaeKaplanMeierLOW <- s1_2proptimeLOW %>% 
                        ggsurvfit(linetype_aes = FALSE, linewidth = 1) +
                        scale_color_manual(values = c("green","forestgreen")) +
                        scale_fill_manual(values = c("green","forestgreen")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        add_risktable() +
                        add_risktable_strata_symbol() + # symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                     #   add_pvalue(location  = "annotation", x = 8.5)
```
* F1 & F2 - Explore generation effect within treatment - moderate ONLY as proportion time to metamorphosis
```{r F1 & F2 plot moderate ONLY as prop time to metamorphosis }

F1_F2_larvaeKaplanMeierMOD <- s1_2proptimeMOD %>% 
                        ggsurvfit(linetype_aes = FALSE, linewidth = 1) +
                        scale_color_manual(values = c("orange","darkorange")) +
                        scale_fill_manual(values = c("orange","darkorange")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        add_risktable() +
                        add_risktable_strata_symbol() + # symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                     #   add_pvalue(location  = "annotation", x = 8.5)
```


## Estimating x-time survival
* per 10 days post fertilization?
```{r}

F1_10d_surv_prob <- s1 %>% 
                          tbl_survfit(
                            times = 10,
                            label_header = "**survival to 10 days post-fertilization (95% CI)**"
                          )

F2_10d_surv_prob <- s2 %>% 
                          tbl_survfit(
                            times = 10,
                            label_header = "**survival to 10 days post-fertilization (95% CI)**"
                          )


F1_F2_10d_surv_prob <- s1_2 %>% 
                          tbl_survfit(
                            times = 10,
                            label_header = "**survival to 10 days post-fertilization (95% CI)**"
                          )

```

## Estimating x-time survival
* per 8 days from day 2 to 18 post fertilization
```{r}

F1_surv_prob <- s1_all %>% 
                          tbl_survfit(
                            times = 18,
                            label_header = "**survival to 20 days post-fertilization (95% CI)**")
F1_surv_prob$table_body$stat_1[2]
F1_surv_prob$table_body$stat_1[3]


# for loop every 50 days between 100 - 450 days post fertilization to a summary table!
x <- seq.int(2,18, by = 8) # call the integer
summarytable <- data.frame() # summary table 
for (i in 1:length(x)) { # for loop
  survtable <- s1_all %>% tbl_survfit(times = x[i])
  survtable$table_body$stat_1[2]
  survtable$table_body$stat_1[3]
  loopDF  <- data.frame(matrix(0,ncol = 4, nrow = 1)) 
  colnames(loopDF) = c('Age','LowpCO2_Probabilty','ModpCO2_Probabilty','Diff')
  loopDF  <- loopDF %>% 
        dplyr::mutate(
                      Age                = x[i],
                      LowpCO2_Probabilty = survtable$table_body$stat_1[2], 
                      ModpCO2_Probabilty = survtable$table_body$stat_1[3],
                      Diff               = abs(
                                           as.numeric(substr(survtable$table_body$stat_1[2],1,2)) -
                                           as.numeric(substr(survtable$table_body$stat_1[3],1,2)) 
                                           )
                      ) 
  loopDF       <- data.frame(loopDF) # name dataframe for this single row
  summarytable <- rbind(summarytable,loopDF) #bind to a cumulative list dataframe
}
```

# Comparing survival times between groups (log-rank test)
* We can conduct between-group significance tests using a log-rank test. The log-rank test equally weights observations over the entire follow-up time and is the most common way to compare survival times between groups. There are versions that more heavily weight the early or late follow-up that could be more appropriate depending on the research question (see ?survdiff for different test options).

* We get the log-rank p-value using the survdiff function
```{r log-rank test survdiff}

F1_logrank_test <- survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = df_F1_MEANS_larvae)
#                              N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2      1340     1154     1216      3.14      10.6
# Larvae_pCO2=Moderate pCO2 1084      946     1066     13.46      42.5
# Larvae_pCO2=High pCO2      805      762      580     56.82     126.1

F2_logrank_test <- survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = df_F2_MEANS_larvae)
#                             N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2      280      199      188     0.688      1.98
# Larvae_pCO2=Moderate pCO2 403      268      279     0.462      1.98

F1_F2_logrank_test <- survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2+Generation, data = df_F1_F2_MEANS_larvae)
#                                            N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2, Generation=1      1340     1154      977      32.2      84.2
# Larvae_pCO2=Low pCO2, Generation=2       334      250      343      25.2      54.7
# Larvae_pCO2=Moderate pCO2, Generation=1 1084      946      861       8.3      20.3
# Larvae_pCO2=Moderate pCO2, Generation=2  481      335      504      56.7     131.7
```


# The Cox regression model

 * We may want to quantify an effect size for a single variable, or include **more than one variable** into a regression model to account for the effects of multiple variables.

* The Cox regression model is a semi-parametric model that can be used to fit univariable and multivariable regression models that have survival outcomes.

```{r cox for multiple variables - random tank effect with all data}
library(survminer) # ggforect hazard ratio plot

# F1 all treatments
F1_coxph_test       <- coxph(formula = Surv(Age, Count_dead) ~ Larvae_pCO2+(1|Tank), data = df_F1_all_larvae)
F1_hazardratio_plot <- ggforest(F1_coxph_test, data = df_F1_all_larvae)
F1_coxph_test_table <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2+(1|Tank), data = df_F1_all_larvae) %>%  tbl_regression(exp = TRUE) 
#                              coef exp(coef) se(coef)      z        p
# Larvae_pCO2Moderate pCO2 -0.08286   0.92048  0.01939 -4.272 1.93e-05
# Larvae_pCO2High pCO2      0.46824   1.59718  0.02121 22.080  < 2e-16
# 1 | TankTRUE                   NA        NA  0.00000     NA       NA

F2_coxph_test <- coxph(formula = Surv(Age, Count_dead) ~ Larvae_pCO2+(1|Tank), data = df_F2_all_larvae)
F2_hazardratio_plot <- ggforest(F2_coxph_test, data = df_F2_all_larvae)
F2_coxph_test_table <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, data = df_F2_all_larvae) %>% 
                              tbl_regression(exp = TRUE) 
#                              coef exp(coef) se(coef)      z       p
# Larvae_pCO2Moderate pCO2 -0.15145   0.85946  0.05402 -2.803 0.00506

df_F1_F2_all_larvae$Generation <- as.character(df_F1_F2_all_larvae$Generation)
F1_F2_coxph_test <- coxph(formula = Surv(Pro, Count_dead) ~ Larvae_pCO2+Generation+(1|Tank), data = df_F1_F2_all_larvae)
F1_F2_hazardratio_plot <- ggforest(F1_F2_coxph_test, data = df_F1_F2_all_larvae)
F1_F2_coxph_test_table <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2*Generation+(1|Tank), data = df_F1_F2_all_larvae) %>% tbl_regression(exp = TRUE) 
#                                         coef exp(coef) se(coef)      z        p
# Larvae_pCO2Moderate pCO2            -0.01726   0.98289  0.06645 -0.260    0.795
# Generation                          -0.25022   0.77863  0.04297 -5.823 5.79e-09
# 1 | TankTRUE                              NA        NA  0.00000     NA       NA
# Larvae_pCO2Moderate pCO2:Generation -0.06632   0.93584  0.05733 -1.157    0.247



# Low gen 1 vs Low Gen 2
df_F1_F2_all_larvaeLOW$Generation <- as.character(df_F1_F2_all_larvaeLOW$Generation)
F1_F2_coxph_testLOW <- coxph(formula = Surv(Age, Count_dead) ~ Generation, 
                            data = df_F1_F2_all_larvaeLOW)
F1_F2_hazardratio_plotLOW <- ggforest(F1_F2_coxph_testLOW, data = df_F1_F2_all_larvaeLOW)


# Mod gen 1 vs Low Gen 2
df_F1_F2_all_larvaeMOD$Generation <- as.character(df_F1_F2_all_larvaeMOD$Generation)
F1_F2_coxph_testMOD <- coxph(formula = Surv(Age, Count_dead) ~ Generation, 
                            data = df_F1_F2_all_larvaeMOD)
F1_F2_hazardratio_plotMOD <- ggforest(F1_F2_coxph_testMOD, data = df_F1_F2_all_larvaeMOD)


```


```{r cox  for multiple variables - mean data}
library(survminer) # ggforect hazard ratio plot


# F1 - low mod and high  AGE
F1_coxph_testMEAN        <- coxph(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, 
                                  data = df_F1_MEANS_larvae)
F1_hazardratio_plotMEAN <- ggforest(F1_coxph_testMEAN, data = df_F1_MEANS_larvae)
F1_coxph_test_tableMEAN <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2,
                                 data = df_F1_MEANS_larvae) %>%  tbl_regression(exp = TRUE) 
#                              coef exp(coef) se(coef)      z      p
# Larvae_pCO2Moderate pCO2 -0.08374   0.91967  0.04337 -1.931 0.0535
# Larvae_pCO2High pCO2      0.46589   1.59343  0.04740  9.830 <2e-16







# F2 - low mod and high  AGE
df_F2_MEANS_larvae$Larvae_pCO2 <- droplevels(df_F2_MEANS_larvae$Larvae_pCO2) 
levels(df_F2_MEANS_larvae$Larvae_pCO2) <- c('Low pCO2', 'Moderate pCO2')
F2_coxph_testMEANS       <- coxph(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, 
                                  data = df_F2_MEANS_larvae)
F2_hazardratio_plotMEANS <- ggforest(F2_coxph_testMEANS, data = df_F2_MEANS_larvae)
F2_coxph_test_tableMEANS <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, 
                                  data = df_F1_MEANS_larvae) %>% tbl_regression(exp = TRUE) 
#                              coef exp(coef) se(coef)      z     p
# Larvae_pCO2Moderate pCO2 -0.14927   0.86133  0.09371 -1.593 0.111






# F1 and F2 - as AGE
df_F1_F2_MEANS_larvae$Generation <- as.character(df_F1_F2_MEANS_larvae$Generation) # make generation a character not a integer
df_F1_F2_MEANS_larvae$Gen_pCO2   <- paste0(df_F1_F2_MEANS_larvae$Generation, df_F1_F2_MEANS_larvae$Larvae_pCO2) # new column for hazard ratio
F1_F2_coxph_testBOTH       <- coxph(formula = Surv(Prop_time_metamophosis, Count_dead) ~ Gen_pCO2, 
                                    data = df_F1_F2_MEANS_larvae)
F1_F2_hazardratio_plotBOTH <- ggforest(F1_F2_coxph_testBOTH, data = df_F1_F2_MEANS_larvae)


pdf("Output/Survival/F1_F2_Mod_Low_Survival_hazardratio.pdf",width=6, height=2)
print(F1_F2_hazardratio_plotBOTH)
dev.off()

# F1 and F2 WITHIN TREATMENT AND AS PROPORTION TIME TO METAMORPHOSIS
# Low gen 1 vs Low Gen 2
df_F1_F2_MEANS_larvaeLOW$Generation <- as.character(df_F1_F2_MEANS_larvaeLOW$Generation)
F1_F2_coxph_testLOW <- coxph(formula = Surv(Prop_time_metamophosis, Count_dead) ~ Generation, 
                            data = df_F1_F2_MEANS_larvaeLOW)
F1_F2_hazardratio_plotLOW <- ggforest(F1_F2_coxph_testLOW, data = df_F1_F2_MEANS_larvaeLOW)


# Mod gen 1 vs Low Gen 2
df_F1_F2_MEANS_larvaeMOD$Generation <- as.character(df_F1_F2_MEANS_larvaeMOD$Generation)
F1_F2_coxph_testMOD <- coxph(formula = Surv(Prop_time_metamophosis, Count_dead) ~ Generation, 
                            data = df_F1_F2_MEANS_larvaeMOD)
F1_F2_hazardratio_plotMOD <- ggforest(F1_F2_coxph_testMOD, data = df_F1_F2_MEANS_larvaeMOD)

```

```{r}
# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F1_all  <- Surv(time = as.numeric(binary_df_F1_all$Age), 
                     event = as.numeric(binary_df_F1_all$Count_dead), type = "right")

# plot the surv data
ggsurvplot(survfit(surv_F1_all ~ Larvae_pCO2, binary_df_F1_all), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")




# (2) Mean counts alive and dead by tank - better presentative for plots!
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F1_means_all  <- Surv(time = as.numeric(binary_df_F1_MEANS$Age), 
                     event = as.numeric(binary_df_F1_MEANS$Count_dead), type = "right")
# with the mean data by  tank replicate! better - save this one
ggsurvplot(survfit(surv_F1_means_all ~ Larvae_pCO2, binary_df_F1_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")

s1 <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS)
str(s1)

View(binary_df_F1_MEANS)
s1 %>% 
  ggsurvfit(linetype_aes = TRUE) +
  scale_color_manual(values = c("forestgreen","orange","purple")) +
  scale_fill_manual(values = c("forestgreen","orange","purple")) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(
    x = "Days",
    y = "Overall survival probability"
  )

?ggsurvfit
s1 %>% 
  ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
  scale_color_manual(values = c("forestgreen","orange","purple")) +
  scale_fill_manual(values = c("forestgreen","orange","purple")) +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) + 
  add_confidence_interval() +
  add_risktable() +
  add_risktable_strata_symbol(symbol = "\U25CF", size = 10) +
  scale_ggsurvfit(x_scales = list(breaks = 0:10)) +
  add_pvalue(location  = "annotation", x = 8.5)

# use gtsummary 
s1 %>% 
  tbl_survfit(
    times = 10,
    label_header = "**survival to 10 days post-fertilization (95% CI)**"
  )
s1 %>% 
  tbl_survfit(
    probs = 0.5,
    label_header =  "**Median survival to 10 days post-fertilization (95% CI)**"
  )
s1
# Call: survfit(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS)
# 
#    2 observations deleted due to missingness 
#                              n events median 0.95LCL 0.95UCL
# Larvae_pCO2=Low pCO2      1844   1631      8       8       8
# Larvae_pCO2=Moderate pCO2 1460   1304      8       8       8
# Larvae_pCO2=High pCO2      805    762      4       4       4
survdiff(Surv(Age, Count_dead) ~ Larvae_pCO2 + (1|Tank), data = binary_df_F1_all)
binary_df_F1_all$Tank <- as.factor(as.numeric(binary_df_F1_all$Tank))
# Call:
# survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS)
# 
# n=4109, 2 observations deleted due to missingness.
# 
#                              N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2      1844     1631     1917     42.67    144.62
# Larvae_pCO2=Moderate pCO2 1460     1304     1372      3.38      8.63
# Larvae_pCO2=High pCO2      805      762      408    307.29    500.45


coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS)
coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS) %>% 
  tbl_regression(exp = TRUE) 


?survfit2
Surv(lung$time, lung$status)[1:10]
Surv(binary_df_F1_MEANS$Age, binary_df_F1_MEANS$Count_dead)[1:10]
head(lung[, c("time", "status", "sex")])

# plot the surv data
pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F1_survival.pdf")
ggsurvplot(survfit(surv_F1_means_all ~ Larvae_pCO2, binary_df_F1_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")
dev.off()

```
### Survival status 
* run the stats with the 'all' data and address the Tank as random effect
```{r F1 run coxme on surv stats}

head(binary_df_F1_all)
# use this data from the first fr loop - has live and dead counts with reoslution of treatment AND replicate tank!!!

# now run the model
F1_survFIT_cox <- coxme(surv_F1_all ~ Larvae_pCO2 + (1|Tank), binary_df_F1_all)
summary(F1_survFIT_cox)
#                               coef exp(coef)   se(coef)     z p
# Larvae_pCO2Moderate pCO2 0.1969625  1.217698 0.01697219 11.61 0
# Larvae_pCO2High pCO2     1.1161450  3.053062 0.02143035 52.08 0



#The standard deviation of the tank random effect is very small (0.006017691) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (312321 ), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 


#pairwise comparisons

library(multcomp)
glht(F1_survFIT_cox, linfct = mcp(Larvae_pCO2 = "Dunnett"),alternative = "greater")
# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0   0.1059
# High pCO2 - Low pCO2 <= 0       0.8664


```

```{r F1 run coxph on surv stats}
?coxme
head(binary_df_F1_all)
# use this data from the first fr loop - has live and dead counts with reoslution of treatment AND replicate tank!!!

# now run the model
F1_survFIT_cox <- coxme(surv_F1_all ~ Larvae_pCO2 + (1|Tank), binary_df_F1_all)
summary(F1_survFIT_cox)

#                               coef exp(coef)   se(coef)     z p
# Larvae_pCO2Moderate pCO2 0.1969625  1.217698 0.01697219 11.61 0
# Larvae_pCO2High pCO2     1.1161450  3.053062 0.02143035 52.08 0



#The standard deviation of the tank random effect is very small (0.006017691) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (312321 ), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 


#pairwise comparisons

library(multcomp)
?glht
glht(F1_survFIT_cox, linfct = mcp(Larvae_pCO2 = "Dunnett"),alternative = "greater")
# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0   0.1059
# High pCO2 - Low pCO2 <= 0       0.8664


```



```{r F2 build surv object and plot em}


# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F2_all  <- Surv(time = as.numeric(binary_df_F2_all$Age), 
                     event = as.numeric(binary_df_F2_all$Count_dead), type = "right")
# plot the surv data
ggsurvplot(survfit(surv_F2_all ~ Larvae_pCO2, binary_df_F2_all), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")




# (2) Mean counts alive and dead by tank - better presentative for plots!
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F2_means_all  <- Surv(time = as.numeric(binary_df_F2_MEANS$Age), 
                     event = as.numeric(binary_df_F2_MEANS$Count_dead), type = "right")
# with the mean data by  tank replicate! better - save this one
ggsurvplot(survfit(surv_F2_means_all ~ Larvae_pCO2, binary_df_F2_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange"), c(1,0.8)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")

# plot the surv data
pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F2_survival.pdf")
ggsurvplot(survfit(surv_F2_means_all ~ Larvae_pCO2, binary_df_F2_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")
dev.off()

```

```{r F2 run coxme on surv stats}

head(binary_df_F2_all)
# use this data from the first fr loop - has live and dead counts with reoslution of treatment AND replicate tank!!!

# omit the level high pCO2 because it is not actually present - to be a carry over from the full dataset?
is.na(binary_df_F2_all$Larvae_pCO2) <- binary_df_F2_all$Larvae_pCO2 == "High pCO2"
binary_df_F2_all$Larvae_pCO2 <- factor(binary_df_F2_all$Larvae_pCO2)
str(binary_df_F2_all) # Larvae_pCO2 now has two factors - correct!!

# now run the model
F2_survFIT_cox <- coxme(surv_F2_all ~ Larvae_pCO2 + (1|Tank), binary_df_F2_all)
summary(F2_survFIT_cox)

#                               coef exp(coef)   se(coef)   z      p
# Larvae_pCO2Moderate pCO2 0.1269179  1.135324 0.04536691 2.8 0.0051


#The standard deviation of the tank random effect is very small (0.006017691) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (312321 ), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 


#pairwise comparisons

library(multcomp)

glht(F2_survFIT_cox, linfct = mcp(Larvae_pCO2 = "Dunnett"),alternative = "greater")

# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0   0.1269


```



# F1 and F2 responses across generations 
  * truncate data to day 14 that overlaps between both datasets 
  * plot and run stats 
```{r}


# filter and merge up to d14 (common between datasets)

# all data
unique(binary_df_F1_all$Age)
binary_df_F1_d10 <- binary_df_F1_all %>% dplyr::filter(!Age %in% c(14,18)) %>% na.omit()
unique(binary_df_F1_d10$Age) #  2  4  8 10
unique(binary_df_F2_all$Age)
binary_df_F2_d10 <- binary_df_F2_all %>% dplyr::filter(!Age %in% c(14,21)) %>% na.omit()
unique(binary_df_F2_d10$Age) # 2  6  8 10

binary_F1_F2_master   <- rbind(binary_df_F1_d10, binary_df_F2_d10) %>% dplyr::filter(!(Generation ==1 & 
                                                                                       Larvae_pCO2 %in% 'High pCO2'))
unique(binary_F1_F2_master$Larvae_pCO2)
binary_F1_F2_moderate <- binary_F1_F2_master %>% dplyr::filter(Larvae_pCO2 %in% 'Moderate pCO2')
binary_F1_F2_low      <- binary_F1_F2_master %>% dplyr::filter(Larvae_pCO2 %in% 'Low pCO2')

# by means
unique(binary_df_F2_MEANS$Age)
binary_df_F1_d10 <- binary_df_F1_MEANS %>% dplyr::filter(!Age %in% c(14,18)) %>% na.omit()
unique(binary_df_F1_d10$Age) #  2  4  8 10
unique(binary_df_F2_MEANS$Age)
binary_df_F2_d10 <- binary_df_F2_MEANS %>% dplyr::filter(!Age %in% c(14,21)) %>% na.omit()
unique(binary_df_F2_d10$Age) # 2  6  8 10

binary_F1_F2_masterMEANS   <- rbind(binary_df_F1_d10, binary_df_F2_d10) %>% dplyr::filter(!(Generation ==1 & 
                                                                                       Larvae_pCO2 %in% 'High pCO2'))
binary_F1_F2_moderateMEANS <- binary_F1_F2_masterMEANS %>% dplyr::filter(Larvae_pCO2 %in% 'Moderate pCO2')
binary_F1_F2_lowMEANS      <- binary_F1_F2_masterMEANS %>% dplyr::filter(Larvae_pCO2 %in% 'Low pCO2')



# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting
# build the surv object! 
# with tank included in the data frame (first for loop above)


surv_F1_F2_allMEANS  <- Surv(time = as.numeric(binary_F1_F2_masterMEANS$Age), 
                     event = as.numeric(binary_F1_F2_masterMEANS$Count_dead), type = "right")


surv_F1_F2_all  <- Surv(time = as.numeric(binary_F1_F2_master$Age), 
                     event = as.numeric(binary_F1_F2_master$Count_dead), type = "right")
surv_F1_F2_modMEANS  <- Surv(time = as.numeric(binary_F1_F2_moderateMEANS$Age), 
                     event = as.numeric(binary_F1_F2_moderateMEANS$Count_dead), type = "right")
surv_F1_F2_lowMEANS  <- Surv(time = as.numeric(binary_F1_F2_lowMEANS$Age), 
                     event = as.numeric(binary_F1_F2_lowMEANS$Count_dead), type = "right")
# plot the surv data
# all 
ggsurvplot(survfit(surv_F1_F2_all ~ Generation+Larvae_pCO2, binary_F1_F2_masterMEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("yellowgreen","orange","darkgreen","darkorange"), 
                             c(1,0.8,1,0.8)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "Generation")
# low
# moderate
ggsurvplot(survfit(surv_F1_F2_modMEANS ~ Generation, binary_F1_F2_moderateMEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("darkorange","orange"), c(1,0.8)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,
           xlab = "Age (days post fertilization)", 
           legend.title = "Generation")
# low
ggsurvplot(survfit(surv_F1_F2_lowMEANS ~ Generation, binary_F1_F2_lowMEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("darkgreen","green"), c(1,0.8)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "Generation")


# Estimate median survival time 
survfit(Surv(as.numeric(Generation), Larvae_pCO2) ~ 1, data = binary_F1_F2_master)
survfit(Surv(as.numeric(Generation), Larvae_pCO2) ~ 1, data = binary_F1_F2_master) %>% 
  tbl_survfit(
    probs = 0.5,
    label_header = "**Median survival (95% CI)**"
  )

# statistics moderate
binary_F1_F2_master$Generation <- factor(binary_F1_F2_master$Generation)
str(binary_F1_F2_master) # Larvae_pCO2 now has two factors - correct!!
F1F2_survFIT_all_cox <- coxme(surv_F1_F2_all ~ Generation + (1|Tank), binary_F1_F2_master)
summary(F1F2_survFIT_Moderate_cox)


# as factor for the Gneration (the ID for this model)
binary_F1_F2_moderate$Generation <- factor(binary_F1_F2_moderate$Generation)
str(binary_F1_F2_moderate) # Larvae_pCO2 now has two factors - correct!!
# now run the model

survdiff(Surv(time, status) ~ sex, data = lung)


F1F2_survFIT_Moderate_cox <- coxme(surv_F1_F2_mod ~ Generation + (1|Tank), binary_F1_F2_moderate)
summary(F1F2_survFIT_Moderate_cox)
# Fixed coefficients
#                 coef exp(coef)   se(coef)     z       p
# Generation -0.208019 0.8121916 0.03484548 -5.97 2.4e-09
library(multcomp)
summary(glht(F1F2_survFIT_Moderate_cox, linfct = mcp(Generation = "Dunnett"),alternative = "greater"))
# Linear Hypotheses:
#            Estimate Std. Error z value Pr(>z)
# 2 - 1 <= 0 -0.20802    0.03485   -5.97      1



# statistics loq

# as factor for the Gneration (the ID for this model)
binary_F1_F2_low$Generation <- factor(binary_F1_F2_low$Generation)
str(binary_F1_F2_low) # Larvae_pCO2 now has two factors - correct!!
# now run the model
F1F2_survFIT_Low_cox <- coxme(surv_F1_F2_low ~ Generation + (1|Tank), binary_F1_F2_low)
summary(F1F2_survFIT_Low_cox)
# Fixed coefficients
#                   coef exp(coef)   se(coef)     z       p
# Generation2 -0.2371305 0.7888884 0.03939462 -6.02 1.8e-09
library(multcomp)
summary(glht(F1F2_survFIT_Low_cox, linfct = mcp(Generation = "Dunnett"),alternative = "greater"))
# Linear Hypotheses:
#            Estimate Std. Error z value Pr(>z)
# 2 - 1 <= 0 -0.23713    0.03939  -6.019      1


```
# F3 Larval Survival
## Plot it!
```{r F3 summary and plot}

stL_F3 <- summarySE(df_F3, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2", "Parent_pCO2"))
stL_F3 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Gen3_PlotMeanSE <- ggplot(stL_F3, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2))+ 
                        geom_line()+
                        geom_point()+ 
                        geom_errorbar(aes(ymin=Percent_Survival-se, ymax=Percent_Survival+se), width=.2,
                                      position=position_dodge(.1)) +
                        scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                        theme(legend.position="right", legend.direction="vertical", 
                              legend.title = element_blank())+ theme_bw() + 
                        labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                        ggtitle("F3 larvae survival (Mean +- SE)") + 
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              panel.background = element_blank(), 
                              axis.line = element_line(colour = "black"))+
                        facet_wrap("Parent_pCO2")
Gen3_PlotMeanSE


# Summarise Percent Deformities for plotting 
Gen3_mean_survival <- df_F3 %>% 
                          dplyr::select(Parent_pCO2,Larvae_pCO2,Age, Percent_Survival) %>% 
                          na.omit() %>% 
                          dplyr::group_by(Age, Parent_pCO2,Larvae_pCO2) %>% 
                          dplyr::summarise(mean_Perc_surv = mean(Percent_Survival), 
                                           n           = n(),
                                           sd_Perc_surv   = sd(Percent_Survival),
                                           se_Perc_surv   = sd_Perc_surv/(sqrt(n)))

# Barplots 
unique(Gen3_mean_survival$Age)
Gen3_barplots_Age2 <- Gen3_mean_survival %>% dplyr::filter(Age ==2) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 2 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)
Gen3_barplots_Age5 <- Gen3_mean_survival %>% dplyr::filter(Age ==5) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 5 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

Gen3_barplots_Age7 <- Gen3_mean_survival %>% dplyr::filter(Age ==7) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 7 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

Gen3_barplots_Age12 <- Gen3_mean_survival %>% dplyr::filter(Age ==12) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 12 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)
Gen3_barplots_Age16 <- Gen3_mean_survival %>% dplyr::filter(Age ==16) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 16 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

# print the plot

pdf("Output/Survival/F3_survival_meanSE.pdf", width=10, height=5)
print(Gen3_PlotMeanSE_All)
dev.off()
# 
# 
# pdf("Output/Survival/PercSurvival_Gen3_2DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age2)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_5DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age5)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_7DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age7)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_12DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age12)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_16DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age16)
# dev.off()

```

```{r F3 build binary live and dead column}
#make a surv object to pass into coxme
df_F3 <- df_F3 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(Count_dead = 
                          (1/as.numeric(proportion))*as.numeric(Count_alive)) # proportion is that of alive


# Run two for loops 

# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting

# keep tank to use as a random effect in the cox model!
# Call the cumulative dataframe that we will write to in the for loop below
binary_df_F3_all     <- data.frame() # start dataframe 
binary_df_F3_high    <- data.frame() # start dataframe 
binary_df_F3_mod     <- data.frame() # start dataframe 
binary_df_F3_low     <- data.frame() # start dataframe 
# run it 
for (i in 1:nrow(df_F3)) {
   #count_alive <- round(df_F3[i,]$Count_alive/1000) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_F3[i,]$Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   count_alive <- round(df_F3[i,]$Count_alive) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F3[i,]$Count_dead)  # divide all counts by 1000, round to nearest

   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_F3[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_F3[i,]$Larvae_pCO2,
                            Generation  = df_F3[i,]$Generation,
                            Tank        = df_F3[i,]$Tank,  
                            Age         = df_F3[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF       <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F3_all <- rbind(binary_df_F3_all,loopDF) #bind to a cumulative list dataframe
  # print(binary_df_F3) # print to monitor progress

     
  if (loopDF$Parent_pCO2[1] == 'High pCO2') {
      binary_df_F3_high <- rbind(binary_df_F3_high,loopDF) #bind to a cumulative list dataframe
  } else if (loopDF$Parent_pCO2[1] == 'Moderate pCO2') {
      binary_df_F3_mod <- rbind(binary_df_F3_mod,loopDF) #bind to a cumulative list dataframe
  } else {
      binary_df_F3_low <- rbind(binary_df_F3_low,loopDF) #bind to a cumulative list dataframe
    }
      
  
}

binary_df_F3_all$All_pCO2 <- as.factor(paste0(binary_df_F3_all$Parent_pCO2, ' x ', binary_df_F3_all$Larvae_pCO2))


# (2) Mean counts alive and dead by tank - better presentative for plots!

# for loop for plotting!! 
# difference here is we take a mean for counts alive and dead within treatment, no tank included and a singluar line for plotting 
df_F3_MEANS <- df_F3 %>% 
  dplyr::select(-c(Generation, Tank, Date, Percent_Survival, proportion)) %>% 
  dplyr::group_by(Parent_pCO2, Larvae_pCO2, Age) %>% 
  dplyr::summarise(mean_Count_alive = mean(Count_alive), 
            mean_Count_dead = mean(Count_dead))
# call datafarme binary for plotting
binary_df_F3_MEANS       <- data.frame() # start dataframe 
binary_df_F3_MEANS_high  <- data.frame() # start dataframe 
binary_df_F3_MEANS_mod   <- data.frame() # start dataframe 
binary_df_F3_MEANS_low   <- data.frame() # start dataframe 

# run it 
for (i in 1:nrow(df_F3_MEANS)) {
   #count_alive <- round(df_F3_MEANS[i,]$mean_Count_alive/1000) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_F3_MEANS[i,]$mean_Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   count_alive <- round(df_F3_MEANS[i,]$mean_Count_alive) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F3_MEANS[i,]$mean_Count_dead)  # divide all counts by 1000, round to nearest

   
   loopDF           <- data.frame(matrix(0,ncol = 4, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_F3_MEANS[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_F3_MEANS[i,]$Larvae_pCO2,
                            Age         = df_F3_MEANS[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),4] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),4] = 0 # 0 for alive

  loopDF             <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F3_MEANS <- rbind(binary_df_F3_MEANS,loopDF) #bind to a cumulative list dataframe
     
  if (loopDF$Parent_pCO2[1] == 'High pCO2') {
      binary_df_F3_MEANS_high <- rbind(binary_df_F3_MEANS_high,loopDF) #bind to a cumulative list dataframe
  } else if (loopDF$Parent_pCO2[1] == 'Moderate pCO2') {
      binary_df_F3_MEANS_mod <- rbind(binary_df_F3_MEANS_mod,loopDF) #bind to a cumulative list dataframe
  } else {
      binary_df_F3_MEANS_low <- rbind(binary_df_F3_MEANS_low,loopDF) #bind to a cumulative list dataframe
  }
    
}
```

```{r F3 build surv object}
nrow(binary_df_F3_all)
# create surv object
# with tank included in the data frame (first for loop above)
surv_F3_all  <- Surv(time = as.numeric(binary_df_F3_all$Age), 
                     event = as.numeric(binary_df_F3_all$Count_dead), type = "right")

surv_F3_high <- Surv(time = as.numeric(binary_df_F3_high$Age), 
                     event = as.numeric(binary_df_F3_high$Count_dead), type = "right")

surv_F3_mod  <- Surv(time = as.numeric(binary_df_F3_mod$Age), 
                     event = as.numeric(binary_df_F3_mod$Count_dead), type = "right")

surv_F3_low  <- Surv(time = as.numeric(binary_df_F3_low$Age), 
                     event = as.numeric(binary_df_F3_low$Count_dead), type = "right")


F3_low_survplot <- ggsurvplot(survfit(surv_F3_low ~ Larvae_pCO2, binary_df_F3_low), 
                       risk.table = F, pval = F , conf.int = TRUE,
                       palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8, 0.3)), 
                       ggtheme = theme_survminer(),
                       break.x.by = 2 ,xlab = "Age (days post fertilization)", 
                       legend.title = "treatment")

F3_mod_survplot <- ggsurvplot(survfit(surv_F3_mod ~ Larvae_pCO2, binary_df_F3_mod), 
                       risk.table = F, pval = F , conf.int = TRUE,
                       palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8, 0.3)), 
                       ggtheme = theme_survminer(),
                       break.x.by = 2 ,xlab = "Age (days post fertilization)", 
                       legend.title = "treatment")

F3_high_survplot <- ggsurvplot(survfit(surv_F3_high ~ Larvae_pCO2, binary_df_F3_high), 
                       risk.table = F, pval = F , conf.int = TRUE,
                       palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8, 0.3)), 
                       ggtheme = theme_survminer(),
                       break.x.by = 2 ,xlab = "Age (days post fertilization)", 
                       legend.title = "treatment")

F3_full_plot <- ggarrange(F3_low_survplot, F3_mod_survplot, F3_high_survplot, ncol = 3)
print(F3_full_plot)


# create surv object
# without tank included - means wihtin treatment SEPCIFICALLY FOR PLOTTING  second for loop above 
surv_F3_MEANS  <- Surv(time = as.numeric(binary_df_F3_MEANS$Age), 
                     event = as.numeric(binary_df_F3_MEANS$Count_dead), type = "right")


is.na(binary_df_F3_MEANS_low$Parent_pCO2) <- binary_df_F3_MEANS_low$Parent_pCO2 == "None"
binary_df_F3_MEANS_low$Parent_pCO2 <- factor(binary_df_F3_MEANS_low$Parent_pCO2)
str(binary_df_F3_MEANS_low) # youll see now its fixed

surv_F3_MEANS_low  <- Surv(time = as.numeric(binary_df_F3_MEANS_low$Age), 
                     event = as.numeric(binary_df_F3_MEANS_low$Count_dead), type = "right")

is.na(binary_df_F3_MEANS_mod$Parent_pCO2) <- binary_df_F3_MEANS_mod$Parent_pCO2 == "None"
binary_df_F3_MEANS_mod$Parent_pCO2 <- factor(binary_df_F3_MEANS_mod$Parent_pCO2)
str(binary_df_F3_MEANS_mod) # youll see now its fixed

surv_F3_MEANS_mod  <- Surv(time = as.numeric(binary_df_F3_MEANS_mod$Age), 
                     event = as.numeric(binary_df_F3_MEANS_mod$Count_dead), type = "right")


is.na(binary_df_F3_MEANS_high$Parent_pCO2) <- binary_df_F3_MEANS_high$Parent_pCO2 == "None"
binary_df_F3_MEANS_high$Parent_pCO2 <- factor(binary_df_F3_MEANS_high$Parent_pCO2)
str(binary_df_F3_MEANS_high) # youll see now its fixed

surv_F3_MEANS_high  <- Surv(time = as.numeric(binary_df_F3_MEANS_high$Age), 
                     event = as.numeric(binary_df_F3_MEANS_high$Count_dead), type = "right")

```

```{r F3 plot em}

pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_survival.pdf", width = 10)
ggsurvplot(survfit(surv_F3_all ~ Larvae_pCO2, binary_df_F3_all), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.2)), 
           break.x.by = 2 ,xlab = "Age (days post fertilization)",
           facet.by = "Parent_pCO2",
           legend.title = "treatment")
dev.off()

pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_survival.pdf", width = 10)
ggsurvplot(survfit(surv_F3_MEANS ~ Larvae_pCO2, binary_df_F3_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.2)), 
           break.x.by = 2 ,xlab = "Age (days post fertilization)",
           facet.by = "Parent_pCO2",
           legend.title = "treatment")
dev.off()
# 
# pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_surival_low_history.pdf")
 ggsurvplot(survfit(surv_F3_MEANS_low ~ Larvae_pCO2, binary_df_F3_MEANS_low), 
            risk.table = F, pval = F , conf.int = TRUE,
            palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.3)), 
            break.x.by = 2 ,xlab = "Age (days post fertilization)", 
            legend.title = "treatment")
# dev.off()
# 
# pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_moderate_history.pdf")
ggsurvplot(survfit(surv_F3_MEANS_mod ~ Larvae_pCO2, binary_df_F3_MEANS_mod), 
            risk.table = F, pval = F , conf.int = TRUE,
            palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.2)), 
            break.x.by = 2 ,xlab = "Age (days post fertilization)", 
            legend.title = "treatment")
# dev.off()
# 
# pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_high_history.pdf")
# ggsurvplot(survfit(surv_F3_MEANS_high ~ Larvae_pCO2, binary_df_F3_MEANS_high), 
#            risk.table = F, pval = F , conf.int = TRUE,
#            palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.3)), 
#            break.x.by = 2 ,xlab = "Age (days post fertilization)", 
#            legend.title = "treatment")
# dev.off()

```
run cox proportional test
include tank as a random effect
```{r}

# remove "none" and the level - for some reason this lingered, dont know why
str(binary_df_F3_all) # before running, youll see PArent_pCO2 has 4 levels - we only want 3!
is.na(binary_df_F3_all$Parent_pCO2) <- binary_df_F3_all$Parent_pCO2 == "None"
binary_df_F3_all$Parent_pCO2 <- factor(binary_df_F3_all$Parent_pCO2)
str(binary_df_F3_all) # youll see now its fixed

# now run the model

survFitME_all_tank <- coxme(surv_F3_all ~ Larvae_pCO2*Parent_pCO2 + (1|Tank), binary_df_F3_all)
summary(survFitME_all_tank) 
#                                                           coef exp(coef)   se(coef)     z
# Larvae_pCO2Moderate pCO2                          -0.025569607 0.9747545 0.01848915 -1.38
# Larvae_pCO2High pCO2                               0.009866940 1.0099158 0.01873860  0.53
# Parent_pCO2Moderate pCO2                          -0.037194870 0.9634884 0.01855713 -2.00
# Parent_pCO2High pCO2                              -0.076000227 0.9268160 0.01845043 -4.12
# Larvae_pCO2Moderate pCO2:Parent_pCO2Moderate pCO2  0.030348803 1.0308140 0.02614267  1.16
# Larvae_pCO2High pCO2:Parent_pCO2Moderate pCO2      0.008771872 1.0088105 0.02630283  0.33
# Larvae_pCO2Moderate pCO2:Parent_pCO2High pCO2      0.035694517 1.0363392 0.02605141  1.37
# Larvae_pCO2High pCO2:Parent_pCO2High pCO2          0.006232918 1.0062524 0.02611643  0.24
# p
#  1.7e-01
#      6.0e-01
#   4.5e-02
#     3.8e-05
#  2.5e-01
#    7.4e-01
#  1.7e-01
#   8.1e-01

#The standard deviation of the tank random effect is very small (0.015) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (527597), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 

#pairwise comparisons

library(multcomp)

glht(survFitME_all_tank, linfct = mcp(Parent_pCO2 = "Dunnett"),alternative = "greater")

# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0 -0.03719
# High pCO2 - Low pCO2 <= 0     -0.07600


```




# Stats
Lenght data do not meet the assumptions of noramlity even with an asin(sqrt()) transformation
```{r}
leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F1)
shapiro.test(df_F1$Percent_Survival_trans)

leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F2)
shapiro.test(df_F2$Percent_Survival_trans)

leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F3)
shapiro.test(df_F3$Percent_Survival_trans)
```

#Non parametric Stats
### F1 Survival
```{r}
scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Age,
                data = df_F1)

dunnTest(Percent_Survival ~ Larvae_pCO2,
                data = df_F1,
              method="bh")

```

### F2 Survival
```{r}
scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Age,
                data = df_F2)

res.kruskal <- kruskal.test(Percent_Survival ~ Larvae_pCO2, data = df_F2)
res.kruskal

```

### F3 Survival
```{r}


scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Parent_pCO2,
                data = df_F3)

dunnTest(Percent_Survival ~ Larvae_pCO2,
                data = df_F3,
              method="bh")

dunnTest(Percent_Survival ~ Parent_pCO2,
                data = df_F3,
              method="bh")
```