---
title: "Larval Growth and Survival"
author: "Katie McFarland"
date: "9/29/2023"
output:
  pdf_document: default
  html_document: default
---

# 20231217 - started manuscript Github repostiory 'EAD-ASEB-Airradians_Larvae_crossgen_OA'
# SGurr edits
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
# knitr::opts_knit$set(root.dir = "C:/Users/katherine.mcfarland/Documents/GitHub/EAD-ASEB-Airradians_multigen_OA/larvae") # Katie's
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis") # Sam's
```


```{r, echo=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(rcompanion)
library(FSA)
library(car)
library(forcats)
library(kableExtra) # nice Rmd tables
library(emmeans)
library(ggpubr)
library(survival)
library(Rmisc)
library(coxme)
library(survminer)
```


#Larval Survival
```{r, echo=FALSE}
df <- read.csv("Data/larval_survival.csv", header = T) %>% 
  dplyr::mutate(Percent_Survival = as.numeric(Percent_Survival)) %>% 
  dplyr::select(-treatment) %>%  # redundant column with Exposure_OA
  
  # rename the variables for consistency using 'case_when'
  dplyr::mutate(Parental_OA = case_when(Parental_OA ==  "LOW" ~ "Low pCO2", 
                                        Parental_OA ==  "MODERATE" ~ "Moderate pCO2", 
                                        Parental_OA ==  "HIGH" ~ "High pCO2",
                                        Parental_OA ==  "WILD" ~ "None")) %>% 
  dplyr::mutate(Exposure_OA = case_when(Exposure_OA ==  "LOW" ~ "Low pCO2", 
                                        Exposure_OA ==  "MODERATE" ~ "Moderate pCO2", 
                                        Exposure_OA ==  "HIGH" ~ "High pCO2")) %>% 
  
  # change the levels for order when plotting - call as factors too!
  dplyr::mutate(Parental_OA = as.factor(forcats::fct_relevel(Parental_OA, 
                                                             c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  dplyr::mutate(Exposure_OA = as.factor(forcats::fct_relevel(Exposure_OA, 
                                                             c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  
  # rename them
  dplyr::rename(Parent_pCO2 = Parental_OA) %>% 
  dplyr::rename(Larvae_pCO2  = Exposure_OA) # done 
  
  
df$Percent_Survival_trans <- asin(sqrt(df$Percent_Survival))

head(df)
str(df)
```

## About data 

* output tables for each generation where 'n' is the number of tanks!

```{r about data}

# number of runs 
unique(df$Generation) # 1, 2, and 3

Rep_Summary_all <- df %>% dplyr::select(c(Generation, Age, Parent_pCO2, Larvae_pCO2)) %>% 
                       dplyr::group_by_all() %>% 
                       dplyr::summarise(n=n())

# print the runs
# summary: 

Rep_Summary_G1 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 1) # 
Rep_Summary_G1 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_G2 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 2) #
Rep_Summary_G2 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_G3 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 3) # 
Rep_Summary_G3 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

```


### make data frames for each generation
```{r}
df_F1 <- df%>%
  filter(Generation=="1")

df_F2 <- df%>%
  filter(Generation=="2")

df_F3 <- df%>%
  filter(Generation=="3")
```

# F1 Larval Survival

## Plot it!
```{r F1 summary and plot}

stL_F1 <- summarySE(df_F1, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2"))
stL_F1 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")


F1_PlotMeanSE <-  ggplot(stL_F1, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                          geom_point()+ 
                           geom_errorbar(aes(ymin=Percent_Survival-se, 
                                             ymax=Percent_Survival+se), width=.2,
                                        position=position_dodge(.1)) +
                          scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                          theme(legend.position="right", legend.direction="vertical", legend.title = element_blank())+ theme_bw() + 
                          labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                          ggtitle("F1 larvae survival (Mean +- SE)") + 
                          theme(panel.grid.major = element_blank(), 
                                panel.grid.minor = element_blank(), 
                                panel.background = element_blank(), 
                                axis.line = element_line(colour = "black"))
 
F1_PlotMeanSE

pdf("Output/Survival/F1_survival_meanSE.pdf", width=10, height=5)
print(F1_PlotMeanSE)
dev.off()

```

## Build 'surv' object for statistical analysis 

* What the hell does this mean?? 
  
  - a 'surv' object reads ALL events as that of a dead or alive individual
  
  - but but but... we have volumetric count data of live and dead! Does this not throw a wrench into this type of analysis? 
  
    - Nope! Below we use logic to create a GIGANTIC table formatted correctly for this purpose!
  
```{r F1 build binary live and dead column}

#make a surv object to pass into coxme
# create new columns for live and dead based on the count and proportion (of alive) tobackcalculate dead - these are integers! 
df_F1 <- df_F1 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(Count_dead = case_when(proportion == 0 ~ 0,
                                               proportion >0 ~ (1/as.numeric(proportion))*as.numeric(Count_alive))) # proportion is that of alive


# Run two for loops 

# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting

# keep tank to use as a random effect in the cox model!
# Call the cumulative dataframe that we will write to in the for loop below
binary_df_F1_all     <- data.frame() # start dataframe  - all
binary_df_F1_mod     <- data.frame() # start dataframe  - only mod history
binary_df_F1_low     <- data.frame() # start dataframe  - only low history
# run it 
for (i in 1:nrow(df_F1)) {
   #count_alive <- round(df_F1[i,]$Count_alive/1000) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_F1[i,]$Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   count_alive <- round(df_F1[i,]$Count_alive) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F1[i,]$Count_dead)  # divide all counts by 1000, round to nearest
   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_F1[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_F1[i,]$Larvae_pCO2,
                            Generation  = df_F1[i,]$Generation,
                            Tank        = df_F1[i,]$Tank,  
                            Age         = df_F1[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF           <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F1_all <- rbind(binary_df_F1_all,loopDF) #bind to a cumulative list dataframe
  
}




# (2) Mean counts alive and dead by tank - better representated for plots!

# for loop for plotting!! 
# difference here is we take a mean for counts alive and dead within treatment, no tank included and a singluar line for plotting 
df_F1_MEANS <- df_F1 %>% 
              dplyr::select(-c(Generation, Tank, Date, Percent_Survival, proportion)) %>% 
              dplyr::group_by(Larvae_pCO2, Age) %>% 
              dplyr::summarise(mean_Count_alive = mean(Count_alive), 
                               mean_Count_dead = mean(Count_dead))
# call datafarme binary for plotting
binary_df_F1_MEANS       <- data.frame() # start dataframe 

# run it 
for (i in 1:nrow(df_F1_MEANS)) {
   count_alive <- round(df_F1_MEANS[i,]$mean_Count_alive/1000) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F1_MEANS[i,]$mean_Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   loopDF           <- data.frame(matrix(0,ncol = 3, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Larvae_pCO2','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Larvae_pCO2 = df_F1_MEANS[i,]$Larvae_pCO2,
                            Age         = df_F1_MEANS[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),3] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),3] = 0 # 0 for alive

  loopDF             <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F1_MEANS <- rbind(binary_df_F1_MEANS,loopDF) #bind to a cumulative list dataframe
  
}

# view the output binary format for building the surv object!
head(binary_df_F1_all) %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria") # view exmaple of the output data 


```

```{r F1 build surv object and plot em}

?Surv
# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F1_all  <- Surv(time = as.numeric(binary_df_F1_all$Age), 
                     event = as.numeric(binary_df_F1_all$Count_dead), type = "right")
# plot the surv data
ggsurvplot(survfit(surv_F1_all ~ Larvae_pCO2, binary_df_F1_all), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")




# (2) Mean counts alive and dead by tank - better presentative for plots!
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F1_means_all  <- Surv(time = as.numeric(binary_df_F1_MEANS$Age), 
                     event = as.numeric(binary_df_F1_MEANS$Count_dead), type = "right")
# with the mean data by  tank replicate! better - save this one
ggsurvplot(survfit(surv_F1_means_all ~ Larvae_pCO2, binary_df_F1_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")

# plot the surv data
pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F1_survival.pdf")
ggsurvplot(survfit(surv_F1_means_all ~ Larvae_pCO2, binary_df_F1_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")
dev.off()

```

```{r F1 run coxme on surv stats}
?coxme
head(binary_df_F1_all)
# use this data from the first fr loop - has live and dead counts with reoslution of treatment AND replicate tank!!!

# now run the model
F1_survFIT_cox <- coxme(surv_F1_all ~ Larvae_pCO2 + (1|Tank), binary_df_F1_all)
summary(F1_survFIT_cox)

#                               coef exp(coef)   se(coef)     z       p
# Larvae_pCO2Moderate pCO2 0.1059100  1.111722 0.01277808  8.29 1.1e-16
# Larvae_pCO2High pCO2     0.8663613  2.378241 0.01519810 57.00 0.0e+00


#The standard deviation of the tank random effect is very small (0.006017691) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (312321 ), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 


#pairwise comparisons

library(multcomp)
?glht
glht(F1_survFIT_cox, linfct = mcp(Larvae_pCO2 = "Dunnett"),alternative = "greater")
# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0   0.1059
# High pCO2 - Low pCO2 <= 0       0.8664


```


# F2 Larval Survival

## Plot it!
```{r F2 summary and plot}

stL_F2 <- summarySE(df_F2, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2"))
stL_F2 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

F2_PlotMeanSE <- ggplot(stL_F2, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2))+ geom_line()+
                            geom_point()+ 
                             geom_errorbar(aes(ymin=Percent_Survival-se, ymax=Percent_Survival+se), width=.2,
                                          position=position_dodge(.1)) +
                            scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                            theme(legend.position="right", legend.direction="vertical", legend.title = element_blank())+ theme_bw() + 
                            labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                            ggtitle("F2 larvae survival (Mean +- SE)") + 
                            theme(panel.grid.major = element_blank(), 
                                  panel.grid.minor = element_blank(), 
                                  panel.background = element_blank(), 
                                  axis.line = element_line(colour = "black"))
 
F2_PlotMeanSE

pdf("Output/Survival/F2_survival_meanSE.pdf", width=10, height=5)
print(F2_PlotMeanSE)
dev.off()

```

```{r F2 build binary live and dead column}

#make a surv object to pass into coxme
# create new columns for live and dead based on the count and proportion (of alive) tobackcalculate dead - these are integers! 
df_F2 <- df_F2 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(Count_dead = case_when(proportion == 0 ~ 0,
                                               proportion >0 ~ (1/as.numeric(proportion))*as.numeric(Count_alive))) # proportion is that of alive


# Run two for loops 

# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting

# keep tank to use as a random effect in the cox model!
# Call the cumulative dataframe that we will write to in the for loop below
binary_df_F2_all     <- data.frame() # start dataframe  - all

# run it 
for (i in 1:nrow(df_F2)) {
   count_alive <- round(df_F2[i,]$Count_alive/1000) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F2[i,]$Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_F2[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_F2[i,]$Larvae_pCO2,
                            Generation  = df_F2[i,]$Generation,
                            Tank        = df_F2[i,]$Tank,  
                            Age         = df_F2[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF           <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F2_all <- rbind(binary_df_F2_all,loopDF) #bind to a cumulative list dataframe
  
}




# (2) Mean counts alive and dead by tank - better presentative for plots!

# for loop for plotting!! 
# difference here is we take a mean for counts alive and dead within treatment, no tank included and a singluar line for plotting 
df_F2_MEANS <- df_F2 %>% 
              dplyr::select(-c(Generation, Tank, Date, Percent_Survival, proportion)) %>% 
              dplyr::group_by(Larvae_pCO2, Age) %>% 
              dplyr::summarise(mean_Count_alive = mean(Count_alive), 
                               mean_Count_dead = mean(Count_dead))
# call datafarme binary for plotting
binary_df_F2_MEANS       <- data.frame() # start dataframe 

# run it 
for (i in 1:nrow(df_F2_MEANS)) {
   count_alive <- round(df_F2_MEANS[i,]$mean_Count_alive/1000) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F2_MEANS[i,]$mean_Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   loopDF           <- data.frame(matrix(0,ncol = 3, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Larvae_pCO2','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Larvae_pCO2 = df_F2_MEANS[i,]$Larvae_pCO2,
                            Age         = df_F2_MEANS[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),3] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),3] = 0 # 0 for alive

  loopDF             <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F2_MEANS <- rbind(binary_df_F2_MEANS,loopDF) #bind to a cumulative list dataframe
  
}

# view the output binary format for building the surv object!
head(binary_df_F2_all) %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria") # view exmaple of the output data 


```

```{r F2 build surv object and plot em}


# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F2_all  <- Surv(time = as.numeric(binary_df_F2_all$Age), 
                     event = as.numeric(binary_df_F2_all$Count_dead), type = "right")
# plot the surv data
ggsurvplot(survfit(surv_F2_all ~ Larvae_pCO2, binary_df_F2_all), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")




# (2) Mean counts alive and dead by tank - better presentative for plots!
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F2_means_all  <- Surv(time = as.numeric(binary_df_F2_MEANS$Age), 
                     event = as.numeric(binary_df_F2_MEANS$Count_dead), type = "right")
# with the mean data by  tank replicate! better - save this one
ggsurvplot(survfit(surv_F2_means_all ~ Larvae_pCO2, binary_df_F2_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange"), c(1,0.8)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")

# plot the surv data
pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F2_survival.pdf")
ggsurvplot(survfit(surv_F2_means_all ~ Larvae_pCO2, binary_df_F2_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")
dev.off()

```

```{r F2 run coxme on surv stats}

head(binary_df_F2_all)
# use this data from the first fr loop - has live and dead counts with reoslution of treatment AND replicate tank!!!

# omit the level high pCO2 because it is not actually present - to be a carry over from the full dataset?
is.na(binary_df_F2_all$Larvae_pCO2) <- binary_df_F2_all$Larvae_pCO2 == "High pCO2"
binary_df_F2_all$Larvae_pCO2 <- factor(binary_df_F2_all$Larvae_pCO2)
str(binary_df_F2_all) # Larvae_pCO2 now has two factors - correct!!

# now run the model
F2_survFIT_cox <- coxme(surv_F2_all ~ Larvae_pCO2 + (1|Tank), binary_df_F2_all)
summary(F2_survFIT_cox)

#                               coef exp(coef)   se(coef)   z      p
# Larvae_pCO2Moderate pCO2 0.1269179  1.135324 0.04536691 2.8 0.0051


#The standard deviation of the tank random effect is very small (0.006017691) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (312321 ), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 


#pairwise comparisons

library(multcomp)

glht(F2_survFIT_cox, linfct = mcp(Larvae_pCO2 = "Dunnett"),alternative = "greater")

# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0   0.1269


```

# F3 Larval Survival
## Plot it!
```{r F3 summary and plot}

stL_F3 <- summarySE(df_F3, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2", "Parent_pCO2"))
stL_F3 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Gen3_PlotMeanSE <- ggplot(stL_F3, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2))+ 
                        geom_line()+
                        geom_point()+ 
                        geom_errorbar(aes(ymin=Percent_Survival-se, ymax=Percent_Survival+se), width=.2,
                                      position=position_dodge(.1)) +
                        scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                        theme(legend.position="right", legend.direction="vertical", 
                              legend.title = element_blank())+ theme_bw() + 
                        labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                        ggtitle("F3 larvae survival (Mean +- SE)") + 
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              panel.background = element_blank(), 
                              axis.line = element_line(colour = "black"))+
                        facet_wrap("Parent_pCO2")
Gen3_PlotMeanSE


# Summarise Percent Deformities for plotting 
Gen3_mean_survival <- df_F3 %>% 
                          dplyr::select(Parent_pCO2,Larvae_pCO2,Age, Percent_Survival) %>% 
                          na.omit() %>% 
                          dplyr::group_by(Age, Parent_pCO2,Larvae_pCO2) %>% 
                          dplyr::summarise(mean_Perc_surv = mean(Percent_Survival), 
                                           n           = n(),
                                           sd_Perc_surv   = sd(Percent_Survival),
                                           se_Perc_surv   = sd_Perc_surv/(sqrt(n)))

# Barplots 
unique(Gen3_mean_survival$Age)
Gen3_barplots_Age2 <- Gen3_mean_survival %>% dplyr::filter(Age ==2) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 2 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)
Gen3_barplots_Age5 <- Gen3_mean_survival %>% dplyr::filter(Age ==5) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 5 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

Gen3_barplots_Age7 <- Gen3_mean_survival %>% dplyr::filter(Age ==7) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 7 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

Gen3_barplots_Age12 <- Gen3_mean_survival %>% dplyr::filter(Age ==12) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 12 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)
Gen3_barplots_Age16 <- Gen3_mean_survival %>% dplyr::filter(Age ==16) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 16 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

# print the plot

pdf("Output/Survival/F3_survival_meanSE.pdf", width=10, height=5)
print(Gen3_PlotMeanSE_All)
dev.off()
# 
# 
# pdf("Output/Survival/PercSurvival_Gen3_2DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age2)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_5DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age5)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_7DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age7)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_12DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age12)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_16DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age16)
# dev.off()

```

```{r F3 build binary live and dead column}
#make a surv object to pass into coxme
df_F3 <- df_F3 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(Count_dead = 
                          (1/as.numeric(proportion))*as.numeric(Count_alive)) # proportion is that of alive


# Run two for loops 

# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting

# keep tank to use as a random effect in the cox model!
# Call the cumulative dataframe that we will write to in the for loop below
binary_df_F3_all     <- data.frame() # start dataframe 
binary_df_F3_high    <- data.frame() # start dataframe 
binary_df_F3_mod     <- data.frame() # start dataframe 
binary_df_F3_low     <- data.frame() # start dataframe 
# run it 
for (i in 1:nrow(df_F3)) {
   #count_alive <- round(df_F3[i,]$Count_alive/1000) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_F3[i,]$Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   count_alive <- round(df_F3[i,]$Count_alive) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F3[i,]$Count_dead)  # divide all counts by 1000, round to nearest

   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_F3[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_F3[i,]$Larvae_pCO2,
                            Generation  = df_F3[i,]$Generation,
                            Tank        = df_F3[i,]$Tank,  
                            Age         = df_F3[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF       <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F3_all <- rbind(binary_df_F3_all,loopDF) #bind to a cumulative list dataframe
  # print(binary_df_F3) # print to monitor progress

     
  if (loopDF$Parent_pCO2[1] == 'High pCO2') {
      binary_df_F3_high <- rbind(binary_df_F3_high,loopDF) #bind to a cumulative list dataframe
  } else if (loopDF$Parent_pCO2[1] == 'Moderate pCO2') {
      binary_df_F3_mod <- rbind(binary_df_F3_mod,loopDF) #bind to a cumulative list dataframe
  } else {
      binary_df_F3_low <- rbind(binary_df_F3_low,loopDF) #bind to a cumulative list dataframe
    }
      
  
}

binary_df_F3_all$All_pCO2 <- as.factor(paste0(binary_df_F3_all$Parent_pCO2, ' x ', binary_df_F3_all$Larvae_pCO2))


# (2) Mean counts alive and dead by tank - better presentative for plots!

# for loop for plotting!! 
# difference here is we take a mean for counts alive and dead within treatment, no tank included and a singluar line for plotting 
df_F3_MEANS <- df_F3 %>% 
  dplyr::select(-c(Generation, Tank, Date, Percent_Survival, proportion)) %>% 
  dplyr::group_by(Parent_pCO2, Larvae_pCO2, Age) %>% 
  dplyr::summarise(mean_Count_alive = mean(Count_alive), 
            mean_Count_dead = mean(Count_dead))
# call datafarme binary for plotting
binary_df_F3_MEANS       <- data.frame() # start dataframe 
binary_df_F3_MEANS_high  <- data.frame() # start dataframe 
binary_df_F3_MEANS_mod   <- data.frame() # start dataframe 
binary_df_F3_MEANS_low   <- data.frame() # start dataframe 

# run it 
for (i in 1:nrow(df_F3_MEANS)) {
   #count_alive <- round(df_F3_MEANS[i,]$mean_Count_alive/1000) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_F3_MEANS[i,]$mean_Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   count_alive <- round(df_F3_MEANS[i,]$mean_Count_alive) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F3_MEANS[i,]$mean_Count_dead)  # divide all counts by 1000, round to nearest

   
   loopDF           <- data.frame(matrix(0,ncol = 4, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_F3_MEANS[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_F3_MEANS[i,]$Larvae_pCO2,
                            Age         = df_F3_MEANS[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),4] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),4] = 0 # 0 for alive

  loopDF             <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F3_MEANS <- rbind(binary_df_F3_MEANS,loopDF) #bind to a cumulative list dataframe
     
  if (loopDF$Parent_pCO2[1] == 'High pCO2') {
      binary_df_F3_MEANS_high <- rbind(binary_df_F3_MEANS_high,loopDF) #bind to a cumulative list dataframe
  } else if (loopDF$Parent_pCO2[1] == 'Moderate pCO2') {
      binary_df_F3_MEANS_mod <- rbind(binary_df_F3_MEANS_mod,loopDF) #bind to a cumulative list dataframe
  } else {
      binary_df_F3_MEANS_low <- rbind(binary_df_F3_MEANS_low,loopDF) #bind to a cumulative list dataframe
  }
    
}
```

```{r F3 build surv object}
nrow(binary_df_F3_all)
# create surv object
# with tank included in the data frame (first for loop above)
surv_F3_all  <- Surv(time = as.numeric(binary_df_F3_all$Age), 
                     event = as.numeric(binary_df_F3_all$Count_dead), type = "right")

surv_F3_high <- Surv(time = as.numeric(binary_df_F3_high$Age), 
                     event = as.numeric(binary_df_F3_high$Count_dead), type = "right")

surv_F3_mod  <- Surv(time = as.numeric(binary_df_F3_mod$Age), 
                     event = as.numeric(binary_df_F3_mod$Count_dead), type = "right")

surv_F3_low  <- Surv(time = as.numeric(binary_df_F3_low$Age), 
                     event = as.numeric(binary_df_F3_low$Count_dead), type = "right")


F3_low_survplot <- ggsurvplot(survfit(surv_F3_low ~ Larvae_pCO2, binary_df_F3_low), 
                       risk.table = F, pval = F , conf.int = TRUE,
                       palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8, 0.3)), 
                       ggtheme = theme_survminer(),
                       break.x.by = 2 ,xlab = "Age (days post fertilization)", 
                       legend.title = "treatment")

F3_mod_survplot <- ggsurvplot(survfit(surv_F3_mod ~ Larvae_pCO2, binary_df_F3_mod), 
                       risk.table = F, pval = F , conf.int = TRUE,
                       palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8, 0.3)), 
                       ggtheme = theme_survminer(),
                       break.x.by = 2 ,xlab = "Age (days post fertilization)", 
                       legend.title = "treatment")

F3_high_survplot <- ggsurvplot(survfit(surv_F3_high ~ Larvae_pCO2, binary_df_F3_high), 
                       risk.table = F, pval = F , conf.int = TRUE,
                       palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8, 0.3)), 
                       ggtheme = theme_survminer(),
                       break.x.by = 2 ,xlab = "Age (days post fertilization)", 
                       legend.title = "treatment")

F3_full_plot <- ggarrange(F3_low_survplot, F3_mod_survplot, F3_high_survplot, ncol = 3)
print(F3_full_plot)


# create surv object
# without tank included - means wihtin treatment SEPCIFICALLY FOR PLOTTING  second for loop above 
surv_F3_MEANS  <- Surv(time = as.numeric(binary_df_F3_MEANS$Age), 
                     event = as.numeric(binary_df_F3_MEANS$Count_dead), type = "right")


is.na(binary_df_F3_MEANS_low$Parent_pCO2) <- binary_df_F3_MEANS_low$Parent_pCO2 == "None"
binary_df_F3_MEANS_low$Parent_pCO2 <- factor(binary_df_F3_MEANS_low$Parent_pCO2)
str(binary_df_F3_MEANS_low) # youll see now its fixed

surv_F3_MEANS_low  <- Surv(time = as.numeric(binary_df_F3_MEANS_low$Age), 
                     event = as.numeric(binary_df_F3_MEANS_low$Count_dead), type = "right")

is.na(binary_df_F3_MEANS_mod$Parent_pCO2) <- binary_df_F3_MEANS_mod$Parent_pCO2 == "None"
binary_df_F3_MEANS_mod$Parent_pCO2 <- factor(binary_df_F3_MEANS_mod$Parent_pCO2)
str(binary_df_F3_MEANS_mod) # youll see now its fixed

surv_F3_MEANS_mod  <- Surv(time = as.numeric(binary_df_F3_MEANS_mod$Age), 
                     event = as.numeric(binary_df_F3_MEANS_mod$Count_dead), type = "right")


is.na(binary_df_F3_MEANS_high$Parent_pCO2) <- binary_df_F3_MEANS_high$Parent_pCO2 == "None"
binary_df_F3_MEANS_high$Parent_pCO2 <- factor(binary_df_F3_MEANS_high$Parent_pCO2)
str(binary_df_F3_MEANS_high) # youll see now its fixed

surv_F3_MEANS_high  <- Surv(time = as.numeric(binary_df_F3_MEANS_high$Age), 
                     event = as.numeric(binary_df_F3_MEANS_high$Count_dead), type = "right")

```

```{r F3 plot em}

pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_survival.pdf", width = 10)
ggsurvplot(survfit(surv_F3_all ~ Larvae_pCO2, binary_df_F3_all), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.2)), 
           break.x.by = 2 ,xlab = "Age (days post fertilization)",
           facet.by = "Parent_pCO2",
           legend.title = "treatment")
dev.off()

pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_survival.pdf", width = 10)
ggsurvplot(survfit(surv_F3_MEANS ~ Larvae_pCO2, binary_df_F3_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.2)), 
           break.x.by = 2 ,xlab = "Age (days post fertilization)",
           facet.by = "Parent_pCO2",
           legend.title = "treatment")
dev.off()
# 
# pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_surival_low_history.pdf")
 ggsurvplot(survfit(surv_F3_MEANS_low ~ Larvae_pCO2, binary_df_F3_MEANS_low), 
            risk.table = F, pval = F , conf.int = TRUE,
            palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.3)), 
            break.x.by = 2 ,xlab = "Age (days post fertilization)", 
            legend.title = "treatment")
# dev.off()
# 
# pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_moderate_history.pdf")
ggsurvplot(survfit(surv_F3_MEANS_mod ~ Larvae_pCO2, binary_df_F3_MEANS_mod), 
            risk.table = F, pval = F , conf.int = TRUE,
            palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.2)), 
            break.x.by = 2 ,xlab = "Age (days post fertilization)", 
            legend.title = "treatment")
# dev.off()
# 
# pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_high_history.pdf")
# ggsurvplot(survfit(surv_F3_MEANS_high ~ Larvae_pCO2, binary_df_F3_MEANS_high), 
#            risk.table = F, pval = F , conf.int = TRUE,
#            palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.3)), 
#            break.x.by = 2 ,xlab = "Age (days post fertilization)", 
#            legend.title = "treatment")
# dev.off()

```
run cox proportional test
include tank as a random effect
```{r}

# remove "none" and the level - for some reason this lingered, dont know why
str(binary_df_F3_all) # before running, youll see PArent_pCO2 has 4 levels - we only want 3!
is.na(binary_df_F3_all$Parent_pCO2) <- binary_df_F3_all$Parent_pCO2 == "None"
binary_df_F3_all$Parent_pCO2 <- factor(binary_df_F3_all$Parent_pCO2)
str(binary_df_F3_all) # youll see now its fixed

# now run the model

survFitME_all_tank <- coxme(surv_F3_all ~ Larvae_pCO2*Parent_pCO2 + (1|Tank), binary_df_F3_all)
summary(survFitME_all_tank) 
#                                                           coef exp(coef)   se(coef)     z
# Larvae_pCO2Moderate pCO2                          -0.025569607 0.9747545 0.01848915 -1.38
# Larvae_pCO2High pCO2                               0.009866940 1.0099158 0.01873860  0.53
# Parent_pCO2Moderate pCO2                          -0.037194870 0.9634884 0.01855713 -2.00
# Parent_pCO2High pCO2                              -0.076000227 0.9268160 0.01845043 -4.12
# Larvae_pCO2Moderate pCO2:Parent_pCO2Moderate pCO2  0.030348803 1.0308140 0.02614267  1.16
# Larvae_pCO2High pCO2:Parent_pCO2Moderate pCO2      0.008771872 1.0088105 0.02630283  0.33
# Larvae_pCO2Moderate pCO2:Parent_pCO2High pCO2      0.035694517 1.0363392 0.02605141  1.37
# Larvae_pCO2High pCO2:Parent_pCO2High pCO2          0.006232918 1.0062524 0.02611643  0.24
# p
#  1.7e-01
#      6.0e-01
#   4.5e-02
#     3.8e-05
#  2.5e-01
#    7.4e-01
#  1.7e-01
#   8.1e-01

#The standard deviation of the tank random effect is very small (0.015) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (527597), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 

#pairwise comparisons

library(multcomp)

glht(survFitME_all_tank, linfct = mcp(Parent_pCO2 = "Dunnett"),alternative = "greater")

# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0 -0.03719
# High pCO2 - Low pCO2 <= 0     -0.07600


```




# Stats
Lenght data do not meet the assumptions of noramlity even with an asin(sqrt()) transformation
```{r}
leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F1)
shapiro.test(df_F1$Percent_Survival_trans)

leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F2)
shapiro.test(df_F2$Percent_Survival_trans)

leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F3)
shapiro.test(df_F3$Percent_Survival_trans)
```

#Non parametric Stats
### F1 Survival
```{r}
scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Age,
                data = df_F1)

dunnTest(Percent_Survival ~ Larvae_pCO2,
                data = df_F1,
              method="bh")

```

### F2 Survival
```{r}
scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Age,
                data = df_F2)

res.kruskal <- kruskal.test(Percent_Survival ~ Larvae_pCO2, data = df_F2)
res.kruskal

```

### F3 Survival
```{r}


scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Parent_pCO2,
                data = df_F3)

dunnTest(Percent_Survival ~ Larvae_pCO2,
                data = df_F3,
              method="bh")

dunnTest(Percent_Survival ~ Parent_pCO2,
                data = df_F3,
              method="bh")
```