---
title: "Larval Growth and Survival"
author: "Katie McFarland"
date: "9/29/2023"
output:
  pdf_document: default
  html_document: default
---

### SET UP
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
# knitr::opts_knit$set(root.dir = "C:/Users/katherine.mcfarland/Documents/GitHub/EAD-ASEB-Airradians_multigen_OA/larvae") # Katie's
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis") # Sam's
```

### LOAD PACKAGES
```{r, echo=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(rcompanion)
library(FSA)
library(car)
library(forcats)
library(kableExtra) # nice Rmd tables
library(emmeans)
library(ggpubr)
library(survival)
library(Rmisc)
library(coxme)
library(survminer)
library(ggsurvfit) # survfit2
library(gtsummary) # tbl_survfit
```

### LAOD DATA
```{r, echo=FALSE}
df <- read.csv("Data/larval_survival.csv", header = T) %>% 
  dplyr::mutate(Percent_Survival = as.numeric(Percent_Survival)) %>% 
  # dplyr::select(-treatment) %>%  # redundant column with Exposure_OA
  
  # rename the variables for consistency using 'case_when'
  dplyr::mutate(Parental_OA = case_when(Parental_OA ==  "LOW" ~ "Low pCO2", 
                                        Parental_OA ==  "MODERATE" ~ "Moderate pCO2", 
                                        Parental_OA ==  "HIGH" ~ "High pCO2",
                                        Parental_OA ==  "WILD" ~ "None")) %>% 
  dplyr::mutate(Exposure_OA = case_when(Exposure_OA ==  "LOW" ~ "Low pCO2", 
                                        Exposure_OA ==  "MODERATE" ~ "Moderate pCO2", 
                                        Exposure_OA ==  "HIGH" ~ "High pCO2")) %>% 
  
  # change the levels for order when plotting - call as factors too!
  dplyr::mutate(Parental_OA = as.factor(forcats::fct_relevel(Parental_OA, 
                                                             c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  dplyr::mutate(Exposure_OA = as.factor(forcats::fct_relevel(Exposure_OA, 
                                                             c("Low pCO2", "Moderate pCO2", "High pCO2")))) %>% 
  
  # rename them
  dplyr::rename(Parent_pCO2 = Parental_OA) %>% 
  dplyr::rename(Larvae_pCO2  = Exposure_OA) # done 
  
  
df$Percent_Survival_trans <- asin(sqrt(df$Percent_Survival))

# head(df)
# str(df)
```

# About data 

### Experiments: 

  - (2) F1 larvae under three OA levels, *grown out to adult stage*

  - (4) F2 larvae under two OA levels, *grown out to adult stage*

  - (6) F3 larvae under 3 x 3 parent x offsring OA, *no grow out past juvenile takedown of the experiment*

### * NOTE: N = 3 for Runs 3 and 6 for parental x offspring pCO2 treamtents 

```{r about data}

# number of runs 
unique(df$Generation) # 1, 2, and 3

Rep_Summary_all <- df %>% dplyr::select(c(Generation, Age, Parent_pCO2, Larvae_pCO2)) %>% 
                       dplyr::group_by_all() %>% 
                       dplyr::summarise(n=n())

# print the runs
# summary: 

Rep_Summary_G1 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 1) # 
Rep_Summary_G1 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_G2 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 2) #
Rep_Summary_G2 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Rep_Summary_G3 <- Rep_Summary_all %>%  dplyr::filter(Generation %in% 3) # 
Rep_Summary_G3 %>%  kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

```


### Build data frames for each expeirment 


```{r}
df_Exp2 <- df%>%
  filter(Generation=="1") %>% dplyr::mutate(Experiment = 2)

df_Exp4 <- df%>%
  filter(Generation=="2") %>% dplyr::mutate(Experiment = 4)

df_Exp6 <- df%>%
  filter(Generation=="3") %>% dplyr::mutate(Experiment = 6)
```


# Plot data as barplots mean +- SE for each run

## Manuscript 1; Expeirments 2 and 4

**output folder** = '1_Survival_Growth'

* note that metamorphosis was observed at 10 dpf in F1s and 14 dpf in F2s, therfore 
larval survival comapred across gnerations should only show < 10 dpf survival 
 
```{r Mean table + plots; F1 Experiment #2}

stL_Exp2 <- summarySE(df_Exp2, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2"))
stL_Exp2 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Exp2_PlotMeanSE <-  ggplot(stL_Exp2, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                          geom_point()+ 
                           geom_errorbar(aes(ymin=Percent_Survival-se, 
                                             ymax=Percent_Survival+se), width=.2,
                                        position=position_dodge(.1)) +
                          scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                          theme(legend.position="right", legend.direction="vertical", 
                                legend.title = element_blank())+ theme_bw() + 
                          labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                          ggtitle("F1 larvae + post-metamorphosis survival (Mean +- SE)") + 
                          theme(panel.grid.major = element_blank(), 
                                panel.grid.minor = element_blank(), 
                                panel.background = element_blank(), 
                                axis.line = element_line(colour = "black"))
Exp2_PlotMeanSE

Exp2_larvae_PlotMeanSE <-  stL_Exp2 %>%  dplyr::filter(Age <= 10) %>% 
                                ggplot(aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                                        geom_point()+ 
                                         geom_errorbar(aes(ymin=Percent_Survival-se, 
                                                           ymax=Percent_Survival+se), width=.2,
                                                      position=position_dodge(.1)) +
                                        scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                                        theme(legend.position="right", legend.direction="vertical", 
                                              legend.title = element_blank())+ theme_bw() + 
                                        labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                                        ggtitle("F1 larvae survival (Mean +- SE)") + 
                                        theme(panel.grid.major = element_blank(), 
                                              panel.grid.minor = element_blank(), 
                                              panel.background = element_blank(), 
                                              axis.line = element_line(colour = "black"))
 
Exp2_larvae_PlotMeanSE

# output the data !!

write.csv(stL_Exp2,"Output/1_Survival_Growth/Experiment1_Mean_Survival.csv")

pdf("Output/1_Survival_Growth/Experiment1_F1larvae_Survival_MeanSE.pdf", width=10, height=5)
print(Exp2_larvae_PlotMeanSE)
dev.off()

pdf("Output/1_Survival_Growth/Experiment1_F1_Survival_MeanSE.pdf", width=10, height=5)
print(Exp2_PlotMeanSE)
dev.off()
```

```{r Mean table + plots; F2 Experiment #4}

stL_Exp4 <- summarySE(df_Exp4, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2"))
stL_Exp4 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Exp4_PlotMeanSE <-  ggplot(stL_Exp4, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                          geom_point()+ 
                           geom_errorbar(aes(ymin=Percent_Survival-se, 
                                             ymax=Percent_Survival+se), width=.2,
                                        position=position_dodge(.1)) +
                          scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                          theme(legend.position="right", legend.direction="vertical", 
                                legend.title = element_blank())+ theme_bw() + 
                          labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                          ggtitle("F2 larvae + post-metamorphosis survival (Mean +- SE)") + 
                          theme(panel.grid.major = element_blank(), 
                                panel.grid.minor = element_blank(), 
                                panel.background = element_blank(), 
                                axis.line = element_line(colour = "black"))
Exp4_PlotMeanSE

Exp4_larvae_PlotMeanSE <-  stL_Exp4 %>%  dplyr::filter(Age <= 14) %>% 
                                ggplot(aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                                        geom_point()+ 
                                         geom_errorbar(aes(ymin=Percent_Survival-se, 
                                                           ymax=Percent_Survival+se), width=.2,
                                                      position=position_dodge(.1)) +
                                        scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                                        theme(legend.position="right", legend.direction="vertical", 
                                              legend.title = element_blank())+ theme_bw() + 
                                        labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                                        ggtitle("F2 larvae survival (Mean +- SE)") + 
                                        theme(panel.grid.major = element_blank(), 
                                              panel.grid.minor = element_blank(), 
                                              panel.background = element_blank(), 
                                              axis.line = element_line(colour = "black"))
 
Exp4_larvae_PlotMeanSE

# output the data !!

write.csv(stL_Exp4,"Output/1_Survival_Growth/Experiment4_Mean_Survival.csv")

pdf("Output/1_Survival_Growth/Experiment4_F2larvae_Survival_MeanSE.pdf", width=10, height=5)
print(Exp4_larvae_PlotMeanSE)
dev.off()

pdf("Output/1_Survival_Growth/Experiment4_F2_Survival_MeanSE.pdf", width=10, height=5)
print(Exp4_PlotMeanSE)
dev.off()
```

## Manuscript 4; Expeirments 6

**output folder** = '4_F3_Full_factorial'

```{r Mean table + plots; F3 Experiment #6}

stL_Exp6 <- summarySE(df_Exp6, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2"))
stL_Exp6 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Exp6_PlotMeanSE <-  ggplot(stL_Exp6, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                          geom_point()+ 
                           geom_errorbar(aes(ymin=Percent_Survival-se, 
                                             ymax=Percent_Survival+se), width=.2,
                                        position=position_dodge(.1)) +
                          scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                          theme(legend.position="right", legend.direction="vertical", 
                                legend.title = element_blank())+ theme_bw() + 
                          labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                          ggtitle("F3 larvae + post-metamorphosis survival (Mean +- SE)") + 
                          theme(panel.grid.major = element_blank(), 
                                panel.grid.minor = element_blank(), 
                                panel.background = element_blank(), 
                                axis.line = element_line(colour = "black"))
Exp6_PlotMeanSE

Exp6_larvae_PlotMeanSE <-  stL_Exp6 %>%  dplyr::filter(Age <= 14) %>% 
                                ggplot(aes(x=Age, y=Percent_Survival, color=Larvae_pCO2)) + geom_line()+
                                        geom_point()+ 
                                         geom_errorbar(aes(ymin=Percent_Survival-se, 
                                                           ymax=Percent_Survival+se), width=.2,
                                                      position=position_dodge(.1)) +
                                        scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                                        theme(legend.position="right", legend.direction="vertical", 
                                              legend.title = element_blank())+ theme_bw() + 
                                        labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                                        ggtitle("F3 larvae survival (Mean +- SE)") + 
                                        theme(panel.grid.major = element_blank(), 
                                              panel.grid.minor = element_blank(), 
                                              panel.background = element_blank(), 
                                              axis.line = element_line(colour = "black"))
 
Exp6_larvae_PlotMeanSE

# output the data !!

write.csv(stL_Exp6,"Output/4_F3_Full_factorial/Experiment6_Mean_Survival.csv")

pdf("Output/4_F3_Full_factorial/Experiment6_F2larvae_Survival_MeanSE.pdf", width=10, height=5)
print(Exp6_larvae_PlotMeanSE)
dev.off()

pdf("Output/4_F3_Full_factorial/Experiment6_F2_Survival_MeanSE.pdf", width=10, height=5)
print(Exp6_PlotMeanSE)
dev.off()
```

# Kaplan-Meier plots


## About:

* We will use the {ggsurvfit} package to generate Kaplan-Meier plots. This package aims to ease plotting of time-to-event endpoints using the power of the {ggplot2} package. See http://www.danieldsjoberg.com/ggsurvfit/index.html for details.

## Build 'surv' object for statistical analysis 

* useful links [https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html#Comparing_survival_times_between_groups]

* What the hell does a surv object mean?? 
  
  - a 'surv' object reads ALL events as that of a dead (1) or alive (0) individual
  
  - but but but... we have volumetric count data of live and dead! Does this not throw a wrench into this type of analysis? 
  
    - Nope! Below we use logic to create a GIGANTIC table formatted correctly for this purpose!

## Format

 * Build new column for count dead - take the count alive data from time 0 and porp survival to calcualted number dead

## Convert to binary formant (super long data tbale!) using two for loops:

 * (1) Including tank - to use as surv object in coxme statistics (random factor included!) not suitable for plotting
 
 * (2) Mean counts alive and dead by tank - better represented for plots!






## Manuscript 1; Expeirments 2 and 4

**output folder** = '1_Survival_Growth'

**note**: out ultimate objsetice with the sruvival data in this manuscript is to investigate change insurvivroship in reponse to treamtent across treatment 
and generational timescales - we reached metamorphosis at 10 DPF for the Experiment 2 and at 14 DPF or experiment 4, so we will need 
to cater to this difference to make meaningful comparisons





### Experiment 2 F1s




```{r format F1 Experiment #2}
# create new columns for live and dead based on the count and proportion (of alive) tobackcalculate dead - these are integers! 
df_Exp2 <- df_Exp2 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(Count_dead = case_when(proportion == 0 ~ 0,
                                               proportion >0 ~
                                                 (1/as.numeric(proportion))*
                                                 as.numeric(Count_alive))) # proportion is that of alive
```

```{r convert to binary F1 Experiment #2}

# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not suitable for plotting

# Call the cumulative dataframe that we will write to in the for loop below
binary_Exp2_all     <- data.frame() # start dataframe  - all
# run it 
for (i in 1:nrow(df_Exp2)) {
   count_alive <- round(df_Exp2[i,]$Count_alive/1000) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_Exp2[i,]$Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   #count_alive <- round(df_Exp2[i,]$Count_alive) # raw counts
   #count_dead  <- round(df_Exp2[i,]$Count_dead)  # raw counts
   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_Exp2[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_Exp2[i,]$Larvae_pCO2,
                            Generation  = df_Exp2[i,]$Generation,
                            Tank        = df_Exp2[i,]$Tank,  
                            Age         = df_Exp2[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF           <- data.frame(loopDF) # name dataframe for this single row
  binary_Exp2_all <- rbind(binary_Exp2_all,loopDF) #bind to a cumulative list dataframe
}
# To address tank as a random factor it MUST be numeric, change this here
unique(binary_Exp2_all$Tank) # "A" "B" "C" "D" "E" NA
binary_Exp2_all <- binary_Exp2_all %>% 
                    dplyr::mutate(Generation = 1,
                                  Tank = case_when(Tank== 'A'~1,
                                                   Tank== 'B'~2,
                                                   Tank== 'C'~3,
                                                   Tank== 'D'~4,
                                                   Tank== 'E'~5)) %>% 
                    na.omit()


# (2) Mean counts alive and dead by tank - better representated for plots!
# difference here is we take a mean for counts alive and dead within treatment, no tank included and a singluar line for plotting 
df_Exp2_MEANS <- df_Exp2 %>% 
              dplyr::select(-c(Generation, Tank, Date, Percent_Survival, proportion)) %>% 
              dplyr::group_by(Larvae_pCO2, Age) %>% 
              dplyr::summarise(mean_Count_alive = mean(Count_alive), 
                               mean_Count_dead = mean(Count_dead))

# call datafarme binary for plotting
binary_Exp2_MEANS       <- data.frame() # start dataframe 

# run it 
for (i in 1:nrow(df_Exp2_MEANS)) {
   count_alive <- round(df_Exp2_MEANS[i,]$mean_Count_alive/1000) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_Exp2_MEANS[i,]$mean_Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   loopDF           <- data.frame(matrix(0,ncol = 3, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Larvae_pCO2','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Larvae_pCO2 = df_Exp2_MEANS[i,]$Larvae_pCO2,
                            Age         = df_Exp2_MEANS[i,]$Age
                          ) 
  loopDF[c(1:count_dead),3] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),3] = 0 # 0 for alive

  loopDF             <- data.frame(loopDF) # name dataframe for this single row
  binary_Exp2_MEANS <- rbind(binary_Exp2_MEANS,loopDF) #bind to a cumulative list dataframe
  
}
binary_Exp2_MEANS <- binary_Exp2_MEANS %>% 
                      dplyr::mutate(Generation = 1) %>% 
                      na.omit()

```

**note**: larval period ended at first sign of metamorphossi at 10 DPF  - truncate before this age to compare
```{r filter larval period F1 Experiment #2}

binary_Exp2_larvae      <- binary_Exp2_all %>% filter(Age <=10) # F1 - aligned timepoint prior to F1 metamorph for Gen 1 and 2 comparisons

binary_Exp2_larvaeMEANS <- binary_Exp2_MEANS %>% filter(Age <=10) # F1 - aligned timepoint prior to F1 metamorph for Gen 1 and 2 comparisons

```

```{r surv object F1 Experiment #2}

survExp2_all <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp2_larvaeMEANS)


# convert Age to the proportion of time to metamorphosis 
binary_Exp2_larvaeMEANS$Prop_time_metamophosis <- (binary_Exp2_larvaeMEANS$Age / 10)



survExp2          <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp2_larvaeMEANS)
survExp2_proptime <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Larvae_pCO2, data = binary_Exp2_larvaeMEANS)


```

```{r plot Kaplan Meier F1 Experiment #2}

Exp2_larvaeKaplanMeier <- survExp2 %>% 
                        ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                        scale_color_manual(values = c("forestgreen","orange","purple")) +
                        scale_fill_manual(values = c("forestgreen","orange","purple")) +
                        labs(
                          x = "Days",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:10)) #+
                       # add_pvalue(location  = "annotation", x = 8.5)

Exp2_larvaeKaplanMeier <- survExp2_proptime %>% 
                        ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                        scale_color_manual(values = c("forestgreen","orange","purple")) +
                        scale_fill_manual(values = c("forestgreen","orange","purple")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                       # add_pvalue(location  = "annotation", x = 8.5)

# plot the surv data
pdf("Output/1_Survival_Growth/Expeirment2_Survival_KaplanMeier.pdf", width=5, height = 4)
Exp2_larvaeKaplanMeier
dev.off()

Exp_10d_surv_prob <- survExp2 %>% 
                          tbl_survfit(
                            times = 10,
                            label_header = "**survival to 10 days post-fertilization (95% CI)**"
                          )


```

# Comparing survival times between groups (log-rank test)
* We can conduct between-group significance tests using a log-rank test. The log-rank test equally weights observations over the entire follow-up time and is the most common way to compare survival times between groups. There are versions that more heavily weight the early or late follow-up that could be more appropriate depending on the research question (see ?survdiff for different test options).
* We get the log-rank p-value using the survdiff function
```{r log-rank test survdiff F1 Experiment #2}

Exp2_logrank_test <- survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp2_larvaeMEANS)
#                              N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2      1340     1154     1216      3.14      10.6
# Larvae_pCO2=Moderate pCO2 1084      946     1066     13.46      42.5
# Larvae_pCO2=High pCO2      805      762      580     56.82     126.1

```

# The Cox regression model
* We may want to quantify an effect size for a single variable, or include **more than one variable** into a regression model to account for the effects of multiple variables.
* The Cox regression model is a semi-parametric model that can be used to fit univariable and multivariable regression models that have survival outcomes.
```{r cox $ hazard ratio F1 Experiment #2}
library(survminer) # ggforect hazard ratio plot

#  Experimetn 1 all data 
Exp2_coxph_test         <- coxph(formula = Surv(Age, Count_dead) ~ Larvae_pCO2+(1|Tank), data = binary_Exp2_larvae)
Exp2_hazardratio_plot   <- ggforest(Exp2_coxph_test, data = binary_Exp2_larvae)
Exp2_coxph_test_table   <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2+(1|Tank), data = binary_Exp2_larvae) %>%  tbl_regression(exp = TRUE) 
#                              coef exp(coef) se(coef)      z       p
# Larvae_pCO2Moderate pCO2 -0.08592   0.91767  0.01963 -4.377 1.2e-05
# Larvae_pCO2High pCO2      0.46739   1.59583  0.02121 22.036 < 2e-16
# 1 | TankTRUE                   NA        NA  0.00000     NA      NA
# 
# Likelihood ratio test=655  on 2 df, p=< 2.2e-16
# n= 16145, number of events= 14310


# mean data - no tank 
Exp2_coxph_testMEANS         <- coxph(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp2_larvaeMEANS)
Exp2_hazardratio_plotMEANS   <- ggforest(Exp2_coxph_testMEANS, data = binary_Exp2_larvaeMEANS)
Exp2_coxph_test_tableMEANS   <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp2_larvaeMEANS) %>%  tbl_regression(exp = TRUE) 
#                             coef exp(coef) se(coef)     z      p
# Larvae_pCO2Moderate pCO2 -0.0869    0.9168   0.0439 -1.98 0.0477
# Larvae_pCO2High pCO2      0.4650    1.5920   0.0474  9.81 <2e-16
# 
# Likelihood ratio test=130.3  on 2 df, p=< 2.2e-16
# n= 3229, number of events= 2862


# output yo stuff homie!
capture.output(Exp2_coxph_test,file="Output/1_Survival_Growth/Experiment2_Survival_CoxReg.doc")
capture.output(Exp2_coxph_testMEANS,file="Output/1_Survival_Growth/Experiment2_Survival_CoxRegMEANS.doc")


pdf("Output/1_Survival_Growth/Experiment2_HazardRatio.pdf", width = 10, height = 6) # three figures
ggarrange(
  Exp2_hazardratio_plot,
  Exp2_hazardratio_plotMEANS,
  nrow = 2
)
dev.off()

```






### Experiment 4 F2s

```{r format F2 Experiment #4}
# create new columns for live and dead based on the count and proportion (of alive) tobackcalculate dead - these are integers! 
df_Exp4 <- df_Exp4 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(Count_dead = case_when(proportion == 0 ~ 0,
                                               proportion >0 ~ (1/as.numeric(proportion))*
                                                 as.numeric(Count_alive))) # proportion is that of alive
```

```{r convert to binary F2 Experiment #4}

# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not suitable for plotting

# Call the cumulative dataframe that we will write to in the for loop below
binary_Exp4_all     <- data.frame() # start dataframe  - all
# run it 
for (i in 1:nrow(df_Exp4)) {
   count_alive <- round(df_Exp4[i,]$Count_alive/1000) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_Exp4[i,]$Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   #count_alive <- round(df_Exp4[i,]$Count_alive) # raw counts
   #count_dead  <- round(df_Exp4[i,]$Count_dead)  # raw counts
   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_Exp4[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_Exp4[i,]$Larvae_pCO2,
                            Generation  = df_Exp4[i,]$Generation,
                            Tank        = df_Exp4[i,]$Tank,  
                            Age         = df_Exp4[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF           <- data.frame(loopDF) # name dataframe for this single row
  binary_Exp4_all <- rbind(binary_Exp4_all,loopDF) #bind to a cumulative list dataframe
}
# To address tank as a random factor it MUST be numeric, change this here
unique(binary_Exp4_all$Tank) # "A" "B" "C" "D" "E" NA
binary_Exp4_all <- binary_Exp4_all %>% 
                    dplyr::mutate(Generation = 2,
                                  Tank = case_when(Tank== 'A'~1,
                                                   Tank== 'B'~2,
                                                   Tank== 'C'~3)) %>% 
                    na.omit()


# (2) Mean counts alive and dead by tank - better representated for plots!
# difference here is we take a mean for counts alive and dead within treatment, no tank included and a singluar line for plotting 
df_Exp4_MEANS <- df_Exp4 %>% 
              dplyr::select(-c(Generation, Tank, Date, Percent_Survival, proportion)) %>% 
              dplyr::group_by(Larvae_pCO2, Age) %>% 
              dplyr::summarise(mean_Count_alive = mean(Count_alive), 
                               mean_Count_dead = mean(Count_dead))

# call datafarme binary for plotting
binary_Exp4_MEANS       <- data.frame() # start dataframe 

# run it 
for (i in 1:nrow(df_Exp4_MEANS)) {
   count_alive <- round(df_Exp4_MEANS[i,]$mean_Count_alive/1000) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_Exp4_MEANS[i,]$mean_Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   loopDF           <- data.frame(matrix(0,ncol = 3, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Larvae_pCO2','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Larvae_pCO2 = df_Exp4_MEANS[i,]$Larvae_pCO2,
                            Age         = df_Exp4_MEANS[i,]$Age
                          ) 
  loopDF[c(1:count_dead),3] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),3] = 0 # 0 for alive

  loopDF             <- data.frame(loopDF) # name dataframe for this single row
  binary_Exp4_MEANS <- rbind(binary_Exp4_MEANS,loopDF) #bind to a cumulative list dataframe
  
}
binary_Exp4_MEANS <- binary_Exp4_MEANS %>% 
                      dplyr::mutate(Generation = 2) %>% 
                      na.omit()

```

**note**: larval period ended at first sign of metamorphossi at 14 DPF  - truncate before this age to compare
```{r filter larval period F2 Experiment #4}

binary_Exp4_larvae      <- binary_Exp4_all %>% filter(Age <=14) # F1 - aligned timepoint prior to F1 metamorph for Gen 1 and 2 comparisons

binary_Exp4_larvaeMEANS <- binary_Exp4_MEANS %>% filter(Age <=14) # F1 - aligned timepoint prior to F1 metamorph for Gen 1 and 2 comparisons

```

```{r surv object F2 Experiment #4}

survExp4_all <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp4_larvaeMEANS)

# convert Age to the proportion of time to metamorphosis 
binary_Exp4_larvaeMEANS$Prop_time_metamophosis <- (binary_Exp4_larvaeMEANS$Age / 14)

survExp4          <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp4_larvaeMEANS)
survExp4_proptime <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Larvae_pCO2, data = binary_Exp4_larvaeMEANS)
```

```{r plot Kaplan Meier F2 Experiment #4}

Exp4_larvaeKaplanMeier <- survExp4 %>% 
                        ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                        scale_color_manual(values = c("forestgreen","orange","purple")) +
                        scale_fill_manual(values = c("forestgreen","orange","purple")) +
                        labs(
                          x = "Days",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        add_risktable() +
                        add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:10)) #+
                       # add_pvalue(location  = "annotation", x = 8.5)

Exp4_larvaeKaplanMeier <- survExp4_proptime %>% 
                        ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
                        scale_color_manual(values = c("forestgreen","orange","purple")) +
                        scale_fill_manual(values = c("forestgreen","orange","purple")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                       # add_pvalue(location  = "annotation", x = 8.5)
# plot the surv data
pdf("Output/1_Survival_Growth/Expeirment4_Survival_KaplanMeier.pdf", width=5, height = 4)
Exp4_larvaeKaplanMeier
dev.off()

Exp4_14d_surv_prob <- survExp4 %>% 
                          tbl_survfit(
                            times = 14,
                            label_header = "**survival to 10 days post-fertilization (95% CI)**"
                          )
```

# Comparing survival times between groups (log-rank test)
* We can conduct between-group significance tests using a log-rank test. The log-rank test equally weights observations over the entire follow-up time and is the most common way to compare survival times between groups. There are versions that more heavily weight the early or late follow-up that could be more appropriate depending on the research question (see ?survdiff for different test options).
* We get the log-rank p-value using the survdiff function
```{r log-rank test survdiff F2 Experiment #4}

Exp4_logrank_test <- survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp4_larvaeMEANS)
#                             N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2      334      250      237     0.689       1.9
# Larvae_pCO2=Moderate pCO2 481      335      348     0.470       1.9

```

# The Cox regression model
* We may want to quantify an effect size for a single variable, or include **more than one variable** into a regression model to account for the effects of multiple variables.
* The Cox regression model is a semi-parametric model that can be used to fit univariable and multivariable regression models that have survival outcomes.
```{r cox & hazard ratio F2 Experiment #4}
library(survminer) # ggforect hazard ratio plot

binary_Exp4_larvae$Larvae_pCO2 <- factor(binary_Exp4_larvae$Larvae_pCO2, exclude = 'High pCO2')
#  Experimetn 1 all data 
Exp4_coxph_test         <- coxph(formula = Surv(Age, Count_dead) ~ Larvae_pCO2+(1|Tank), data = binary_Exp4_larvae)
Exp4_hazardratio_plot   <- ggforest(Exp4_coxph_test, data = binary_Exp4_larvae)
Exp4_coxph_test_table   <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2+(1|Tank), data = binary_Exp4_larvae) %>%  tbl_regression(exp = TRUE) 
#                              coef exp(coef) se(coef)      z       p
# Larvae_pCO2Moderate pCO2 -0.12657   0.88111  0.04822 -2.625 0.00867
# 
# Likelihood ratio test=6.84  on 1 df, p=0.008898
# n= 2451, number of events= 1761 

binary_Exp4_larvaeMEANS$Larvae_pCO2 <- factor(binary_Exp4_larvaeMEANS$Larvae_pCO2, exclude = 'High pCO2')
# mean data - no tank 
Exp4_coxph_testMEANS         <- coxph(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp4_larvaeMEANS)
Exp4_hazardratio_plotMEANS   <- ggforest(Exp4_coxph_testMEANS, data = binary_Exp4_larvaeMEANS)
Exp4_coxph_test_tableMEANS   <- coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_Exp4_larvaeMEANS) %>%  tbl_regression(exp = TRUE) 
#                              coef exp(coef) se(coef)      z     p
# Larvae_pCO2Moderate pCO2 -0.13089   0.87732  0.08366 -1.564 0.118
# 
# Likelihood ratio test=2.43  on 1 df, p=0.119
# n= 815, number of events= 585 



# output yo stuff homie!
capture.output(Exp4_coxph_test,file="Output/1_Survival_Growth/Experiment4_Survival_CoxReg.doc")
capture.output(Exp4_coxph_testMEANS,file="Output/1_Survival_Growth/Experiment4_Survival_CoxRegMEANS.doc")


pdf("Output/1_Survival_Growth/Experiment4_HazardRatio.pdf", width = 10, height = 6) # three figures
ggarrange(
  Exp4_hazardratio_plot,
  Exp4_hazardratio_plotMEANS,
  nrow = 2
)
dev.off()


```




### Experiment 2 and 4 MERGE DATA

**about**: we have parallel datasets to use in a signle manuscript - merge the binary files and run stats / plots

**note**: the proportion time to metamorphosis column is *already accounted for* with the different times to metamorphossis!

* NOTE: omit the 'high' treatment in F1 as this was not present in F2 larvae
```{r merge Experiments 2 and 4}
# all - with tank replicate included
binary_Exp2_Exp4_all             <- rbind(binary_Exp2_larvae, binary_Exp4_larvae) %>% 
                                         dplyr::filter(!(Larvae_pCO2 == 'High pCO2')) %>% 
                                         dplyr::mutate(Prop_time_metamophosis = 
                                                            case_when(Generation == 1 ~ (Age / 10),
                                                                      Generation == 2 ~ (Age / 14))) %>% 
                                         # omit timpoints after 10 days for Gen 1 and after 14 days for Gen 2
                                         dplyr::filter(!Prop_time_metamophosis >1) 
binary_Exp2_Exp4_all$Larvae_pCO2 <- factor(binary_Exp2_Exp4_all$Larvae_pCO2, exclude = 'High pCO2')

#  by mean tank
binary_Exp2_Exp4_MEANS             <- rbind(binary_Exp2_larvaeMEANS, binary_Exp4_larvaeMEANS ) %>% 
                                           dplyr::filter(!(Larvae_pCO2 == 'High pCO2')) %>% 
                                           dplyr::mutate(Prop_time_metamophosis = 
                                                            case_when(Generation == 1 ~ (Age / 10),
                                                                      Generation == 2 ~ (Age / 14))) %>% 
                                         # omit timpoints after 10 days for Gen 1 and after 14 days for Gen 2
                                         dplyr::filter(!Prop_time_metamophosis >1) 
binary_Exp2_Exp4_MEANS$Larvae_pCO2 <- factor(binary_Exp2_Exp4_MEANS$Larvae_pCO2, exclude = 'High pCO2')


# sanity check
binary_Exp2_Exp4_all %>% dplyr::filter(Prop_time_metamophosis > 1) # should be no data!
binary_Exp2_Exp4_MEANS %>% dplyr::filter(Prop_time_metamophosis > 1) # should be no data!


# Subsets to explore further 
binary_Exp2_Exp4_larvaeLOW       <- binary_Exp2_Exp4_all %>% filter(Larvae_pCO2 %in% "Low pCO2")
binary_Exp2_Exp4_MEANS_larvaeLOW <- binary_Exp2_Exp4_MEANS %>% filter(Larvae_pCO2 %in% "Low pCO2") 

binary_Exp2_Exp4_larvaeMOD       <- binary_Exp2_Exp4_all %>% filter(Larvae_pCO2 %in% "Moderate pCO2")
binary_Exp2_Exp4_MEANS_larvaeMOD <- binary_Exp2_Exp4_MEANS %>% filter(Larvae_pCO2 %in% "Moderate pCO2") 

# what days are included?
unique(binary_Exp2_larvae$Age) # 2  4  8 10
unique(binary_Exp4_larvae$Age) # 2  6  8 10
unique(binary_Exp2_Exp4_all$Age) #  2  4  8 10  6 14

unique(binary_Exp2_larvaeMEANS$Age) # 2  4  8 10
unique(binary_Exp4_larvaeMEANS$Age) # 2  6  8 10
unique(binary_Exp2_Exp4_MEANS$Age) #  2  4  8 10  6 14

```

## Surv object with all data v. mean data
* note you can see the CI for the 'all' data surv object is small - these are 0s and 1s for each
* note there is a large CI for the 'mean' data 
```{r surv object merge F1 & F2}


s1_2          <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2+Generation, data = binary_Exp2_Exp4_MEANS)
s1_2_proptime <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Larvae_pCO2+Generation, data = binary_Exp2_Exp4_MEANS)


s1_2LOW          <- survfit2(Surv(Age, Count_dead) ~ Generation, data = binary_Exp2_Exp4_MEANS_larvaeLOW)
s1_2proptimeLOW  <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Generation, data = binary_Exp2_Exp4_MEANS_larvaeLOW)

s1_2MOD          <- survfit2(Surv(Age, Count_dead) ~ Generation, data = binary_Exp2_Exp4_MEANS_larvaeMOD)
s1_2proptimeMOD  <- survfit2(Surv(Prop_time_metamophosis, Count_dead) ~ Generation, data = binary_Exp2_Exp4_MEANS_larvaeMOD)
```


```{r plot Kaplan Meier merge F1 & F2}

Exp2_Exp4_larvaeKaplanMeier <- s1_2_proptime %>% 
                        ggsurvfit(linetype_aes = FALSE, linewidth = 1) +
                        scale_color_manual(values = c("green","forestgreen",
                                                      "orange","darkorange4")) +
                        scale_fill_manual(values = c("green","forestgreen",
                                                     "orange","darkorange4")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                     #   add_pvalue(location  = "annotation", x = 8.5)

# output plot 
pdf(paste0("Output/1_Survival_Growth/Experiments2&4_KaplanMeier.pdf"),width=8, height=4)
print(Exp2_Exp4_larvaeKaplanMeier)
dev.off()


Exp2_Exp4_larvaeKaplanMeierLOW <- s1_2proptimeLOW %>% 
                        ggsurvfit(linetype_aes = FALSE, linewidth = 1) +
                        scale_color_manual(values = c("green","forestgreen")) +
                        scale_fill_manual(values = c("green","forestgreen")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                     #   add_pvalue(location  = "annotation", x = 8.5)
# output plot 
pdf(paste0("Output/1_Survival_Growth/Experiments2&4_KaplanMeier_LowpCO2.pdf"),width=8, height=4)
print(Exp2_Exp4_larvaeKaplanMeierLOW)
dev.off()



Exp2_Exp4_larvaeKaplanMeierMOD <- s1_2proptimeMOD %>% 
                        ggsurvfit(linetype_aes = FALSE, linewidth = 1) +
                        scale_color_manual(values = c("orange","darkorange")) +
                        scale_fill_manual(values = c("orange","darkorange")) +
                        labs(
                          x = "Proportion time to metamorphosis",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:1)) #+
                     #   add_pvalue(location  = "annotation", x = 8.5)
# output plot 
pdf(paste0("Output/1_Survival_Growth/Experiments2&4_KaplanMeier_ModeratepCO2.pdf"),width=8, height=4)
print(Exp2_Exp4_larvaeKaplanMeierMOD)
dev.off()


```


```{r log-rank test survdiff  merge F1 & F2}

F1_F2_logrank_test <- survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2+Generation, data = binary_Exp2_Exp4_MEANS)
#                                            N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2, Generation=1      1340     1154      977      32.2      84.2
# Larvae_pCO2=Low pCO2, Generation=2       334      250      343      25.2      54.7
# Larvae_pCO2=Moderate pCO2, Generation=1 1084      946      861       8.3      20.3
# Larvae_pCO2=Moderate pCO2, Generation=2  481      335      504      56.7     131.7
# 
#  Chisq= 227  on 3 degrees of freedom, p= <2e-16

```


```{r cox & hazard ratio merge F1 & F2}


binary_Exp2_Exp4_MEANS$Generation <- as.character(binary_Exp2_Exp4_MEANS$Generation)
binary_Exp2_Exp4_MEANS$Gen_pCO2 <- paste0(as.character(binary_Exp2_Exp4_MEANS$Generation),'_',(binary_Exp2_Exp4_MEANS$Larvae_pCO2))


Exp2_Exp4_coxph_test <- coxph(formula = Surv(Prop_time_metamophosis, Count_dead) ~ Gen_pCO2, data = binary_Exp2_Exp4_MEANS)
Exp2_Exp4_hazardratio_plot <- ggforest(Exp2_Exp4_coxph_test, data = binary_Exp2_Exp4_MEANS)
Exp2_Exp4_coxph_test_table <- coxph(Surv(Prop_time_metamophosis, Count_dead) ~ Gen_pCO2, data = binary_Exp2_Exp4_MEANS) %>% tbl_regression(exp = TRUE) 
#                             coef exp(coef) se(coef)      z      p
# Gen_pCO21_Moderate pCO2 -0.09762   0.90700  0.04390 -2.224 0.0262
# Gen_pCO22_Low pCO2      -0.01664   0.98350  0.07004 -0.238 0.8122
# Gen_pCO22_Moderate pCO2 -0.14228   0.86738  0.06237 -2.281 0.0225
# 
# Likelihood ratio test=8.14  on 3 df, p=0.0432
# n= 3239, number of events= 2685 

capture.output(Exp2_Exp4_coxph_test,file="Output/1_Survival_Growth/Experiments2&4_Survival_CoxRegMEANS.doc")

pdf("Output/1_Survival_Growth/Experiments2&4_HazardRatio.pdf", width = 10, height = 3) # three figures
print(Exp2_Exp4_hazardratio_plot)
dev.off()



# Low gen 1 vs Low Gen 2
binary_Exp2_Exp4_MEANS_larvaeLOW$Generation <- as.character(binary_Exp2_Exp4_MEANS_larvaeLOW$Generation)

Exp2_Exp4_coxph_testLOW       <- coxph(formula = Surv(Prop_time_metamophosis, Count_dead) ~ Generation, 
                                       data = binary_Exp2_Exp4_MEANS_larvaeLOW)
Exp2_Exp4_hazardratio_plotLOW <- ggforest(Exp2_Exp4_coxph_testLOW, 
                                          data = binary_Exp2_Exp4_MEANS_larvaeLOW)
Exp2_Exp4_coxph_test_tableLOW <- coxph(Surv(Prop_time_metamophosis, Count_dead) ~ Generation, 
                                       data = binary_Exp2_Exp4_MEANS_larvaeLOW) %>% tbl_regression(exp = TRUE) 

                # coef exp(coef) se(coef)      z     p
# Generation2 -0.04029   0.96051  0.07030 -0.573 0.567

capture.output(Exp2_Exp4_coxph_testLOW,file="Output/1_Survival_Growth/Experiments2&4_Survival_CoxRegMEANS_LowpCO2.doc")

pdf("Output/1_Survival_Growth/Experiments2&4_HazardRatio_LowpCO2.pdf", width = 10, height = 3) # three figures
print(Exp2_Exp4_hazardratio_plotLOW)
dev.off()

# Mod gen 1 vs Low Gen 2
binary_Exp2_Exp4_MEANS_larvaeMOD$Generation <- as.character(binary_Exp2_Exp4_MEANS_larvaeMOD$Generation)

Exp2_Exp4_coxph_testMOD       <- coxph(formula = Surv(Prop_time_metamophosis, Count_dead) ~ Generation, 
                                       data = binary_Exp2_Exp4_MEANS_larvaeMOD)
Exp2_Exp4_hazardratio_plotMOD <- ggforest(Exp2_Exp4_coxph_testMOD, 
                                          data = binary_Exp2_Exp4_MEANS_larvaeMOD)
Exp2_Exp4_coxph_test_tableMOD <- coxph(Surv(Prop_time_metamophosis, Count_dead) ~ Generation, 
                                       data = binary_Exp2_Exp4_MEANS_larvaeMOD) %>% tbl_regression(exp = TRUE) 

                # coef exp(coef) se(coef)      z     p
# Generation2 -0.01658   0.98355  0.06429 -0.258 0.796

capture.output(Exp2_Exp4_coxph_testMOD,file="Output/1_Survival_Growth/Experiments2&4_Survival_CoxRegMEANS_ModeratepCO2.doc")

pdf("Output/1_Survival_Growth/Experiments2&4_HazardRatio_ModeratepCO2.pdf", width = 10, height = 3) # three figures
print(Exp2_Exp4_hazardratio_plotMOD)
dev.off()
```










```{r format F3 Experiment #6}
# create new columns for live and dead based on the count and proportion (of alive) tobackcalculate dead - these are integers! 
df_Exp6 <- df_Exp6 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(Count_dead = case_when(proportion == 0 ~ 0,
                                               proportion >0 ~ (1/as.numeric(proportion))*
                                                 as.numeric(Count_alive))) # proportion is that of alive
```






## Estimating x-time survival
* per 8 days from day 2 to 18 post fertilization
```{r}

F1_surv_prob <- s1_all %>% 
                          tbl_survfit(
                            times = 18,
                            label_header = "**survival to 20 days post-fertilization (95% CI)**")
F1_surv_prob$table_body$stat_1[2]
F1_surv_prob$table_body$stat_1[3]


# for loop every 50 days between 100 - 450 days post fertilization to a summary table!
x <- seq.int(2,18, by = 8) # call the integer
summarytable <- data.frame() # summary table 
for (i in 1:length(x)) { # for loop
  survtable <- s1_all %>% tbl_survfit(times = x[i])
  survtable$table_body$stat_1[2]
  survtable$table_body$stat_1[3]
  loopDF  <- data.frame(matrix(0,ncol = 4, nrow = 1)) 
  colnames(loopDF) = c('Age','LowpCO2_Probabilty','ModpCO2_Probabilty','Diff')
  loopDF  <- loopDF %>% 
        dplyr::mutate(
                      Age                = x[i],
                      LowpCO2_Probabilty = survtable$table_body$stat_1[2], 
                      ModpCO2_Probabilty = survtable$table_body$stat_1[3],
                      Diff               = abs(
                                           as.numeric(substr(survtable$table_body$stat_1[2],1,2)) -
                                           as.numeric(substr(survtable$table_body$stat_1[3],1,2)) 
                                           )
                      ) 
  loopDF       <- data.frame(loopDF) # name dataframe for this single row
  summarytable <- rbind(summarytable,loopDF) #bind to a cumulative list dataframe
}
```

# Comparing survival times between groups (log-rank test)
* We can conduct between-group significance tests using a log-rank test. The log-rank test equally weights observations over the entire follow-up time and is the most common way to compare survival times between groups. There are versions that more heavily weight the early or late follow-up that could be more appropriate depending on the research question (see ?survdiff for different test options).






```{r}
# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F1_all  <- Surv(time = as.numeric(binary_Exp2_all$Age), 
                     event = as.numeric(binary_Exp2_all$Count_dead), type = "right")

# plot the surv data
ggsurvplot(survfit(surv_F1_all ~ Larvae_pCO2, binary_Exp2_all), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")




# (2) Mean counts alive and dead by tank - better presentative for plots!
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F1_means_all  <- Surv(time = as.numeric(binary_Exp2_MEANS$Age), 
                     event = as.numeric(binary_Exp2_MEANS$Count_dead), type = "right")
# with the mean data by  tank replicate! better - save this one
ggsurvplot(survfit(surv_F1_means_all ~ Larvae_pCO2, binary_Exp2_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")

s1 <- survfit2(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS)
str(s1)

View(binary_df_F1_MEANS)
s1 %>% 
  ggsurvfit(linetype_aes = TRUE) +
  scale_color_manual(values = c("forestgreen","orange","purple")) +
  scale_fill_manual(values = c("forestgreen","orange","purple")) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(
    x = "Days",
    y = "Overall survival probability"
  )

?ggsurvfit
s1 %>% 
  ggsurvfit(linetype_aes = TRUE, linewidth = 1) +
  scale_color_manual(values = c("forestgreen","orange","purple")) +
  scale_fill_manual(values = c("forestgreen","orange","purple")) +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) + 
  add_confidence_interval() +
  add_risktable() +
  add_risktable_strata_symbol(symbol = "\U25CF", size = 10) +
  scale_ggsurvfit(x_scales = list(breaks = 0:10)) +
  add_pvalue(location  = "annotation", x = 8.5)

# use gtsummary 
s1 %>% 
  tbl_survfit(
    times = 10,
    label_header = "**survival to 10 days post-fertilization (95% CI)**"
  )
s1 %>% 
  tbl_survfit(
    probs = 0.5,
    label_header =  "**Median survival to 10 days post-fertilization (95% CI)**"
  )
s1
# Call: survfit(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS)
# 
#    2 observations deleted due to missingness 
#                              n events median 0.95LCL 0.95UCL
# Larvae_pCO2=Low pCO2      1844   1631      8       8       8
# Larvae_pCO2=Moderate pCO2 1460   1304      8       8       8
# Larvae_pCO2=High pCO2      805    762      4       4       4
survdiff(Surv(Age, Count_dead) ~ Larvae_pCO2 + (1|Tank), data = binary_df_F1_all)
binary_df_F1_all$Tank <- as.factor(as.numeric(binary_df_F1_all$Tank))
# Call:
# survdiff(formula = Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS)
# 
# n=4109, 2 observations deleted due to missingness.
# 
#                              N Observed Expected (O-E)^2/E (O-E)^2/V
# Larvae_pCO2=Low pCO2      1844     1631     1917     42.67    144.62
# Larvae_pCO2=Moderate pCO2 1460     1304     1372      3.38      8.63
# Larvae_pCO2=High pCO2      805      762      408    307.29    500.45


coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS)
coxph(Surv(Age, Count_dead) ~ Larvae_pCO2, data = binary_df_F1_MEANS) %>% 
  tbl_regression(exp = TRUE) 


?survfit2
Surv(lung$time, lung$status)[1:10]
Surv(binary_df_F1_MEANS$Age, binary_df_F1_MEANS$Count_dead)[1:10]
head(lung[, c("time", "status", "sex")])

# plot the surv data
pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F1_survival.pdf")
ggsurvplot(survfit(surv_F1_means_all ~ Larvae_pCO2, binary_df_F1_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")
dev.off()

```
### Survival status 
* run the stats with the 'all' data and address the Tank as random effect
```{r F1 run coxme on surv stats}

head(binary_df_F1_all)
# use this data from the first fr loop - has live and dead counts with reoslution of treatment AND replicate tank!!!

# now run the model
F1_survFIT_cox <- coxme(surv_F1_all ~ Larvae_pCO2 + (1|Tank), binary_df_F1_all)
summary(F1_survFIT_cox)
#                               coef exp(coef)   se(coef)     z p
# Larvae_pCO2Moderate pCO2 0.1969625  1.217698 0.01697219 11.61 0
# Larvae_pCO2High pCO2     1.1161450  3.053062 0.02143035 52.08 0



#The standard deviation of the tank random effect is very small (0.006017691) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (312321 ), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 


#pairwise comparisons

library(multcomp)
glht(F1_survFIT_cox, linfct = mcp(Larvae_pCO2 = "Dunnett"),alternative = "greater")
# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0   0.1059
# High pCO2 - Low pCO2 <= 0       0.8664


```

```{r F1 run coxph on surv stats}
?coxme
head(binary_df_F1_all)
# use this data from the first fr loop - has live and dead counts with reoslution of treatment AND replicate tank!!!

# now run the model
F1_survFIT_cox <- coxme(surv_F1_all ~ Larvae_pCO2 + (1|Tank), binary_df_F1_all)
summary(F1_survFIT_cox)

#                               coef exp(coef)   se(coef)     z p
# Larvae_pCO2Moderate pCO2 0.1969625  1.217698 0.01697219 11.61 0
# Larvae_pCO2High pCO2     1.1161450  3.053062 0.02143035 52.08 0



#The standard deviation of the tank random effect is very small (0.006017691) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (312321 ), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 


#pairwise comparisons

library(multcomp)
?glht
glht(F1_survFIT_cox, linfct = mcp(Larvae_pCO2 = "Dunnett"),alternative = "greater")
# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0   0.1059
# High pCO2 - Low pCO2 <= 0       0.8664


```



```{r F2 build surv object and plot em}


# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F2_all  <- Surv(time = as.numeric(binary_df_F2_all$Age), 
                     event = as.numeric(binary_df_F2_all$Count_dead), type = "right")
# plot the surv data
ggsurvplot(survfit(surv_F2_all ~ Larvae_pCO2, binary_df_F2_all), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")




# (2) Mean counts alive and dead by tank - better presentative for plots!
# build the surv object! 
# with tank included in the data frame (first for loop above)
surv_F2_means_all  <- Surv(time = as.numeric(binary_df_F2_MEANS$Age), 
                     event = as.numeric(binary_df_F2_MEANS$Count_dead), type = "right")
# with the mean data by  tank replicate! better - save this one
ggsurvplot(survfit(surv_F2_means_all ~ Larvae_pCO2, binary_df_F2_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange"), c(1,0.8)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")

# plot the surv data
pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F2_survival.pdf")
ggsurvplot(survfit(surv_F2_means_all ~ Larvae_pCO2, binary_df_F2_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange","purple"), c(1,0.8, 0.3)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "treatment")
dev.off()

```

```{r F2 run coxme on surv stats}

head(binary_df_F2_all)
# use this data from the first fr loop - has live and dead counts with reoslution of treatment AND replicate tank!!!

# omit the level high pCO2 because it is not actually present - to be a carry over from the full dataset?
is.na(binary_df_F2_all$Larvae_pCO2) <- binary_df_F2_all$Larvae_pCO2 == "High pCO2"
binary_df_F2_all$Larvae_pCO2 <- factor(binary_df_F2_all$Larvae_pCO2)
str(binary_df_F2_all) # Larvae_pCO2 now has two factors - correct!!

# now run the model
F2_survFIT_cox <- coxme(surv_F2_all ~ Larvae_pCO2 + (1|Tank), binary_df_F2_all)
summary(F2_survFIT_cox)

#                               coef exp(coef)   se(coef)   z      p
# Larvae_pCO2Moderate pCO2 0.1269179  1.135324 0.04536691 2.8 0.0051


#The standard deviation of the tank random effect is very small (0.006017691) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (312321 ), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 


#pairwise comparisons

library(multcomp)

glht(F2_survFIT_cox, linfct = mcp(Larvae_pCO2 = "Dunnett"),alternative = "greater")

# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0   0.1269


```



# F1 and F2 responses across generations 
  * truncate data to day 14 that overlaps between both datasets 
  * plot and run stats 
```{r}


# filter and merge up to d14 (common between datasets)

# all data
unique(binary_df_F1_all$Age)
binary_df_F1_d10 <- binary_df_F1_all %>% dplyr::filter(!Age %in% c(14,18)) %>% na.omit()
unique(binary_df_F1_d10$Age) #  2  4  8 10
unique(binary_df_F2_all$Age)
binary_df_F2_d10 <- binary_df_F2_all %>% dplyr::filter(!Age %in% c(14,21)) %>% na.omit()
unique(binary_df_F2_d10$Age) # 2  6  8 10

binary_F1_F2_master   <- rbind(binary_df_F1_d10, binary_df_F2_d10) %>% dplyr::filter(!(Generation ==1 & 
                                                                                       Larvae_pCO2 %in% 'High pCO2'))
unique(binary_F1_F2_master$Larvae_pCO2)
binary_F1_F2_moderate <- binary_F1_F2_master %>% dplyr::filter(Larvae_pCO2 %in% 'Moderate pCO2')
binary_F1_F2_low      <- binary_F1_F2_master %>% dplyr::filter(Larvae_pCO2 %in% 'Low pCO2')

# by means
unique(binary_df_F2_MEANS$Age)
binary_df_F1_d10 <- binary_df_F1_MEANS %>% dplyr::filter(!Age %in% c(14,18)) %>% na.omit()
unique(binary_df_F1_d10$Age) #  2  4  8 10
unique(binary_df_F2_MEANS$Age)
binary_df_F2_d10 <- binary_df_F2_MEANS %>% dplyr::filter(!Age %in% c(14,21)) %>% na.omit()
unique(binary_df_F2_d10$Age) # 2  6  8 10

binary_F1_F2_masterMEANS   <- rbind(binary_df_F1_d10, binary_df_F2_d10) %>% dplyr::filter(!(Generation ==1 & 
                                                                                       Larvae_pCO2 %in% 'High pCO2'))
binary_F1_F2_moderateMEANS <- binary_F1_F2_masterMEANS %>% dplyr::filter(Larvae_pCO2 %in% 'Moderate pCO2')
binary_F1_F2_lowMEANS      <- binary_F1_F2_masterMEANS %>% dplyr::filter(Larvae_pCO2 %in% 'Low pCO2')



# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting
# build the surv object! 
# with tank included in the data frame (first for loop above)


surv_F1_F2_allMEANS  <- Surv(time = as.numeric(binary_F1_F2_masterMEANS$Age), 
                     event = as.numeric(binary_F1_F2_masterMEANS$Count_dead), type = "right")


surv_F1_F2_all  <- Surv(time = as.numeric(binary_F1_F2_master$Age), 
                     event = as.numeric(binary_F1_F2_master$Count_dead), type = "right")
surv_F1_F2_modMEANS  <- Surv(time = as.numeric(binary_F1_F2_moderateMEANS$Age), 
                     event = as.numeric(binary_F1_F2_moderateMEANS$Count_dead), type = "right")
surv_F1_F2_lowMEANS  <- Surv(time = as.numeric(binary_F1_F2_lowMEANS$Age), 
                     event = as.numeric(binary_F1_F2_lowMEANS$Count_dead), type = "right")
# plot the surv data
# all 
ggsurvplot(survfit(surv_F1_F2_all ~ Generation+Larvae_pCO2, binary_F1_F2_masterMEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("yellowgreen","orange","darkgreen","darkorange"), 
                             c(1,0.8,1,0.8)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "Generation")
# low
# moderate
ggsurvplot(survfit(surv_F1_F2_modMEANS ~ Generation, binary_F1_F2_moderateMEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("darkorange","orange"), c(1,0.8)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,
           xlab = "Age (days post fertilization)", 
           legend.title = "Generation")
# low
ggsurvplot(survfit(surv_F1_F2_lowMEANS ~ Generation, binary_F1_F2_lowMEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("darkgreen","green"), c(1,0.8)), 
           ggtheme = theme_survminer(),
           break.x.by = 2 ,xlab = "Age (days post fertilization)", 
           legend.title = "Generation")


# Estimate median survival time 
survfit(Surv(as.numeric(Generation), Larvae_pCO2) ~ 1, data = binary_F1_F2_master)
survfit(Surv(as.numeric(Generation), Larvae_pCO2) ~ 1, data = binary_F1_F2_master) %>% 
  tbl_survfit(
    probs = 0.5,
    label_header = "**Median survival (95% CI)**"
  )

# statistics moderate
binary_F1_F2_master$Generation <- factor(binary_F1_F2_master$Generation)
str(binary_F1_F2_master) # Larvae_pCO2 now has two factors - correct!!
F1F2_survFIT_all_cox <- coxme(surv_F1_F2_all ~ Generation + (1|Tank), binary_F1_F2_master)
summary(F1F2_survFIT_Moderate_cox)


# as factor for the Gneration (the ID for this model)
binary_F1_F2_moderate$Generation <- factor(binary_F1_F2_moderate$Generation)
str(binary_F1_F2_moderate) # Larvae_pCO2 now has two factors - correct!!
# now run the model

survdiff(Surv(time, status) ~ sex, data = lung)


F1F2_survFIT_Moderate_cox <- coxme(surv_F1_F2_mod ~ Generation + (1|Tank), binary_F1_F2_moderate)
summary(F1F2_survFIT_Moderate_cox)
# Fixed coefficients
#                 coef exp(coef)   se(coef)     z       p
# Generation -0.208019 0.8121916 0.03484548 -5.97 2.4e-09
library(multcomp)
summary(glht(F1F2_survFIT_Moderate_cox, linfct = mcp(Generation = "Dunnett"),alternative = "greater"))
# Linear Hypotheses:
#            Estimate Std. Error z value Pr(>z)
# 2 - 1 <= 0 -0.20802    0.03485   -5.97      1



# statistics loq

# as factor for the Gneration (the ID for this model)
binary_F1_F2_low$Generation <- factor(binary_F1_F2_low$Generation)
str(binary_F1_F2_low) # Larvae_pCO2 now has two factors - correct!!
# now run the model
F1F2_survFIT_Low_cox <- coxme(surv_F1_F2_low ~ Generation + (1|Tank), binary_F1_F2_low)
summary(F1F2_survFIT_Low_cox)
# Fixed coefficients
#                   coef exp(coef)   se(coef)     z       p
# Generation2 -0.2371305 0.7888884 0.03939462 -6.02 1.8e-09
library(multcomp)
summary(glht(F1F2_survFIT_Low_cox, linfct = mcp(Generation = "Dunnett"),alternative = "greater"))
# Linear Hypotheses:
#            Estimate Std. Error z value Pr(>z)
# 2 - 1 <= 0 -0.23713    0.03939  -6.019      1


```
# F3 Larval Survival
## Plot it!
```{r F3 summary and plot}

stL_F3 <- summarySE(df_F3, measurevar="Percent_Survival", groupvars=c("Age", "Larvae_pCO2", "Parent_pCO2"))
stL_F3 %>% kbl() %>% kable_classic(full_width = F, html_font = "Cambria")

Gen3_PlotMeanSE <- ggplot(stL_F3, aes(x=Age, y=Percent_Survival, color=Larvae_pCO2))+ 
                        geom_line()+
                        geom_point()+ 
                        geom_errorbar(aes(ymin=Percent_Survival-se, ymax=Percent_Survival+se), width=.2,
                                      position=position_dodge(.1)) +
                        scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                        theme(legend.position="right", legend.direction="vertical", 
                              legend.title = element_blank())+ theme_bw() + 
                        labs(x="Age (days)", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                        ggtitle("F3 larvae survival (Mean +- SE)") + 
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              panel.background = element_blank(), 
                              axis.line = element_line(colour = "black"))+
                        facet_wrap("Parent_pCO2")
Gen3_PlotMeanSE


# Summarise Percent Deformities for plotting 
Gen3_mean_survival <- df_F3 %>% 
                          dplyr::select(Parent_pCO2,Larvae_pCO2,Age, Percent_Survival) %>% 
                          na.omit() %>% 
                          dplyr::group_by(Age, Parent_pCO2,Larvae_pCO2) %>% 
                          dplyr::summarise(mean_Perc_surv = mean(Percent_Survival), 
                                           n           = n(),
                                           sd_Perc_surv   = sd(Percent_Survival),
                                           se_Perc_surv   = sd_Perc_surv/(sqrt(n)))

# Barplots 
unique(Gen3_mean_survival$Age)
Gen3_barplots_Age2 <- Gen3_mean_survival %>% dplyr::filter(Age ==2) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 2 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)
Gen3_barplots_Age5 <- Gen3_mean_survival %>% dplyr::filter(Age ==5) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 5 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

Gen3_barplots_Age7 <- Gen3_mean_survival %>% dplyr::filter(Age ==7) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 7 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

Gen3_barplots_Age12 <- Gen3_mean_survival %>% dplyr::filter(Age ==12) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 12 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)
Gen3_barplots_Age16 <- Gen3_mean_survival %>% dplyr::filter(Age ==16) %>% 
                        ggplot() +
                        geom_errorbar(aes(x=Larvae_pCO2, 
                                           ymin=mean_Perc_surv-se_Perc_surv, 
                                           ymax=mean_Perc_surv+se_Perc_surv), 
                                       width=0, # removes the horizontal line
                                       colour="black", 
                                       size=1) +
                        geom_bar(aes(x=Larvae_pCO2, y=mean_Perc_surv,fill=factor(Larvae_pCO2)), 
                                  stat="identity",
                                 width = 0.75,
                                 alpha = 0.5) +
                        labs(title="F3 Age 16 DPF - Perc Survival (Mean +- SE)", 
                            x ="pCO2 Offspring Exposure", 
                            y = "Percent Survival") +
                         scale_fill_manual(breaks=c("Low pCO2", "Moderate pCO2", "High pCO2"), 
                                             values=c("forestgreen","orange", "purple")) +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(), 
                              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                              axis.text=element_text(size=12),
                              legend.position="none") +
                        facet_wrap(~Parent_pCO2)

# print the plot

pdf("Output/Survival/F3_survival_meanSE.pdf", width=10, height=5)
print(Gen3_PlotMeanSE_All)
dev.off()
# 
# 
# pdf("Output/Survival/PercSurvival_Gen3_2DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age2)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_5DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age5)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_7DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age7)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_12DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age12)
# dev.off()
# 
# pdf("Output/Survival/PercSurvival_Gen3_16DPF.pdf", width=5, height=5)
# print(Gen3_barplots_Age16)
# dev.off()

```

```{r F3 build binary live and dead column}
#make a surv object to pass into coxme
df_F3 <- df_F3 %>% 
          dplyr::mutate(Count = 
                          as.numeric(gsub(",","",Count))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::rename(Count_alive = Count) %>% # rename to avoid confusion
          dplyr::mutate(Count_dead = 
                          (1/as.numeric(proportion))*as.numeric(Count_alive)) # proportion is that of alive


# Run two for loops 

# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not surtable for plotting

# keep tank to use as a random effect in the cox model!
# Call the cumulative dataframe that we will write to in the for loop below
binary_df_F3_all     <- data.frame() # start dataframe 
binary_df_F3_high    <- data.frame() # start dataframe 
binary_df_F3_mod     <- data.frame() # start dataframe 
binary_df_F3_low     <- data.frame() # start dataframe 
# run it 
for (i in 1:nrow(df_F3)) {
   #count_alive <- round(df_F3[i,]$Count_alive/1000) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_F3[i,]$Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   count_alive <- round(df_F3[i,]$Count_alive) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F3[i,]$Count_dead)  # divide all counts by 1000, round to nearest

   
   loopDF           <- data.frame(matrix(0,ncol = 6, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Generation','Tank','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_F3[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_F3[i,]$Larvae_pCO2,
                            Generation  = df_F3[i,]$Generation,
                            Tank        = df_F3[i,]$Tank,  
                            Age         = df_F3[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),6] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),6] = 0 # 0 for alive

  loopDF       <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F3_all <- rbind(binary_df_F3_all,loopDF) #bind to a cumulative list dataframe
  # print(binary_df_F3) # print to monitor progress

     
  if (loopDF$Parent_pCO2[1] == 'High pCO2') {
      binary_df_F3_high <- rbind(binary_df_F3_high,loopDF) #bind to a cumulative list dataframe
  } else if (loopDF$Parent_pCO2[1] == 'Moderate pCO2') {
      binary_df_F3_mod <- rbind(binary_df_F3_mod,loopDF) #bind to a cumulative list dataframe
  } else {
      binary_df_F3_low <- rbind(binary_df_F3_low,loopDF) #bind to a cumulative list dataframe
    }
      
  
}

binary_df_F3_all$All_pCO2 <- as.factor(paste0(binary_df_F3_all$Parent_pCO2, ' x ', binary_df_F3_all$Larvae_pCO2))


# (2) Mean counts alive and dead by tank - better presentative for plots!

# for loop for plotting!! 
# difference here is we take a mean for counts alive and dead within treatment, no tank included and a singluar line for plotting 
df_F3_MEANS <- df_F3 %>% 
  dplyr::select(-c(Generation, Tank, Date, Percent_Survival, proportion)) %>% 
  dplyr::group_by(Parent_pCO2, Larvae_pCO2, Age) %>% 
  dplyr::summarise(mean_Count_alive = mean(Count_alive), 
            mean_Count_dead = mean(Count_dead))
# call datafarme binary for plotting
binary_df_F3_MEANS       <- data.frame() # start dataframe 
binary_df_F3_MEANS_high  <- data.frame() # start dataframe 
binary_df_F3_MEANS_mod   <- data.frame() # start dataframe 
binary_df_F3_MEANS_low   <- data.frame() # start dataframe 

# run it 
for (i in 1:nrow(df_F3_MEANS)) {
   #count_alive <- round(df_F3_MEANS[i,]$mean_Count_alive/1000) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_F3_MEANS[i,]$mean_Count_dead/1000)  # divide all counts by 1000, round to nearest
   
   count_alive <- round(df_F3_MEANS[i,]$mean_Count_alive) # divide all counts by 1000, round to nearest
   count_dead  <- round(df_F3_MEANS[i,]$mean_Count_dead)  # divide all counts by 1000, round to nearest

   
   loopDF           <- data.frame(matrix(0,ncol = 4, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Parent_pCO2','Larvae_pCO2','Age','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                            Parent_pCO2 = df_F3_MEANS[i,]$Parent_pCO2,
                            Larvae_pCO2 = df_F3_MEANS[i,]$Larvae_pCO2,
                            Age         = df_F3_MEANS[i,]$Age,
                          ) 
  loopDF[c(1:count_dead),4] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),4] = 0 # 0 for alive

  loopDF             <- data.frame(loopDF) # name dataframe for this single row
  binary_df_F3_MEANS <- rbind(binary_df_F3_MEANS,loopDF) #bind to a cumulative list dataframe
     
  if (loopDF$Parent_pCO2[1] == 'High pCO2') {
      binary_df_F3_MEANS_high <- rbind(binary_df_F3_MEANS_high,loopDF) #bind to a cumulative list dataframe
  } else if (loopDF$Parent_pCO2[1] == 'Moderate pCO2') {
      binary_df_F3_MEANS_mod <- rbind(binary_df_F3_MEANS_mod,loopDF) #bind to a cumulative list dataframe
  } else {
      binary_df_F3_MEANS_low <- rbind(binary_df_F3_MEANS_low,loopDF) #bind to a cumulative list dataframe
  }
    
}
```

```{r F3 build surv object}
nrow(binary_df_F3_all)
# create surv object
# with tank included in the data frame (first for loop above)
surv_F3_all  <- Surv(time = as.numeric(binary_df_F3_all$Age), 
                     event = as.numeric(binary_df_F3_all$Count_dead), type = "right")

surv_F3_high <- Surv(time = as.numeric(binary_df_F3_high$Age), 
                     event = as.numeric(binary_df_F3_high$Count_dead), type = "right")

surv_F3_mod  <- Surv(time = as.numeric(binary_df_F3_mod$Age), 
                     event = as.numeric(binary_df_F3_mod$Count_dead), type = "right")

surv_F3_low  <- Surv(time = as.numeric(binary_df_F3_low$Age), 
                     event = as.numeric(binary_df_F3_low$Count_dead), type = "right")


F3_low_survplot <- ggsurvplot(survfit(surv_F3_low ~ Larvae_pCO2, binary_df_F3_low), 
                       risk.table = F, pval = F , conf.int = TRUE,
                       palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8, 0.3)), 
                       ggtheme = theme_survminer(),
                       break.x.by = 2 ,xlab = "Age (days post fertilization)", 
                       legend.title = "treatment")

F3_mod_survplot <- ggsurvplot(survfit(surv_F3_mod ~ Larvae_pCO2, binary_df_F3_mod), 
                       risk.table = F, pval = F , conf.int = TRUE,
                       palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8, 0.3)), 
                       ggtheme = theme_survminer(),
                       break.x.by = 2 ,xlab = "Age (days post fertilization)", 
                       legend.title = "treatment")

F3_high_survplot <- ggsurvplot(survfit(surv_F3_high ~ Larvae_pCO2, binary_df_F3_high), 
                       risk.table = F, pval = F , conf.int = TRUE,
                       palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8, 0.3)), 
                       ggtheme = theme_survminer(),
                       break.x.by = 2 ,xlab = "Age (days post fertilization)", 
                       legend.title = "treatment")

F3_full_plot <- ggarrange(F3_low_survplot, F3_mod_survplot, F3_high_survplot, ncol = 3)
print(F3_full_plot)


# create surv object
# without tank included - means wihtin treatment SEPCIFICALLY FOR PLOTTING  second for loop above 
surv_F3_MEANS  <- Surv(time = as.numeric(binary_df_F3_MEANS$Age), 
                     event = as.numeric(binary_df_F3_MEANS$Count_dead), type = "right")


is.na(binary_df_F3_MEANS_low$Parent_pCO2) <- binary_df_F3_MEANS_low$Parent_pCO2 == "None"
binary_df_F3_MEANS_low$Parent_pCO2 <- factor(binary_df_F3_MEANS_low$Parent_pCO2)
str(binary_df_F3_MEANS_low) # youll see now its fixed

surv_F3_MEANS_low  <- Surv(time = as.numeric(binary_df_F3_MEANS_low$Age), 
                     event = as.numeric(binary_df_F3_MEANS_low$Count_dead), type = "right")

is.na(binary_df_F3_MEANS_mod$Parent_pCO2) <- binary_df_F3_MEANS_mod$Parent_pCO2 == "None"
binary_df_F3_MEANS_mod$Parent_pCO2 <- factor(binary_df_F3_MEANS_mod$Parent_pCO2)
str(binary_df_F3_MEANS_mod) # youll see now its fixed

surv_F3_MEANS_mod  <- Surv(time = as.numeric(binary_df_F3_MEANS_mod$Age), 
                     event = as.numeric(binary_df_F3_MEANS_mod$Count_dead), type = "right")


is.na(binary_df_F3_MEANS_high$Parent_pCO2) <- binary_df_F3_MEANS_high$Parent_pCO2 == "None"
binary_df_F3_MEANS_high$Parent_pCO2 <- factor(binary_df_F3_MEANS_high$Parent_pCO2)
str(binary_df_F3_MEANS_high) # youll see now its fixed

surv_F3_MEANS_high  <- Surv(time = as.numeric(binary_df_F3_MEANS_high$Age), 
                     event = as.numeric(binary_df_F3_MEANS_high$Count_dead), type = "right")

```

```{r F3 plot em}

pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_survival.pdf", width = 10)
ggsurvplot(survfit(surv_F3_all ~ Larvae_pCO2, binary_df_F3_all), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.2)), 
           break.x.by = 2 ,xlab = "Age (days post fertilization)",
           facet.by = "Parent_pCO2",
           legend.title = "treatment")
dev.off()

pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_survival.pdf", width = 10)
ggsurvplot(survfit(surv_F3_MEANS ~ Larvae_pCO2, binary_df_F3_MEANS), 
           risk.table = F, pval = F , conf.int = TRUE,
           palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.2)), 
           break.x.by = 2 ,xlab = "Age (days post fertilization)",
           facet.by = "Parent_pCO2",
           legend.title = "treatment")
dev.off()
# 
# pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_surival_low_history.pdf")
 ggsurvplot(survfit(surv_F3_MEANS_low ~ Larvae_pCO2, binary_df_F3_MEANS_low), 
            risk.table = F, pval = F , conf.int = TRUE,
            palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.3)), 
            break.x.by = 2 ,xlab = "Age (days post fertilization)", 
            legend.title = "treatment")
# dev.off()
# 
# pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_moderate_history.pdf")
ggsurvplot(survfit(surv_F3_MEANS_mod ~ Larvae_pCO2, binary_df_F3_MEANS_mod), 
            risk.table = F, pval = F , conf.int = TRUE,
            palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.2)), 
            break.x.by = 2 ,xlab = "Age (days post fertilization)", 
            legend.title = "treatment")
# dev.off()
# 
# pdf("C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Airradians_Larvae_crossgen_OA/RAnalysis/Output/Survival/F3_high_history.pdf")
# ggsurvplot(survfit(surv_F3_MEANS_high ~ Larvae_pCO2, binary_df_F3_MEANS_high), 
#            risk.table = F, pval = F , conf.int = TRUE,
#            palette = alpha(c("forestgreen","orange", "purple"), c(1,0.8,0.3)), 
#            break.x.by = 2 ,xlab = "Age (days post fertilization)", 
#            legend.title = "treatment")
# dev.off()

```
run cox proportional test
include tank as a random effect
```{r}

# remove "none" and the level - for some reason this lingered, dont know why
str(binary_df_F3_all) # before running, youll see PArent_pCO2 has 4 levels - we only want 3!
is.na(binary_df_F3_all$Parent_pCO2) <- binary_df_F3_all$Parent_pCO2 == "None"
binary_df_F3_all$Parent_pCO2 <- factor(binary_df_F3_all$Parent_pCO2)
str(binary_df_F3_all) # youll see now its fixed

# now run the model

survFitME_all_tank <- coxme(surv_F3_all ~ Larvae_pCO2*Parent_pCO2 + (1|Tank), binary_df_F3_all)
summary(survFitME_all_tank) 
#                                                           coef exp(coef)   se(coef)     z
# Larvae_pCO2Moderate pCO2                          -0.025569607 0.9747545 0.01848915 -1.38
# Larvae_pCO2High pCO2                               0.009866940 1.0099158 0.01873860  0.53
# Parent_pCO2Moderate pCO2                          -0.037194870 0.9634884 0.01855713 -2.00
# Parent_pCO2High pCO2                              -0.076000227 0.9268160 0.01845043 -4.12
# Larvae_pCO2Moderate pCO2:Parent_pCO2Moderate pCO2  0.030348803 1.0308140 0.02614267  1.16
# Larvae_pCO2High pCO2:Parent_pCO2Moderate pCO2      0.008771872 1.0088105 0.02630283  0.33
# Larvae_pCO2Moderate pCO2:Parent_pCO2High pCO2      0.035694517 1.0363392 0.02605141  1.37
# Larvae_pCO2High pCO2:Parent_pCO2High pCO2          0.006232918 1.0062524 0.02611643  0.24
# p
#  1.7e-01
#      6.0e-01
#   4.5e-02
#     3.8e-05
#  2.5e-01
#    7.4e-01
#  1.7e-01
#   8.1e-01

#The standard deviation of the tank random effect is very small (0.015) which suggests that tank does not really affect the outcome. Also the liklihood ratio test (Chi?sq) value is the same (527597), meaning that including tank as a random effect does not improve the model. We likely do not need to include it as a random effect. 

#pairwise comparisons

library(multcomp)

glht(survFitME_all_tank, linfct = mcp(Parent_pCO2 = "Dunnett"),alternative = "greater")

# Linear Hypotheses:
#                               Estimate
# Moderate pCO2 - Low pCO2 <= 0 -0.03719
# High pCO2 - Low pCO2 <= 0     -0.07600


```




# Stats
Lenght data do not meet the assumptions of noramlity even with an asin(sqrt()) transformation
```{r}
leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F1)
shapiro.test(df_F1$Percent_Survival_trans)

leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F2)
shapiro.test(df_F2$Percent_Survival_trans)

leveneTest(Percent_Survival_trans ~ Larvae_pCO2, data=df_F3)
shapiro.test(df_F3$Percent_Survival_trans)
```

#Non parametric Stats
### F1 Survival
```{r}
scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Age,
                data = df_F1)

dunnTest(Percent_Survival ~ Larvae_pCO2,
                data = df_F1,
              method="bh")

```

### F2 Survival
```{r}
scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Age,
                data = df_F2)

res.kruskal <- kruskal.test(Percent_Survival ~ Larvae_pCO2, data = df_F2)
res.kruskal

```

### F3 Survival
```{r}


scheirerRayHare(Percent_Survival ~ Larvae_pCO2*Parent_pCO2,
                data = df_F3)

dunnTest(Percent_Survival ~ Larvae_pCO2,
                data = df_F3,
              method="bh")

dunnTest(Percent_Survival ~ Parent_pCO2,
                data = df_F3,
              method="bh")
```